---
title: "Sensitivity analysis of constitutive sFLT1 secretion - fraction inhibition by genetic manipulation"
author: "Amy Gill"
date: "2023-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This file imports data from analysis of different fractional inhibition of each parameter starting from different base cases, performed 2/14/23 using the Matlab script "sFlt-model/04_sensitivity/sens_analysis.m" with run_mode = "sens_vary_base_inh_genchem", genetic_chemical = "genetic".

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/04_sensitivity/vary_base_vary_inh_chem/"
dir.create(outfile_prefix, recursive = TRUE)
```

## Load and wrangle files

### Import experimental data from CSV

```{r}
exp_df <- read_csv("../../../saved-data/karina/updated_inhibitor_data_2023-03-29/inh_data_wrangled.csv")
```

```{r}
# calculate mean treatment/control ratio for each variable (X,I) for each treatment type
mean_trt_ctrl_ratios <- exp_df %>%
  group_by(treatment_type, measured_var) %>%
  summarize(mean_ratio = mean(trt_ctrl_ratio))

# function to extract a particular mean ratio
pull_mean_ratio <- function(treatment, var){
  mean_trt_ctrl_ratios %>%
    filter(treatment_type == treatment & measured_var == var) %>%
    pull(mean_ratio)
}

# extract ratios to be used later
bref_cm_ratio <- pull_mean_ratio("Brefeldin", "sFlt1_CM")
bref_lys_ratio <- pull_mean_ratio("Brefeldin", "sFlt1_lysate")
tat_cm_ratio <- pull_mean_ratio("TATNSF", "sFlt1_CM")
tat_lys_ratio <- pull_mean_ratio("TATNSF", "sFlt1_lysate")
chp_cm_ratio <- pull_mean_ratio("Chlorpromazine", "sFlt1_CM")
chp_lys_ratio <- pull_mean_ratio("Chlorpromazine", "sFlt1_lysate")
chq_cm_ratio <- pull_mean_ratio("Chloroquine", "sFlt1_CM")
chq_lys_ratio <- pull_mean_ratio("Chloroquine", "sFlt1_lysate")
```

### Hornig experimental data

```{r}
# load experimental data from Hornig
load("../../../saved-data/hornig.rda")

hornig_basal <- hornig %>%
  filter(condition == "basal") %>%
  select(t = nominal_time, X = Sx_ng_ml)

hornig_basal
```

### Simulation input file

```{r}
# import file of input parameters
param_file <- readMat("../../../saved-data/04_sensitivity/vary_base_vary_inh_chem/input.mat")

# inspect available variables
names(param_file)
```

### Simulation output file

```{r}
# import file 
inh_file <- readMat("../../../saved-data/04_sensitivity/vary_base_vary_inh_chem/sens_diffbase_output.mat")

# inspect available variables
names(inh_file)
```

### Species names and parameter names

```{r}
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")
param_names <- unlist(inh_file$param.names)
```

### Extract beta values and full sets of base parameter values

```{r}
in_betas <- as.vector(param_file$betas)

base_param_mat <- param_file$p.base.all

# rows = parameter names
rownames(base_param_mat) <- param_names

# columns = parameter sets
colnames(base_param_mat) <- sapply(1:length(in_betas), function(i){
  paste0("set_", i)})

base_param_df <- as.data.frame(base_param_mat)
base_param_df$param = rownames(base_param_mat)

# doesn't auto-adjust ti changing # bases
base_param_df <- base_param_df %>%
  pivot_longer(set_1:set_5, names_to = "param_set", values_to = "base_param_value")
```


### Extract outputs to list

```{r}
inh_file_outlist <- inh_file$out[,,1]
names(inh_file_outlist) <- str_replace_all(names(inh_file_outlist), "\\.", "_")
names(inh_file_outlist)
```

```{r}
num_params <- length(param_names)
num_inhs <- length(inh_file$inh.frac)
num_bases <- length(param_file$betas) # number of base cases to test
```

### Extract time points

```{r}
time_vec <- inh_file$sim.time[1,]
```

### Extract baseline time courses for X, I


#### Extract for X

```{r}
X_base_tcs <- inh_file_outlist$X_base_tcs_all

colnames(X_base_tcs) <- sapply(1:length(in_betas), function(i){
  paste0("set_", i)})

X_base_df <- as.data.frame(X_base_tcs)

X_base_df$t <- time_vec

# does not adapt to changing number of bases - manually adjust
X_base_df <- X_base_df %>%
  pivot_longer(set_1:set_5, names_to = "base_params", values_to = "X_base")
```

#### Extract for Y

```{r}
I_base_tcs <- inh_file_outlist$I_base_tcs_all

colnames(I_base_tcs) <- sapply(1:length(in_betas), function(i){
  paste0("set_", i)})

I_base_df <- as.data.frame(I_base_tcs)

I_base_df$t <- time_vec

I_base_df <- I_base_df %>%
  pivot_longer(set_1:set_5, names_to = "base_params", values_to = "I_base")
```

#### Combine baseline time courses

```{r}
base_tc_df <- X_base_df %>%
  left_join(I_base_df, by = c("t", "base_params")) %>%
  right_join(base_param_df, join_by("base_params" == "param_set")) %>%
  mutate(param_altered = "none",
         param = paste0("base_", param)) %>%
  pivot_wider(names_from = "param", values_from = "base_param_value") %>%
  mutate(base_beta = factor(round(base_beta, 3)))

base_tc_df
```


### Generate or load inhibition time course data frame

#### Define function for Y_df extraction

```{r}
# format data frames of simulated species values for time courses
make_sim_Y_df <- function(Y_mat){
  
  # flatten matrix to data frame of time course data
  # (this isn't very "R-style", but it works, so tell Hadley to chill)
  for (i in 1:dim(Y_mat)[3]) {
    # print(i)
    if (i == 1) {    # first run: initialize output data frame
      # extract matrix for first parameter set to a data frame
      Y_temp <- as.data.frame(Y_mat[,,1])
      names(Y_temp) <- species_names
      Y_temp$run <- 1
        
      # initialize full Y data frame to the time course from this first run
      Y_df <- Y_temp
      
    } else {    # later runs: append to data frame
      Y_temp <- as.data.frame(Y_mat[,,i])
      names(Y_temp) <- species_names
      Y_temp$run <- i
      Y_df <- rbind(Y_df, Y_temp)
    }
  }
  
  # add matching time points
  Y_df <- Y_df %>%
    mutate(t = rep(time_vec, num_params * num_inhs * num_bases))
  
  Y_df
}
```


#### Generate or load Y_df

If Y_df has already been generated, load it. Otherwise, generate and save Y_df.

```{r}
Y_df_filename <- paste0(outfile_prefix, "Y_df_imported.rda")

if(file.exists(Y_df_filename)){
  load(Y_df_filename)
} else {
  # extract simulated data
  Y_df <- make_sim_Y_df(inh_file$Y.all)
  save(Y_df, file = paste0(outfile_prefix, "Y_df_imported.rda"))
}
```

### Annotate parameter alteration metadata

Order: For each base parameter set, for each parameter, test each inhibition fraction

```{r}
# annotate base parameter set
param_set <- unlist(lapply(1:num_bases, function(x){
  rep(paste0("set_", x), num_inhs * num_params)
}))

# annotate parameter altered
temp <- unlist(lapply(1:num_params, function(x){
  rep(param_names[x], num_inhs)
}))

param_altered <- rep(temp, num_bases)

# extract inhibition fraction
inh_frac <- rep(rep(inh_file$inh.frac[1,], num_params), num_bases)

param_meta_df <- data.frame(run = 1:(num_bases * num_params * num_inhs),
                            base_param_set = factor(param_set),
                            param_altered = factor(param_altered,
                                                  levels = c("alpha", "beta", "gamma",
                                                             "delta", "tau")), 
                           inh_frac = inh_frac)

param_meta_df
```

### Merge time course data and parameter inhibition metadata

```{r}
Y_df_filename_2 <- paste0(outfile_prefix, "Y_df_meta_added.rda")

if(file.exists(Y_df_filename_2)){
  load(Y_df_filename_2)
} else {

  Y_df <- Y_df %>%
    left_join(param_meta_df, by = "run")
  
  save(Y_df, file = paste0(outfile_prefix, Y_df_filename_2))
}
```

### Extract 72h values, normalize to baseline

```{r}
# all species at 72h in all conditions
Y_72h <- Y_df %>%
  filter(t == 72)

# baseline X at 72h
base_72h_df <- base_tc_df %>% 
  filter(t == 72) %>%
  select(base_param_set = base_params, X_base_t = X_base, I_base_t = I_base, base_beta)

Y_72h <- Y_72h %>%
  full_join(base_72h_df, by = "base_param_set") %>%
  mutate(X_norm = X/X_base_t,
         I_norm = I/I_base_t)
```

### Extract 18h values, normalize to baseline

```{r}
# all species at 18h in all conditions
Y_18h <- Y_df %>%
  filter(t == 18)

# baseline X at 18h
base_18h_df <- base_tc_df %>% 
  filter(t == 18) %>%
  select(base_param_set = base_params, X_base_t = X_base, I_base_t = I_base, base_beta)

Y_18h <- Y_18h %>%
  left_join(base_18h_df, by = "base_param_set") %>%
  mutate(X_norm = X/X_base_t,
         I_norm = I/I_base_t)

Y_18h
```

### Combined 18/72h values

```{r}
Y_combo <- rbind(Y_18h, Y_72h)
Y_combo
```

### WORKSPACE


#### Beta altered

```{r}
Y_18h_beta <- Y_18h %>%
  filter(param_altered == "beta")

# find beta inhibition value that maps to brefeldin effect on extracellular sFLT1
bref_beta_X_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$X_norm - bref_cm_ratio))]

# find beta inhibition value that maps to brefeldin effect on intracellular sFLT1
bref_beta_I_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$I_norm - bref_lys_ratio))]

# find beta inhibition value that maps to chlorpromazine effect on extracellular sFLT1
chp_beta_X_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$X_norm - chp_cm_ratio))]

# find beta inhibition value that maps to chlorpromazine effect on intracellular sFLT1
chp_beta_I_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$I_norm - chp_lys_ratio))]

tat_beta_X_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$X_norm - tat_cm_ratio))]

tat_beta_I_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$I_norm - tat_lys_ratio))]

```

#### Gamma altered

```{r}
Y_18h_gamma <- Y_18h %>%
  filter(param_altered == "gamma")

# find gamma inhibition value that maps to chloroquine effect on extracellular sFLT1
chq_gamma_X_inh_frac <- Y_18h_gamma$inh_frac[which.min(
  abs(Y_18h_gamma$X_norm - chq_cm_ratio))]

# find gamma inhibition value that maps to chloroquine effect on intracellular sFLT1
chq_gamma_I_inh_frac <- Y_18h_gamma$inh_frac[which.min(
  abs(Y_18h_gamma$I_norm - chq_lys_ratio))]
```















### Inhibitors of beta - find inh frac

```{r}
Y_18h_beta <- Y_18h %>%
  filter(param_altered == "beta") %>%
  mutate(bref_x_error = abs(X_norm - bref_cm_ratio),
         bref_i_error = abs(I_norm - bref_lys_ratio),
         chp_x_error = abs(X_norm - chp_cm_ratio),
         chp_i_error = abs(I_norm - chp_lys_ratio),
         tat_x_error = abs(X_norm - tat_cm_ratio),
         tat_i_error = abs(I_norm - tat_lys_ratio))

Y_18h

temp_list <- lapply(1:num_bases, function(x){
  temp_df <- Y_18h %>%
    filter(param_altered == "beta",
           base_param_set == paste0("set_", x))
  
  # find beta inhibition value that maps to brefeldin effect on extracellular sFLT1
  bref_beta_X_inh_frac <- temp_df$inh_frac[which.min(abs(temp_df$X_norm - bref_cm_ratio))]
  
  # find beta inhibition value that maps to brefeldin effect on intracellular sFLT1
  bref_beta_I_inh_frac <- temp_df$inh_frac[which.min(abs(temp_df$I_norm - bref_lys_ratio))]
  
  # find beta inhibition value that maps to chlorpromazine effect on extracellular sFLT1
  chp_beta_X_inh_frac <- temp_df$inh_frac[which.min(abs(temp_df$X_norm - chp_cm_ratio))]
  
  # find beta inhibition value that maps to chlorpromazine effect on intracellular sFLT1
  chp_beta_I_inh_frac <- temp_df$inh_frac[which.min(abs(temp_df$I_norm - chp_lys_ratio))]
  
  # find beta inhibition value that maps to STX6 effect on extracellular sFLT1
  tat_beta_X_inh_frac <- temp_df$inh_frac[which.min(abs(temp_df$X_norm - tat_cm_ratio))]
  
  # find beta inhibition value that maps to STX6 effect on intracellular sFLT1
  tat_beta_I_inh_frac <- temp_df$inh_frac[which.min(abs(temp_df$I_norm - tat_lys_ratio))]
  
  data.frame(inhibitor = c("Brefeldin", "Brefeldin", "Chlorpromazine", "Chlorpromazine", 
                           "TATNSF700", "TATNSF700"),
             base_param_set = paste0("set_", x),
             output = rep(c("X_inh_frac", "I_inh_frac"),3),
             inh_frac = c(bref_beta_X_inh_frac, bref_beta_I_inh_frac, 
                          chp_beta_X_inh_frac, chp_beta_I_inh_frac,
                          tat_beta_X_inh_frac, tat_beta_I_inh_frac)) %>%
    pivot_wider(names_from = "output", values_from = "inh_frac")
  
})

# need to write for an arbitrary number of divides - for now, manually update
inh_frac_df_beta <- rbind(temp_list[[1]], temp_list[[2]], temp_list[[3]], temp_list[[4]], temp_list[[5]])
```

### Inhibitors of gamma - find inh frac

```{r}
Y_18h_gamma <- Y_18h %>%
  filter(param_altered == "gamma") %>%
  mutate(chq_x_error = abs(X_norm - chq_cm_ratio),
         chq_i_error = abs(I_norm - chq_lys_ratio))

temp_list <- lapply(1:num_bases, function(x){
  temp_df <- Y_18h %>%
    filter(param_altered == "gamma",
           base_param_set == paste0("set_", x))
  
  # find beta inhibition value that maps to chloroquine effect on extracellular sFLT1
  chq_beta_X_inh_frac <- temp_df$inh_frac[which.min(abs(temp_df$X_norm - chq_cm_ratio))]
  
  # find beta inhibition value that maps to chloroquine effect on intracellular sFLT1
  chq_beta_I_inh_frac <- temp_df$inh_frac[which.min(abs(temp_df$I_norm - chq_lys_ratio))]
  
  
  
  data.frame(inhibitor = c("Chloroquine", "Chloroquine"),
             base_param_set = paste0("set_", x),
             output = c("X_inh_frac", "I_inh_frac"),
             inh_frac = c(chq_beta_X_inh_frac, chq_beta_I_inh_frac)) %>%
    pivot_wider(names_from = "output", values_from = "inh_frac")
  
})

# need to write for an arbitrary number of divides - for now, manually update
inh_frac_df_gamma <- rbind(temp_list[[1]], temp_list[[2]], temp_list[[3]], temp_list[[4]], temp_list[[5]])

inh_frac_df_gamma
```

```{r}
inh_frac_df <- rbind(inh_frac_df_beta, inh_frac_df_gamma)
inh_frac_df
```





## Base plots

### Base X

```{r}
base_X_tcs <- base_tc_df %>%
  ggplot(aes(t, X_base, col = base_beta)) +
  geom_line(linewidth = 2, alpha = 0.7) +
  xlab("Time (h)") +
  ylab("X (ng/mL)") +
  scale_x_continuous(breaks = seq(0, 72, by=24)) +
  theme_paper +
  scale_color_manual(values = cb_pal) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("base_X_tcs"),
       base_X_tcs, width = 6, height = 4)

base_X_tcs
```


### Base I

```{r}
base_I_tcs <- base_tc_df %>%
  ggplot(aes(t, I_base, col = base_beta)) +
  geom_line(linewidth = 2) +
  xlab("Time (h)") +
  ylab("I (#/cell)") +
  scale_x_continuous(breaks = seq(0, 72, by=24)) +
  scale_y_continuous(limits = c(0, 1.25e6), breaks = seq(0, 1.25e6, 2.5e5)) +
  theme_paper +
  scale_color_manual(values = cb_pal) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("base_I_tcs"),
       base_I_tcs, width = 6, height = 4)

base_I_tcs
```

### Combined X, I

```{r}
X_I_base_tcs <- (base_X_tcs + theme(legend.position = "none")) /
  (base_I_tcs + theme(legend.position = "bottom"))

# save plot
ggsave(out_png("X_I_base_tcs"),
       X_I_base_tcs, width = 7, height = 8)

X_I_base_tcs
```



## #### WORKSPACE ####

From here, plot 72h chemical inhibition and annotate with inhibitor results








## OLD


## 72h plots: fraction of initial amount 

### abgd, facet grid, X and I


```{r}
inhfrac_abgd_xi_grid_72h <- Y_72h %>%
  filter(param_altered != "tau") %>%
  pivot_longer(cols = c(X_norm, I_norm), names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "X", "I"),
         species = factor(species, levels = c("X", "I"))) %>%
  ggplot(aes(inh_frac, amount, color = base_beta)) +
      geom_hline(yintercept = 1, lty = 2, linewidth = 0.5) +
  geom_line(linewidth = 1, alpha = 0.9) +
  facet_grid(species ~ param_altered, scales = "free", labeller = label_parsed) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme_paper +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 1, by = 0.25)) +
  scale_y_continuous(limits = c(0, 4)) +
  scale_color_manual(values = cb_pal) +
  theme(panel.spacing = unit(1.5, "lines")) +
  labs(color = TeX(r'(Base $\beta$)'))

# save plot
ggsave(out_png("inhfrac_abgd_xi_grid_72h"),
       inhfrac_abgd_xi_grid_72h, width = 12, height = 6)

inhfrac_abgd_xi_grid_72h
```

### Individual plots

#### alpha, X

```{r}
X_72h_v_inh_frac_alpha <- Y_72h %>%
  filter(param_altered == "alpha") %>%
  ggplot(aes(inh_frac, X_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\alpha$)')) +
  ylab(TeX(r'($X_{72h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0,4), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_72h_v_inh_frac_alpha"),
       X_72h_v_inh_frac_alpha, width = 6, height = 4)

X_72h_v_inh_frac_alpha
```


#### alpha, I

```{r}
I_72h_v_inh_frac_alpha <- Y_72h %>%
  filter(param_altered == "alpha") %>%
  ggplot(aes(inh_frac, I_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\alpha$)')) +
  ylab(TeX(r'($I_{72h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0,2), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_72h_v_inh_frac_alpha"),
       I_72h_v_inh_frac_alpha, width = 6, height = 4)

I_72h_v_inh_frac_alpha
```


#### beta, X

```{r}
X_72h_v_inh_frac_beta <- Y_72h %>%
  filter(param_altered == "beta") %>%
  ggplot(aes(inh_frac, X_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\beta$)')) +
  ylab(TeX(r'($X_{72h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0,4), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_72h_v_inh_frac_beta"),
       X_72h_v_inh_frac_beta, width = 6, height = 4)

X_72h_v_inh_frac_beta
```


#### beta, I

```{r}
I_72h_v_inh_frac_beta <- Y_72h %>%
  filter(param_altered == "beta") %>%
  ggplot(aes(inh_frac, I_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\beta$)')) +
  ylab(TeX(r'($I_{72h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0, 2), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_72h_v_inh_frac_beta"),
       I_72h_v_inh_frac_beta, width = 6, height = 4)

I_72h_v_inh_frac_beta
```


#### gamma, X

```{r}
X_72h_v_inh_frac_gamma <- Y_72h %>%
  filter(param_altered == "gamma") %>%
  ggplot(aes(inh_frac, X_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\gamma$)')) +
  ylab(TeX(r'($X_{72h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0, 4), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_72h_v_inh_frac_gamma"),
       X_72h_v_inh_frac_gamma, width = 6, height = 4)

X_72h_v_inh_frac_gamma
```


#### gamma, I

```{r}
I_72h_v_inh_frac_gamma <- Y_72h %>%
  filter(param_altered == "gamma") %>%
  ggplot(aes(inh_frac, I_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\gamma$)')) +
  ylab(TeX(r'($I_{72h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0, 2), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_72h_v_inh_frac_gamma"),
       I_72h_v_inh_frac_gamma, width = 6, height = 4)

I_72h_v_inh_frac_gamma
```


#### delta, X

```{r}
X_72h_v_inh_frac_delta <- Y_72h %>%
  filter(param_altered == "delta") %>%
  ggplot(aes(inh_frac, X_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\delta$)')) +
  ylab(TeX(r'($X_{72h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0, 4), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_72h_v_inh_frac_delta"),
       X_72h_v_inh_frac_delta, width = 6, height = 4)

X_72h_v_inh_frac_delta
```


#### delta, I


```{r}
I_72h_v_inh_frac_delta <- Y_72h %>%
  filter(param_altered == "delta") %>%
  ggplot(aes(inh_frac, I_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\delta$)')) +
  ylab(TeX(r'($I_{72h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0, 2), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_72h_v_inh_frac_delta"),
       I_72h_v_inh_frac_delta, width = 6, height = 4)

I_72h_v_inh_frac_delta
```

Made an attempt to combine in patchwork, but the guides were not behaving nicely.

## 18h plots: fraction of initial amount 

### abgd, facet grid, X and I

```{r}
inhfrac_abgd_xi_grid_18h <- Y_18h %>%
  filter(param_altered != "tau") %>%
  #mutate(param_color = case_when(param_altered == "alpha" ~ param_pal_list$alpha,
  #                               param_altered == "beta" ~ param_pal_list$beta,
  #                               param_altered == "gamma" ~ param_pal_list$gamma,
  #                               param_altered == "delta" ~ param_pal_list$delta)) %>%
  pivot_longer(cols = c(X_norm, I_norm), names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "X", "I"),
         species = factor(species, levels = c("X", "I"))) %>%
  ggplot(aes(inh_frac, amount, color = base_beta)) +
      geom_hline(yintercept = 1, lty = 2, linewidth = 0.5) +
  geom_line(linewidth = 1, alpha = 0.9) +
  facet_grid(species ~ param_altered, scales = "free", labeller = label_parsed) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme_paper +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 1, by = 0.25)) +
  scale_y_continuous(limits = c(0,4)) +
  scale_color_manual(values = cb_pal) +
  theme(panel.spacing = unit(1.5, "lines")) +
  labs(color = TeX(r'(Base $\beta$)'))

# save plot
ggsave(out_png("inhfrac_abgd_xi_grid_18h"),
       inhfrac_abgd_xi_grid_18h, width = 12, height = 6)

inhfrac_abgd_xi_grid_18h
```

### Beta only - annotate inhibitors

#### Extracellular

```{r}
X_18h_v_inh_frac_beta <- Y_18h %>%
  filter(param_altered == "beta") %>%
  ggplot(aes(inh_frac, X_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\beta$)')) +
  ylab(TeX(r'($X_{18h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_18h_v_inh_frac_beta_nonotes"),
       X_18h_v_inh_frac_beta, width = 6, height = 4)

X_18h_v_inh_frac_beta
```

##### Annotate brefeldin

```{r}
# extract data frame of RAB27a inhibition fraction estimates for each base parameter set
bref_inh_frac_df <- inh_frac_df %>%
  filter(inhibitor == "Brefeldin")

X_18h_v_inh_frac_beta_bref <- X_18h_v_inh_frac_beta +
  geom_hline(yintercept = bref_cm_ratio, lty = 4) +
  annotate("text", x = 0.25, y = bref_cm_ratio + .05, label = "Brefeldin", size = 6) 

for (i in 1:num_bases) {
  
  if(bref_inh_frac_df$X_inh_frac[i] < .99){
    X_18h_v_inh_frac_beta_bref <- X_18h_v_inh_frac_beta_bref +
      annotate("segment", x = bref_inh_frac_df$X_inh_frac[i], xend = bref_inh_frac_df$X_inh_frac[i], 
               y = -Inf, yend = bref_cm_ratio, , color = cb_pal[i]) +  
      annotate("point", x = bref_inh_frac_df$X_inh_frac[i], y = bref_cm_ratio, 
               size = 5, pch = 21, fill = cb_pal[i], color = "black")
  }
  
}

# save plot
ggsave(out_png("X_18h_v_inh_frac_beta_bref"),
       X_18h_v_inh_frac_beta_bref, width = 6, height = 4)

X_18h_v_inh_frac_beta_bref
```

##### Annotate chlorpromazine

```{r}
# extract data frame of RAB27a inhibition fraction estimates for each base parameter set
chp_inh_frac_df <- inh_frac_df %>%
  filter(inhibitor == "Chlorpromazine")

X_18h_v_inh_frac_beta_chp <- X_18h_v_inh_frac_beta +
  geom_hline(yintercept = chp_cm_ratio, lty = 2) +
  annotate("text", x = 0.2, y = chp_cm_ratio + .05, label = "Chlorpromazine", size = 6) 

for (i in 1:num_bases) {
  
  if(chp_inh_frac_df$X_inh_frac[i] < .99){
    X_18h_v_inh_frac_beta_chp <- X_18h_v_inh_frac_beta_chp +
    annotate("segment", x = chp_inh_frac_df$X_inh_frac[i], xend = chp_inh_frac_df$X_inh_frac[i], 
           y = -Inf, yend = chp_cm_ratio, , color = cb_pal[i]) +  
    annotate("point", x = chp_inh_frac_df$X_inh_frac[i], y = chp_cm_ratio, 
             size = 5, pch = 21, fill = cb_pal[i], color = "black")
  }
  
}

# save plot
ggsave(out_png("X_18h_v_inh_frac_beta_chp"),
       X_18h_v_inh_frac_beta_chp, width = 6, height = 4)

X_18h_v_inh_frac_beta_chp
```

##### Annotate TATNSF700

```{r}
# extract data frame of RAB27a inhibition fraction estimates for each base parameter set
tat_inh_frac_df <- inh_frac_df %>%
  filter(inhibitor == "TATNSF700")

X_18h_v_inh_frac_beta_tat <- X_18h_v_inh_frac_beta +
  geom_hline(yintercept = tat_cm_ratio, lty = 3) +
  annotate("text", x = 0.2, y = tat_cm_ratio + .05, label = "TATNSF700", size = 6) 

for (i in 1:num_bases) {
  if(tat_inh_frac_df$X_inh_frac[i] < .99){
  
  X_18h_v_inh_frac_beta_tat <- X_18h_v_inh_frac_beta_tat +
    annotate("segment", x = tat_inh_frac_df$X_inh_frac[i], xend = tat_inh_frac_df$X_inh_frac[i], 
           y = -Inf, yend = tat_cm_ratio, , color = cb_pal[i]) +  
    annotate("point", x = tat_inh_frac_df$X_inh_frac[i], y = tat_cm_ratio, 
             size = 5, pch = 21, fill = cb_pal[i], color = "black")
  }
}

# save plot
ggsave(out_png("X_18h_v_inh_frac_beta_tat"),
       X_18h_v_inh_frac_beta_tat, width = 6, height = 4)

X_18h_v_inh_frac_beta_tat
```

##### Annotate all siRNAs - expt ratios only

```{r}
X_18h_v_inh_frac_beta_annot_3 <- X_18h_v_inh_frac_beta +
  geom_hline(yintercept = bref_cm_ratio, lty = 4) +
  geom_hline(yintercept = chp_cm_ratio, lty = 2) +
  geom_hline(yintercept = tat_cm_ratio, lty = 3) +
  annotate("text", x = 0.5, y = bref_cm_ratio - .05, label = "Brefeldin", size = 6) +
  annotate("text", x = 0.3, y = chp_cm_ratio - .05, label = "Chlorpromazine", size = 6) +
  annotate("text", x = 0.17, y = tat_cm_ratio + .05, label = "TATNSF700", size = 6) +
  scale_y_continuous(expand = c(.01, .01))

# save plot
ggsave(out_png("X_18h_v_inh_frac_beta_annot_3"),
       X_18h_v_inh_frac_beta_annot_3, width = 6, height = 4)

X_18h_v_inh_frac_beta_annot_3
```

#### Intracellular

```{r}
I_18h_v_inh_frac_beta <- Y_18h %>%
  filter(param_altered == "beta") %>%
  ggplot(aes(inh_frac, I_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\beta$)')) +
  ylab(TeX(r'($I_{18h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0,3.5), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_18h_v_inh_frac_beta_nonotes"),
       I_18h_v_inh_frac_beta, width = 6, height = 4)

I_18h_v_inh_frac_beta
```

##### Annotate brefeldin

```{r}
I_18h_v_inh_frac_beta_bref <- I_18h_v_inh_frac_beta +
  geom_hline(yintercept = bref_lys_ratio, lty = 4) +
  annotate("text", x = 0.25, y = bref_lys_ratio + .15, label = "Brefeldin", size = 6)

for (i in 1:num_bases) {
  
  if(bref_inh_frac_df$I_inh_frac[i] < .99){
    I_18h_v_inh_frac_beta_bref <- I_18h_v_inh_frac_beta_bref +
      annotate("segment", x = bref_inh_frac_df$I_inh_frac[i], xend = bref_inh_frac_df$I_inh_frac[i], 
               y = -Inf, yend = bref_lys_ratio, , color = cb_pal[i]) +  
      annotate("point", x = bref_inh_frac_df$I_inh_frac[i], y = bref_lys_ratio, 
               size = 5, pch = 21, fill = cb_pal[i], color = "black")
  }
  
}

# save plot
ggsave(out_png("I_18h_v_inh_frac_beta_bref"),
       I_18h_v_inh_frac_beta_bref, width = 6, height = 4)

I_18h_v_inh_frac_beta_bref
```

##### Annotate chlorpromazine

```{r}
I_18h_v_inh_frac_beta_chp <- I_18h_v_inh_frac_beta +
  geom_hline(yintercept = chp_lys_ratio, lty = 2) +
  annotate("text", x = 0.3, y = chp_lys_ratio + .15, label = "Chlorpromazine", size = 6)

for (i in 1:num_bases) {
  
  if(chp_inh_frac_df$I_inh_frac[i] < 0.99){
    
    I_18h_v_inh_frac_beta_chp <- I_18h_v_inh_frac_beta_chp +
    annotate("segment", x = chp_inh_frac_df$I_inh_frac[i], xend = chp_inh_frac_df$I_inh_frac[i], 
           y = -Inf, yend = chp_lys_ratio, , color = cb_pal[i]) +  
    annotate("point", x = chp_inh_frac_df$I_inh_frac[i], y = chp_lys_ratio, 
             size = 5, pch = 21, fill = cb_pal[i], color = "black")
  }
}

# save plot
ggsave(out_png("I_18h_v_inh_frac_beta_chp"),
       I_18h_v_inh_frac_beta_chp, width = 6, height = 4)

I_18h_v_inh_frac_beta_chp
```

##### Annotate TATNSF700

```{r}
I_18h_v_inh_frac_beta_tat <- I_18h_v_inh_frac_beta +
  geom_hline(yintercept = tat_lys_ratio, lty = 3) +
  annotate("text", x = 0.15, y = tat_lys_ratio + .15, label = "TATNSF700", size = 6) 

for (i in 1:num_bases) {
  
  if(tat_inh_frac_df$I_inh_frac[i] < 0.99){
    
    I_18h_v_inh_frac_beta_tat <- I_18h_v_inh_frac_beta_tat +
    annotate("segment", x = tat_inh_frac_df$I_inh_frac[i], xend = tat_inh_frac_df$I_inh_frac[i], 
           y = -Inf, yend = tat_lys_ratio, , color = cb_pal[i]) +  
    annotate("point", x = tat_inh_frac_df$I_inh_frac[i], y = tat_lys_ratio, 
             size = 5, pch = 21, fill = cb_pal[i], color = "black")
  }
}

# save plot
ggsave(out_png("I_18h_v_inh_frac_beta_tat"),
       I_18h_v_inh_frac_beta_tat, width = 6, height = 4)

I_18h_v_inh_frac_beta_tat
```

##### Annotate all beta inhibitors - expt ratios only

```{r}
I_18h_v_inh_frac_beta_annot_3 <- I_18h_v_inh_frac_beta +
  geom_hline(yintercept = chp_lys_ratio, lty = 2) +
  geom_hline(yintercept = tat_lys_ratio, lty = 3) +
  geom_hline(yintercept = bref_lys_ratio, lty = 4) +
  annotate("text", x = 0.4, y = chp_lys_ratio + .15, label = "Chlorpromazine", size = 6) +
  annotate("text", x = 0.27, y = bref_lys_ratio + .15, label = "Brefeldin", size = 6) +
  annotate("text", x = 0.15, y = tat_lys_ratio + .15, label = "TATNSF700", size = 6) +
  scale_y_continuous(limits = c(-0.2, 3.5), expand = c(0,0)) 


# save plot
ggsave(out_png("I_18h_v_inh_frac_beta_annot_3"),
       I_18h_v_inh_frac_beta_annot_3, width = 6, height = 4)

I_18h_v_inh_frac_beta_annot_3
```

#### X, I combined

##### All beta inhibitors


```{r}
X_I_18h_beta_inh_annot <- (X_18h_v_inh_frac_beta_annot_3 + theme(legend.position = "none")) /
  (I_18h_v_inh_frac_beta_annot_3 + theme(legend.position = "bottom"))

# save plot
ggsave(out_png("X_I_18h_beta_inh_annot"),
       X_I_18h_beta_inh_annot, width = 6, height = 8)

X_I_18h_beta_inh_annot
```

##### Brefeldin

```{r}
bref_18h_beta_inh_annot <- (X_18h_v_inh_frac_beta_bref + 
                              theme(legend.position = "none") + 
                              ggtitle("Brefeldin")) / 
  (I_18h_v_inh_frac_beta_bref + theme(legend.position = "bottom"))

# save plot
ggsave(out_png("bref_18h_beta_inh_annot"),
       bref_18h_beta_inh_annot, width = 6, height = 8)

bref_18h_beta_inh_annot
```

##### Chlorpromazine

```{r}
chp_18h_beta_inh_annot <- (X_18h_v_inh_frac_beta_chp + 
                             ggtitle("Chlorpromazine") +
                             theme(legend.position = "none")) /
  (I_18h_v_inh_frac_beta_chp + theme(legend.position = "bottom"))

# save plot
ggsave(out_png("chp_18h_beta_inh_annot"),
       chp_18h_beta_inh_annot, width = 6, height = 8)

chp_18h_beta_inh_annot
```


##### TATNSF700

```{r}
tat_18h_beta_inh_annot <- (X_18h_v_inh_frac_beta_tat + 
                             ggtitle("TATNSF700") +
                             theme(legend.position = "none")) /
  (I_18h_v_inh_frac_beta_tat + theme(legend.position = "bottom"))

# save plot
ggsave(out_png("tat_18h_beta_inh_annot"),
       tat_18h_beta_inh_annot, width = 6, height = 8)

tat_18h_beta_inh_annot
```

### Gamma only - annotate inhibitors

#### Extracellular

```{r}
X_18h_v_inh_frac_gamma <- Y_18h %>%
  filter(param_altered == "gamma") %>%
  ggplot(aes(inh_frac, X_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\gamma$)')) +
  ylab(TeX(r'($X_{18h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_18h_v_inh_frac_gamma_nonotes"),
       X_18h_v_inh_frac_gamma, width = 6, height = 4)

X_18h_v_inh_frac_gamma
```

##### Annotate chloroquine

```{r}
# extract data frame of RAB27a inhibition fraction estimates for each base parameter set
chq_inh_frac_df <- inh_frac_df %>%
  filter(inhibitor == "Chloroquine")

X_18h_v_inh_frac_gamma_chq <- X_18h_v_inh_frac_gamma +
  geom_hline(yintercept = chq_cm_ratio, lty = 4) +
  annotate("text", x = 0.25, y = chq_cm_ratio + .05, label = "Chloroquine", size = 6) 

for (i in 1:num_bases) {
  
  if(chq_inh_frac_df$X_inh_frac[i] < .99 & chq_inh_frac_df$X_inh_frac[i] > .01){
    X_18h_v_inh_frac_gamma_chq <- X_18h_v_inh_frac_gamma_chq +
      annotate("segment", x = chq_inh_frac_df$X_inh_frac[i], xend = chq_inh_frac_df$X_inh_frac[i], 
               y = -Inf, yend = chq_cm_ratio, , color = cb_pal[i]) +  
      annotate("point", x = chq_inh_frac_df$X_inh_frac[i], y = chq_cm_ratio, 
               size = 5, pch = 21, fill = cb_pal[i], color = "black")
  }
  
}

# save plot
ggsave(out_png("X_18h_v_inh_frac_gamma_chq"),
       X_18h_v_inh_frac_gamma_chq, width = 6, height = 4)

X_18h_v_inh_frac_gamma_chq
```


#### Intracellular

```{r}
I_18h_v_inh_frac_gamma <- Y_18h %>%
  filter(param_altered == "gamma") %>%
  ggplot(aes(inh_frac, I_norm, col = base_beta)) +
  geom_line(linewidth = 2) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition of $\gamma$)')) +
  ylab(TeX(r'($I_{18h}$ (Normalized))'))  +
  scale_color_manual(values = cb_pal) +
  scale_y_continuous(limits = c(0,3.5), expand = c(0.01,0.01)) +
  scale_x_continuous(expand = c(0.01,0.01)) +
  labs(color = TeX(r'(Base $\beta$)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_18h_v_inh_frac_gamma_nonotes"),
       I_18h_v_inh_frac_gamma, width = 6, height = 4)

I_18h_v_inh_frac_gamma
```

##### Annotate chloroquine

```{r}
I_18h_v_inh_frac_gamma_chq <- I_18h_v_inh_frac_gamma +
  geom_hline(yintercept = chq_lys_ratio, lty = 4) +
  annotate("text", x = 0.25, y = chq_lys_ratio + .15, label = "Chloroquine", size = 6)

for (i in 1:num_bases) {
  
  if(chq_inh_frac_df$I_inh_frac[i] < .99 & chq_inh_frac_df$I_inh_frac[i] > .01){
    I_18h_v_inh_frac_gamma_chq <- I_18h_v_inh_frac_gamma_chq +
      annotate("segment", x = chq_inh_frac_df$I_inh_frac[i], xend = chq_inh_frac_df$I_inh_frac[i], 
               y = -Inf, yend = chq_lys_ratio, , color = cb_pal[i]) +  
      annotate("point", x = chq_inh_frac_df$I_inh_frac[i], y = chq_lys_ratio, 
               size = 5, pch = 21, fill = cb_pal[i], color = "black")
  }
  
}

# save plot
ggsave(out_png("I_18h_v_inh_frac_gamma_chq"),
       I_18h_v_inh_frac_gamma_chq, width = 6, height = 4)

I_18h_v_inh_frac_gamma_chq
```

#### X, I combined

```{r}
chq_18h_gamma_inh_annot <- (X_18h_v_inh_frac_gamma_chq + 
                              ggtitle("Chloroquine") +
                              theme(legend.position = "none")) /
  (I_18h_v_inh_frac_gamma_chq + theme(legend.position = "bottom"))

# save plot
ggsave(out_png("chq_18h_gamma_inh_annot"),
       chq_18h_gamma_inh_annot, width = 6, height = 8)

chq_18h_gamma_inh_annot
```

### Combined chem inh figures

#### Fold change vs inh frac, 18h and 72h

```{r}
p1 <- inhfrac_abgd_xi_grid_18h +
  ggtitle("18h") +
  theme(legend.position = "none")

p2 <- inhfrac_abgd_xi_grid_72h +
  ggtitle("72h") +
  theme(legend.text = element_text(size = 18),
        legend.title = element_text(size = 20),
        legend.key.width = unit(2, "cm")) +
  guides(color = guide_legend(override.aes = list(linewidth = 2)))

combo_fig_18h_72h <- p1 / p2 

ggsave(out_png("combo_fig_18h_72h"),
       combo_fig_18h_72h, width = 12, height = 12)

combo_fig_18h_72h
```

#### Comparison to experimental data


```{r}
combo_fig_chem_inh_vary_base <- (X_18h_v_inh_frac_beta_bref + 
   ggtitle("Brefeldin") +
   theme(legend.position = "none")) + 
  (X_18h_v_inh_frac_beta_chp + 
     ggtitle("Chlorpromazine") +
     theme(legend.position = "none")) +
  (X_18h_v_inh_frac_beta_tat + 
     ggtitle("TATNSF700") +
     theme(legend.position = "none")) +
  (X_18h_v_inh_frac_gamma_chq + ggtitle("Chloroquine") +
     theme(legend.position = "none")) +
  (I_18h_v_inh_frac_beta_bref + theme(legend.position = "none")) +
  (I_18h_v_inh_frac_beta_chp + theme(legend.position = "none")) +
  (I_18h_v_inh_frac_beta_tat + theme(legend.position = "none")) +
  (I_18h_v_inh_frac_gamma_chq + theme(legend.position = "none")) +
  plot_layout(ncol = 4, axis_titles = "collect")

ggsave(out_png("combo_fig_chem_inh_vary_base"),
       combo_fig_chem_inh_vary_base, width = 20, height = 10)

combo_fig_chem_inh_vary_base
```


