---
title: "Untitled"
author: "Amy Gill"
date: "2023-06-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/04_sensitivity/explore_c1/"
dir.create(outfile_prefix, recursive = TRUE)
```

## Import data

### Import .mat files

```{r}
param_file_name <- "../../../saved-data/04_sensitivity/explore_c1/input.mat"
alt_file_name <- "../../../saved-data/04_sensitivity/explore_c1/explore_c1.mat"

# read files and extract data
param_file <- readMat(param_file_name)
alt_file <- readMat(alt_file_name)
```

```{r}
names(param_file)
```


```{r}
names(alt_file)
```

### Species names and parameter names

```{r}
# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")
param_names <- unlist(alt_file$param.names)
```

### Extract time points

```{r}
time_vec <- alt_file$sim.time[1,]
```

### Extract time courses

```{r}
# format data frames of simulated species values for time courses
make_sim_Y_df <- function(Y_mat){
  
  # flatten matrix to data frame of time course data
  # (this isn't very "R-style", but it works, so tell Hadley to chill)
  for (i in 1:dim(Y_mat)[3]) {
    if (i == 1) {    # first run: initialize output data frame
      # extract matrix for first parameter set to a data frame
      Y_temp <- as.data.frame(Y_mat[,,1])
      names(Y_temp) <- species_names
      Y_temp$run_y <- 1
        
      # initialize full Y data frame to the time course from this first run
      Y_df <- Y_temp
      
    } else {    # later runs: append to data frame
      Y_temp <- as.data.frame(Y_mat[,,i])
      names(Y_temp) <- species_names
      Y_temp$run_y <- i
      Y_df <- rbind(Y_df, Y_temp)
    }
  }
  
  Y_df
  
}

# extract simulated time courses
Y_df <- make_sim_Y_df(alt_file$Y.all)

Y_df <- Y_df %>%
  mutate(t = rep(time_vec, dim(alt_file$Y.all)[3])) %>%
  select(t, run = run_y, everything())
```

```{r}
Y_df
```
### Extract parameter values

```{r}
param_val_df <- as.data.frame(t(alt_file$p.mod.all))

names(param_val_df) <- param_names

param_val_df$run <- 1:nrow(param_val_df)

param_val_df <- param_val_df %>%
  mutate(ab = alpha*beta,
         bpg = beta + gamma)

param_val_df
```

### Extract outputs to list

```{r}
alt_file_outlist <- alt_file$out[,,1]
names(alt_file_outlist) <- str_replace_all(names(alt_file_outlist), "\\.", "_")
names(alt_file_outlist)
```


## Plot

### X versus t colored by beta

```{r}
c1_fixed_X_v_t <- param_val_df %>%
  select(run, beta) %>%
  mutate(beta = factor(round(beta, 3))) %>%
  right_join(Y_df) %>%
  ggplot(aes(t, X, color = beta)) +
  geom_line(size = 2) +
  scale_color_manual(name = TeX(r'($\beta$ $(h^{-1})$)'), values = viridis(7)) +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  theme_paper +
  xlab("Time (h)") +
  ylab("X (ng/mL)") +
#  ggtitle(TeX(r'($c_1 = \alpha \beta = 7270$ #/cell/h)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("c1_fixed_X_v_t"),
       c1_fixed_X_v_t, width = 6, height = 4)

c1_fixed_X_v_t
```


### X versus t colored by c2

```{r}
c1_fixed_X_v_t_by_c2 <- param_val_df %>%
  select(run, bpg) %>%
  mutate(bpg = factor(round(bpg, 3))) %>%
  right_join(Y_df) %>%
  ggplot(aes(t, X, color = bpg)) +
  geom_line(size = 2) +
  scale_color_manual(name = TeX(r'($c_2 = \beta + \gamma$ $(h^{-1})$)'), values = viridis(7)) +
  scale_x_continuous(breaks = seq(0, 72, by = 24), expand = c(0,0)) +
  theme_paper +
  xlab("Time (h)") +
  ylab("X (ng/mL)") +
 # ggtitle(TeX(r'($c_1 = \alpha \beta = 7270$ #/cell/h)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("c1_fixed_X_v_t_by_c2"),
       c1_fixed_X_v_t_by_c2, width = 6, height = 4)

c1_fixed_X_v_t_by_c2
```

### X at 72h versus alpha

For a given c1, X increases hyperbolically as alpha increases

```{r}
c1_fixed_x72_v_alpha <- param_val_df %>%
  select(run, alpha) %>%
  mutate(X_72 = alt_file_outlist$X_72) %>%
  ggplot(aes(alpha, X_72)) +
  geom_point(size = 5) +
  #geom_line() +
  xlab(TeX(r'($\alpha$ (#/cell/h))')) +
  ylab(TeX(r'($\X_{72h}$ (ng/mL))')) +
  theme_paper +
  scale_x_continuous(breaks = c(0, 6e5, 1.2e6), limits = c(0,1.3e6), labels = scientific)

# save plot
ggsave(out_png("c1_fixed_x72_v_alpha"),
       c1_fixed_x72_v_alpha, width = 6, height = 4)

c1_fixed_x72_v_alpha
```

### X at 72h versus beta

For a given c1, X is inversely proportional to beta

```{r}
c1_fixed_x72_v_beta <- param_val_df %>%
  select(run, beta) %>%
  mutate(X_72 = alt_file_outlist$X_72) %>%
  ggplot(aes(beta, X_72)) +
  geom_point(size = 5) +
  #geom_line() +
  xlab(TeX(r'($\beta$ $(h^{-1})$)')) +
  ylab(TeX(r'($\X_{72h}$ (ng/mL))')) +
  theme_paper

# save plot
ggsave(out_png("c1_fixed_x72_v_beta"),
       c1_fixed_x72_v_beta, width = 6, height = 4)

c1_fixed_x72_v_beta
```


### X at 72h versus c2

For a given c1, X is inversely proportional to c2 and directly proportional to 1/c2

```{r}
c1_fixed_x72_v_c2 <- param_val_df %>%
  select(run, bpg, ab, delta) %>%
  mutate(X_72 = alt_file_outlist$X_72) %>%
  ggplot(aes(bpg, X_72)) +
  geom_point(size = 5) +
  #geom_line() +
  xlab(TeX(r'($c_2 = \beta + \gamma$ $(h^{-1})$)')) +
  ylab(TeX(r'($\X_{72h}$ (ng/mL))')) +
  theme_paper

# save plot
ggsave(out_png("c1_fixed_x72_v_c2"),
       c1_fixed_x72_v_c2, width = 6, height = 4)

c1_fixed_x72_v_c2
```

### X at 72h versus 1/c2

```{r}
c1_fixed_x72_v_invc2 <- param_val_df %>%
  select(run, bpg) %>%
  mutate(X_72 = alt_file_outlist$X_72,
         inv_c2 = 1/bpg) %>%
  ggplot(aes(inv_c2, X_72)) +
  geom_point(size = 5) +
  #geom_line() +
  xlab(TeX(r'($\frac{1}{c_2} = \frac{1}{\beta + \gamma}$ $(h)$)')) +
  ylab(TeX(r'($\X_{72h}$ (ng/mL))')) +
  theme_paper

# save plot
ggsave(out_png("c1_fixed_x72_v_invc2"),
       c1_fixed_x72_v_invc2, width = 6, height = 4)

c1_fixed_x72_v_invc2
```

### X at 72h v nontrivial params

```{r}
c1_fixed_x72_v_params <- c1_fixed_x72_v_alpha + c1_fixed_x72_v_beta + c1_fixed_x72_v_c2 + 
  plot_layout(nrow = 1)

ggsave(out_png("c1_fixed_x72_v_params"),
       c1_fixed_x72_v_params, width = 12, height = 4)

c1_fixed_x72_v_params
```


### I versus t colored by beta

```{r}
c1_fixed_I_v_t <- param_val_df %>%
  select(run, beta) %>%
  mutate(beta = factor(round(beta, 3))) %>%
  right_join(Y_df) %>%
  ggplot(aes(t, I, color = beta)) +
  geom_line(size = 2) +
  scale_color_manual(name = TeX(r'($\beta$ $(h^{-1})$)'), values = viridis(7)) +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_y_continuous(labels = scientific) +
  theme_paper +
  #ylim(c(0, 6e5)) +
  xlab("Time (h)") +
  ylab("I (#/cell)") +
  #ggtitle(TeX(r'($c_1 = \alpha \beta = 7270$ #/cell/h)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("c1_fixed_I_v_t"),
       c1_fixed_I_v_t, width = 6, height = 4)

c1_fixed_I_v_t
```

### I versus t colored by c2

```{r}
c1_fixed_I_v_t_by_c2 <- param_val_df %>%
  select(run, bpg) %>%
  mutate(bpg = factor(round(bpg, 3))) %>%
  right_join(Y_df) %>%
  ggplot(aes(t, I, color = bpg)) +
  geom_line(size = 2) +
  scale_color_manual(name = TeX(r'($c_2 = \beta + \gamma$ $(h^{-1})$)'), values = viridis(7)) +
  scale_x_continuous(breaks = seq(0, 72, by = 24), expand = c(0,0)) +
  theme_paper +
  scale_y_continuous(labels = scientific) +
  xlab("Time (h)") +
  ylab("I (#/cell)") +
  #ggtitle(TeX(r'($c_1 = \alpha \beta = 7270$ #/cell/h)')) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("c1_fixed_I_v_t_by_c2"),
       c1_fixed_I_v_t_by_c2, width = 6, height = 4)

c1_fixed_I_v_t_by_c2
```

### I_ss versus alpha

```{r}
c1_fixed_I_v_alpha <- param_val_df %>%
  select(run, alpha) %>%
  #mutate(beta_factor = factor(round(beta, 3))) %>%
  right_join(Y_df) %>%
  filter(t == 0) %>%
  ggplot(aes(alpha, I/1e6)) + #, color = beta_factor
  geom_point(size = 5) +
  #geom_line() +
  #scale_color_manual(name = TeX(r'($\beta$ $(h^{-1})$)'), values = cb_pal[-6]) +
  theme_paper +
  #ylim(c(0, 6e5)) +
  xlab(TeX(r'($\alpha$ (#/cell/h))')) +
  ylab(TeX(r'($\I_{SS}$ ($10^6$/cell))')) +
  scale_x_continuous(breaks = c(0, 6e5, 1.2e6), limits = c(0,1.3e6), labels = scientific) #+
  #scale_y_continuous(labels = scientific)

# save plot
ggsave(out_png("c1_fixed_I_v_alpha"),
       c1_fixed_I_v_alpha, width = 6, height = 4)

c1_fixed_I_v_alpha
```

### I_ss versus beta

```{r}
c1_fixed_I_v_beta <- param_val_df %>%
  select(run, beta) %>%
  #mutate(beta_factor = factor(round(beta, 3))) %>%
  right_join(Y_df) %>%
  filter(t == 0) %>%
  ggplot(aes(beta, I/1e6)) + #, color = beta_factor
  geom_point(size = 5) +
  #geom_line() +
  #scale_color_manual(name = TeX(r'($\beta$ $(h^{-1})$)'), values = cb_pal[-6]) +
  theme_paper +
  #ylim(c(0, 6e5)) +
  xlab(TeX(r'($\beta$ $(h^{-1})$)')) +
  ylab(TeX(r'($\I_{SS}$ ($10^6$/cell))')) #+
#  scale_y_continuous(labels = scientific)

# save plot
ggsave(out_png("c1_fixed_I_v_beta"),
       c1_fixed_I_v_beta, width = 6, height = 4)

c1_fixed_I_v_beta
```


### I_ss versus 1/beta

Not quite linearly proportional to 1/beta

```{r}
c1_fixed_I_v_invbeta <- param_val_df %>%
  select(run, beta) %>%
  mutate(invbeta = 1/beta) %>%
  #mutate(beta_factor = factor(round(beta, 3))) %>%
  right_join(Y_df) %>%
  filter(t == 0) %>%
  ggplot(aes(invbeta, I/1e6)) + #, color = beta_factor
  geom_point(size = 5) +
  #geom_line() +
  #scale_color_manual(name = TeX(r'($\beta$ $(h^{-1})$)'), values = cb_pal[-6]) +
  theme_paper +
  #ylim(c(0, 6e5)) +
  xlab(TeX(r'($\frac{1}{\beta}$ (h))')) +
  ylab(TeX(r'($\I_{SS}$ ($10^6$/cell))')) #+
#  scale_y_continuous(labels = scientific)

# save plot
ggsave(out_png("c1_fixed_I_v_invbeta"),
       c1_fixed_I_v_invbeta, width = 6, height = 4)

c1_fixed_I_v_invbeta
```

### I_ss versus c2

```{r}
c1_fixed_I_v_c2 <- param_val_df %>%
  select(run, bpg) %>%
  #mutate(beta_factor = factor(round(beta, 3))) %>%
  right_join(Y_df) %>%
  filter(t == 0) %>%
  ggplot(aes(bpg, I/1e6)) + #, color = beta_factor
  geom_point(size = 5) +
  #geom_line() +
  #scale_color_manual(name = TeX(r'($\beta$ $(h^{-1})$)'), values = cb_pal[-6]) +
  theme_paper +
  #ylim(c(0, 6e5)) +
  xlab(TeX(r'($c_2 = \beta + \gamma$ $(h^{-1})$)')) +
  ylab(TeX(r'($\I_{SS}$ ($10^6$/cell))')) #+
#  scale_y_continuous(labels = scientific)

# save plot
ggsave(out_png("c1_fixed_I_v_c2"),
       c1_fixed_I_v_c2, width = 6, height = 4)

c1_fixed_I_v_c2
```


### I_ss versus 1/c2

```{r}
c1_fixed_I_v_invc2 <- param_val_df %>%
  select(run, bpg) %>%
  #mutate(beta_factor = factor(round(beta, 3))) %>%
  mutate(inv_c2 = 1/bpg) %>%
  right_join(Y_df) %>%
  filter(t == 0) %>%
  ggplot(aes(inv_c2, I/1e6)) + #, color = beta_factor
  geom_point(size = 5) +
  #geom_line() +
  #scale_color_manual(name = TeX(r'($\beta$ $(h^{-1})$)'), values = cb_pal[-6]) +
  theme_paper +
  #ylim(c(0, 6e5)) +
  xlab(TeX(r'($\frac{1}{c_2} = \frac{1}{\beta + \gamma}$ (h))')) +
  ylab(TeX(r'($\I_{SS}$ ($10^6$/cell))')) #+
#  scale_y_continuous(labels = scientific)

# save plot
ggsave(out_png("c1_fixed_I_v_invc2"),
       c1_fixed_I_v_invc2, width = 6, height = 4)

c1_fixed_I_v_invc2
```



### I_ss v nontrivial params

```{r}
c1_fixed_I_v_params <- c1_fixed_I_v_alpha + c1_fixed_I_v_beta + c1_fixed_I_v_c2 + 
  plot_layout(nrow = 1)

ggsave(out_png("c1_fixed_I_v_params"),
       c1_fixed_I_v_params, width = 12, height = 4)

c1_fixed_I_v_params
```

### X,I v t by c2 combined

```{r}
c1_fixed_IX_v_t_by_c2 <- (c1_fixed_X_v_t_by_c2 + theme(legend.position = "none")) / 
  (c1_fixed_I_v_t_by_c2 + theme(legend.position = "bottom"))

# save plot
ggsave(out_png("c1_fixed_IX_v_t_by_c2"),
       c1_fixed_IX_v_t_by_c2, width = 6.5, height = 8)

c1_fixed_IX_v_t_by_c2
```

### X_72 and I_ss v nontrivial params

```{r}
c1_fixed_IX_v_params <- (c1_fixed_x72_v_alpha + c1_fixed_x72_v_beta + c1_fixed_x72_v_c2) /
  (c1_fixed_I_v_alpha + c1_fixed_I_v_beta + c1_fixed_I_v_c2)

ggsave(out_png("c1_fixed_IX_v_params"),
       c1_fixed_IX_v_params, width = 12, height = 7)

c1_fixed_IX_v_params
```

