---
title: "Sensitivity analysis of constitutive sFLT1 secretion - fraction inhibition by chemical manipulation"
author: "Amy Gill"
date: "2023-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This file imports data from sensitivity analysis of extracellular sFLT1 (X) and intracellular sFLT1 (I) at 72h performed 9/26/23 using the Matlab script "sFlt-model/04_sensitivity/sens_analysis.m" with run_mode = "sens_vary_inh_genchem", genetic_chemical = "chemical".

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/04_sensitivity/frac_inh_chem/"
dir.create(outfile_prefix, recursive = TRUE)
```

## Load and wrangle files

### Simulation input file

```{r}
# import file of input parameters
param_file <- readMat("../../../saved-data/04_sensitivity/frac_inh_chem/input.mat")

# inspect available variables
names(param_file)
```

### Simulation output file

```{r}
# import file 
inh_file <- readMat("../../../saved-data/04_sensitivity/frac_inh_chem/sens_vary_inh_genchem_chemical.mat")

# inspect available variables
names(inh_file)
```

### Species names and parameter names

```{r}
# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")
param_names <- unlist(inh_file$param.names)
```

### Extract outputs to list

```{r}
inh_file_outlist <- inh_file$out[,,1]
names(inh_file_outlist) <- str_replace_all(names(inh_file_outlist), "\\.", "_")
names(inh_file_outlist)
```
```{r}
names(inh_file)
```

### Extract time points

```{r}
time_vec <- inh_file$sim.time[1,]
```

### Extract baseline time courses for X, I

```{r}
base_df <- data.frame(t = time_vec,
                      X = inh_file$base.sim.y[which(species_names == "X"),],
                      I = inh_file$base.sim.y[which(species_names == "I"),],
                      param_altered = "none")

base_df
```

### Extract time courses

```{r}
# format data frames of simulated species values for time courses
make_sim_Y_df <- function(Y_mat){
  
  # flatten matrix to data frame of time course data
  # (this isn't very "R-style", but it works, so tell Hadley to chill)
  for (i in 1:dim(Y_mat)[3]) {
    if (i == 1) {    # first run: initialize output data frame
      # extract matrix for first parameter set to a data frame
      Y_temp <- as.data.frame(Y_mat[,,1])
      names(Y_temp) <- species_names
      Y_temp$run_y <- 1
        
      # initialize full Y data frame to the time course from this first run
      Y_df <- Y_temp
      
    } else {    # later runs: append to data frame
      Y_temp <- as.data.frame(Y_mat[,,i])
      names(Y_temp) <- species_names
      Y_temp$run_y <- i
      Y_df <- rbind(Y_df, Y_temp)
    }
  }
  
  Y_df
  
}

# extract simulated data for Hornig
Y_df <- make_sim_Y_df(inh_file$Y.all)
```

```{r}
Y_df <- Y_df %>%
  mutate(t = rep(time_vec, 100*length(param_names))) %>%
  select(t, run = run_y, everything())

Y_df 
```

### Extract parameter alteration metadata

```{r}
# extract parameter altered
param_altered <- unlist(lapply(1:length(param_names), function(x){
  rep(param_names[x],100)
}))

# extract inhibition fraction
param_meta_df <- data.frame(run = as.numeric(1:500),
                           param_altered = factor(param_altered,
                                                  levels = c("alpha", "beta", "gamma",
                                                             "delta", "tau")), 
                           inh_frac = rep(inh_file$inh.frac,5))

param_meta_df
```

### Merge time course data and metadata

```{r}
Y_df <- Y_df %>%
  left_join(param_meta_df, by = "run")
```


### Extract 72h values from time courses

```{r}
# all species at 18h in all conditions
Y_72h <- Y_df %>%
  filter(t == 72)

# baseline X at 18h
X0_72h <- base_df %>%
  filter(t == 72) %>%
  pull(X)

# baseline I at 18h
I0_72h <- base_df %>%
  filter(t == 72) %>%
  pull(I)

# calculate X,I normalized to baseline 18h
Y_72h <- Y_72h %>%
  mutate(X_norm = X / X0_72h,
         I_norm = I / I0_72h)

Y_72h
```

### Extract 18h values from time courses

```{r}
# all species at 18h in all conditions
Y_18h <- Y_df %>%
  filter(t == 18)

# baseline X at 18h
X0_18h <- base_df %>%
  filter(t == 18) %>%
  pull(X)

# baseline I at 18h
I0_18h <- base_df %>%
  filter(t == 18) %>%
  pull(I)

# calculate X,I normalized to baseline 18h
Y_18h <- Y_18h %>%
  mutate(X_norm = X / X0_18h,
         I_norm = I / I0_18h)

Y_18h
```

## Combined 18/72h values

```{r}
Y_combo <- rbind(Y_18h, Y_72h)
Y_combo
```

### Kinghorn experimental data from CSV

```{r}
exp_df <- read_csv("../../../saved-data/karina/updated_inhibitor_data_2023-03-29/inh_data_wrangled.csv")
```

### Calculate mean treatment/control ratios (fraction of control)

```{r}
# calculate mean treatment/control ratio for each variable (X,I) for each treatment type
mean_trt_ctrl_ratios <- exp_df %>%
  group_by(treatment_type, measured_var) %>%
  summarize(mean_ratio = mean(trt_ctrl_ratio))

# function to extract a particular mean ratio
pull_mean_ratio <- function(treatment, var){
  mean_trt_ctrl_ratios %>%
    filter(treatment_type == treatment & measured_var == var) %>%
    pull(mean_ratio)
}

# extract ratios to be used later
bref_cm_ratio <- pull_mean_ratio("Brefeldin", "sFlt1_CM")
bref_lys_ratio <- pull_mean_ratio("Brefeldin", "sFlt1_lysate")
tat_cm_ratio <- pull_mean_ratio("TATNSF", "sFlt1_CM")
tat_lys_ratio <- pull_mean_ratio("TATNSF", "sFlt1_lysate")
chp_cm_ratio <- pull_mean_ratio("Chlorpromazine", "sFlt1_CM")
chp_lys_ratio <- pull_mean_ratio("Chlorpromazine", "sFlt1_lysate")
chq_cm_ratio <- pull_mean_ratio("Chloroquine", "sFlt1_CM")
chq_lys_ratio <- pull_mean_ratio("Chloroquine", "sFlt1_lysate")
```

#### Beta altered

```{r}
Y_18h_beta <- Y_18h %>%
  filter(param_altered == "beta")

# find beta inhibition value that maps to brefeldin effect on extracellular sFLT1
bref_beta_X_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$X_norm - bref_cm_ratio))]

# find beta inhibition value that maps to brefeldin effect on intracellular sFLT1
bref_beta_I_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$I_norm - bref_lys_ratio))]

# find beta inhibition value that maps to chlorpromazine effect on extracellular sFLT1
chp_beta_X_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$X_norm - chp_cm_ratio))]

# find beta inhibition value that maps to chlorpromazine effect on intracellular sFLT1
chp_beta_I_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$I_norm - chp_lys_ratio))]

tat_beta_X_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$X_norm - tat_cm_ratio))]

tat_beta_I_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$I_norm - tat_lys_ratio))]
```

#### Gamma altered

```{r}
Y_18h_gamma <- Y_18h %>%
  filter(param_altered == "gamma")

# find gamma inhibition value that maps to chloroquine effect on extracellular sFLT1
chq_gamma_X_inh_frac <- Y_18h_gamma$inh_frac[which.min(
  abs(Y_18h_gamma$X_norm - chq_cm_ratio))]

# find gamma inhibition value that maps to chloroquine effect on intracellular sFLT1
chq_gamma_I_inh_frac <- Y_18h_gamma$inh_frac[which.min(
  abs(Y_18h_gamma$I_norm - chq_lys_ratio))]
```

### Hornig experimental data

```{r}
# load experimental data from Hornig
load("../../../saved-data/hornig.rda")

hornig_basal <- hornig %>%
  filter(condition == "basal") %>%
  select(t = nominal_time, X = Sx_ng_ml)

hornig_basal
```

## Base plots

### Base X

```{r}
base_df %>%
  ggplot(aes(t, X)) +
  geom_line() +
  xlab("Time (h)") +
  ylab("X (ng/mL)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  theme_paper
```

### Base I

```{r}
base_df %>%
  ggplot(aes(t, I)) +
  geom_line() +
  xlab("Time (h)") +
  ylab("I (#/cell)") +
    scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_y_continuous(limits = c(0, 1e6), labels = scientific) +
  theme_paper
```

## 72h plots: fraction of initial amount 

### abgd, facet grid, X and I


```{r}
inhfrac_abgd_xi_grid_72h <- Y_72h %>%
  filter(param_altered != "tau") %>%
  mutate(param_color = case_when(param_altered == "alpha" ~ param_pal_list$alpha,
                                 param_altered == "beta" ~ param_pal_list$beta,
                                 param_altered == "gamma" ~ param_pal_list$gamma,
                                 param_altered == "delta" ~ param_pal_list$delta)) %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "X", "I"),
         species = factor(species, levels = c("X", "I"))) %>%
  #mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  ggplot(aes(inh_frac, amount, color = param_altered)) +
    geom_hline(yintercept = 1, lty = 2, linewidth = 1) +
  geom_line(linewidth = 2) +
  facet_grid(species ~ param_altered, scales = "free", labeller = label_parsed) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme_paper +
  scale_x_continuous(expand = c(0,0)) +
  scale_color_manual(values = unlist(param_pal_list)) +
  theme(legend.position = "none",
        panel.spacing = unit(1.5, "lines"))
  
# save plot
ggsave(out_png("inhfrac_abgd_xi_grid_72h"),
       inhfrac_abgd_xi_grid_72h, width = 12, height = 6)

inhfrac_abgd_xi_grid_72h
```

### abgd, shared axes, X and I

```{r}
inhfrac_abgd_xi_shared_72h <- Y_72h %>%
  filter(!param_altered == "tau") %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "X", "I"),
         species = factor(species, levels = c("X", "I"))) %>%
  #mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  ggplot(aes(inh_frac, amount, col = param_altered)) +
  geom_hline(yintercept = 1, lty = 2, linewidth = 1) +
  geom_line(linewidth = 2) +
  facet_grid(species ~ .) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme_paper +
  scale_color_manual(values = unlist(param_pal_list), label = color_labels) +
    scale_x_continuous(expand = c(0,0)) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Parameter inhibited"))

# save plot
ggsave(out_png("inhfrac_abgd_xi_shared_72h"),
       inhfrac_abgd_xi_shared_72h, width = 4, height = 6)

inhfrac_abgd_xi_shared_72h
```


## 18h plots: fraction of initial amount 

### abgd, facet grid, X and I

```{r}
inhfrac_abgd_xi_grid_18h <- Y_18h %>%
  filter(param_altered != "tau") %>%
  mutate(param_color = case_when(param_altered == "alpha" ~ param_pal_list$alpha,
                                 param_altered == "beta" ~ param_pal_list$beta,
                                 param_altered == "gamma" ~ param_pal_list$gamma,
                                 param_altered == "delta" ~ param_pal_list$delta)) %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  ggplot(aes(inh_frac, amount, color = param_altered)) +
    geom_hline(yintercept = 1, lty = 2, linewidth = 1) +
  geom_line(linewidth = 2) +
  facet_grid(species ~ param_altered, scales = "free", labeller = label_parsed) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme_paper +
  scale_color_manual(values = unlist(param_pal_list)) +
  theme(legend.position = "none",
        panel.spacing = unit(1.5, "lines"))
  
# save plot
ggsave(out_png("inhfrac_abgd_xi_grid_18h"),
       inhfrac_abgd_xi_grid_18h, width = 12, height = 6)

inhfrac_abgd_xi_grid_18h
```


### abgd, shared axes, X and I

```{r}
inhfrac_abgd_xi_shared_18h <- Y_18h %>%
  filter(param_altered != "tau") %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "X", "I"),
         species = factor(species, levels = c("X", "I"))) %>%
  ggplot(aes(inh_frac, amount, col = param_altered)) +
    geom_hline(yintercept = 1, lty = 2, linewidth = 1) +
  geom_line(linewidth = 2) +
  facet_grid(species ~ .) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme_paper +
    scale_x_continuous(expand = c(0,0)) +
  scale_color_manual(values = unlist(param_pal_list), label = color_labels) +
  theme(legend.position = "bottom",
        panel.spacing = unit(1.5, "lines")) +
  guides(color = guide_legend(title = "Parameter inhibited"))

# save plot
ggsave(out_png("inhfrac_abgd_xi_shared_18h"),
       inhfrac_abgd_xi_shared_18h, width = 4, height = 6)

inhfrac_abgd_xi_shared_18h
```

### Beta and gamma

#### Extracellular

```{r}
bg_18h_X_v_inh_annot_4 <- Y_18h %>%
  filter(param_altered %in% c("beta", "gamma")) %>%
  ggplot(aes(inh_frac, X_norm, col = param_altered)) +
  geom_line(linewidth = 2) +
  theme_paper +
  scale_color_manual(values = c(param_pal_list$beta, param_pal_list$gamma)) +
  xlab(TeX(r'(Fraction inhibition)')) +
  ylab(TeX(r'($X_{18h}$ (Normalized))')) +
  annotate("segment", x = -Inf, xend = tat_beta_X_inh_frac, 
           y = tat_cm_ratio, yend = tat_cm_ratio, lty = 2) +
  annotate("segment", x = tat_beta_X_inh_frac, xend = tat_beta_X_inh_frac, 
           y = -Inf, yend = tat_cm_ratio, lty = 2) +
  annotate("segment", x = -Inf, xend = chq_gamma_X_inh_frac, 
           y = chq_cm_ratio, yend = chq_cm_ratio, lty = 4) +
  annotate("segment", x = chq_gamma_X_inh_frac, xend = chq_gamma_X_inh_frac, 
           y = -Inf, yend = chq_cm_ratio, lty = 4) +
  annotate("segment", x = -Inf, xend = bref_beta_X_inh_frac, 
           y = bref_cm_ratio, yend = bref_cm_ratio, lty = 3) +
  annotate("segment", x = bref_beta_X_inh_frac, xend = bref_beta_X_inh_frac, 
           y = -Inf, yend = bref_cm_ratio, lty = 3) +
  annotate("segment", x = -Inf, xend = chp_beta_X_inh_frac, 
           y = chp_cm_ratio, yend = chp_cm_ratio, lty = 1) +
  annotate("segment", x = chp_beta_X_inh_frac, xend = chp_beta_X_inh_frac, 
           y = -Inf, yend = chp_cm_ratio, lty = 1) +
  annotate("text", x = 0.12, y = tat_cm_ratio + .07, label = "TATNSF", size = 6) +
  annotate("text", x = 0.25, y = chq_cm_ratio + .1, label = "chloroquine", size = 6) +
   annotate("text", x = 0.4, y = bref_cm_ratio + .07, label = "brefeldin", size = 6) +
  annotate("text", x = 0.25, y = chp_cm_ratio - .07, label = "chlorpromazine", size = 6) +
  annotate("point", x = tat_beta_X_inh_frac, y = tat_cm_ratio, size = 5) +
  annotate("point", x = chq_gamma_X_inh_frac, y = chq_cm_ratio, size = 5) +
  annotate("point", x = bref_beta_X_inh_frac, y = bref_cm_ratio, size = 5) +
  annotate("point", x = chp_beta_X_inh_frac, y = chp_cm_ratio, size = 5) +
  annotate("text", x = .4, y = .85, label = parse(text = "beta"), 
           color = param_pal_list$beta, size = 12) +
  annotate("text", x = .4, y = 1.4, label = parse(text = "gamma"), 
           color = param_pal_list$gamma, size = 12) +
  scale_x_continuous(expand = c(0,0)) +
  theme(legend.position = "none")

# save plot
ggsave(out_png("bg_18h_X_v_inh_annot_4"),
       bg_18h_X_v_inh_annot_4, width = 6, height = 4)

bg_18h_X_v_inh_annot_4
```

##### Poster version

```{r}
bg_18h_X_v_inh_annot_4_poster <- Y_18h %>%
  filter(param_altered %in% c("beta", "gamma")) %>%
  ggplot(aes(inh_frac*100, X_norm, col = param_altered)) +
  geom_line(linewidth = 2) +
  theme_poster +
  scale_color_manual(values = c(param_pal_list$beta, param_pal_list$gamma)) +
  xlab(TeX(r'(% inhibition)')) +
  ylab(TeX(r'($X_{18h}$)')) +
  annotate("segment", x = -Inf, xend = tat_beta_X_inh_frac*100, 
           y = tat_cm_ratio, yend = tat_cm_ratio, lty = 2) +
  annotate("segment", x = tat_beta_X_inh_frac*100, xend = tat_beta_X_inh_frac*100, 
           y = -Inf, yend = tat_cm_ratio, lty = 2) +
  annotate("segment", x = -Inf, xend = chq_gamma_X_inh_frac*100, 
           y = chq_cm_ratio, yend = chq_cm_ratio, lty = 4) +
  annotate("segment", x = chq_gamma_X_inh_frac*100, xend = chq_gamma_X_inh_frac*100, 
           y = -Inf, yend = chq_cm_ratio, lty = 4) +
  annotate("segment", x = -Inf, xend = bref_beta_X_inh_frac*100, 
           y = bref_cm_ratio, yend = bref_cm_ratio, lty = 3) +
  annotate("segment", x = bref_beta_X_inh_frac*100, xend = bref_beta_X_inh_frac*100, 
           y = -Inf, yend = bref_cm_ratio, lty = 3) +
  annotate("segment", x = -Inf, xend = chp_beta_X_inh_frac*100, 
           y = chp_cm_ratio, yend = chp_cm_ratio, lty = 1) +
  annotate("segment", x = chp_beta_X_inh_frac*100, xend = chp_beta_X_inh_frac*100, 
           y = -Inf, yend = chp_cm_ratio, lty = 1) +
  annotate("text", x = 12, y = tat_cm_ratio + .07, label = "TATNSF", size = 6) +
  annotate("text", x = 25, y = chq_cm_ratio + .1, label = "chloroquine", size = 6) +
   annotate("text", x = 40, y = bref_cm_ratio + .07, label = "brefeldin", size = 6) +
  annotate("text", x = 25, y = chp_cm_ratio - .07, label = "chlorpromazine", size = 6) +
  annotate("point", x = tat_beta_X_inh_frac*100, y = tat_cm_ratio, size = 5) +
  annotate("point", x = chq_gamma_X_inh_frac*100, y = chq_cm_ratio, size = 5) +
  annotate("point", x = bref_beta_X_inh_frac*100, y = bref_cm_ratio, size = 5) +
  annotate("point", x = chp_beta_X_inh_frac*100, y = chp_cm_ratio, size = 5) +
  annotate("text", x = 40, y = .85, label = parse(text = "beta"), 
           color = param_pal_list$beta, size = 12) +
  annotate("text", x = 40, y = 1.4, label = parse(text = "gamma"), 
           color = param_pal_list$gamma, size = 12) +
  scale_x_continuous(expand = c(0,0)) +
  theme(legend.position = "none",
        plot.margin = margin(r = 10))

# save plot
ggsave(out_png("bg_18h_X_v_inh_annot_4_poster"),
       bg_18h_X_v_inh_annot_4_poster, width = 6, height = 4)

bg_18h_X_v_inh_annot_4_poster
```


#### Intracellular

```{r}
bg_18h_I_v_inh_annot_4 <- Y_18h %>%
  filter(param_altered %in% c("beta", "gamma")) %>%
  ggplot(aes(inh_frac, I_norm, col = param_altered)) +
  geom_line(linewidth = 2) +
    scale_color_manual(values = c(param_pal_list$beta, param_pal_list$gamma)) +
  theme_paper +
  xlab(TeX(r'(Fraction inhibition)')) +
  ylab(TeX(r'($I_{18h}$ (Normalized))')) +
  geom_hline(yintercept = bref_lys_ratio, lty = 3) + # annotate X level with brefeldin
  geom_hline(yintercept = chp_lys_ratio) +   # annotate X level with chlorpromazine
  geom_hline(yintercept = tat_lys_ratio, lty = 2) +
    annotate("segment", x = -Inf, xend = chq_gamma_I_inh_frac, 
           y = chq_lys_ratio, yend = chq_lys_ratio, lty = 4) +
  annotate("segment", x = chq_gamma_I_inh_frac, xend = chq_gamma_I_inh_frac, 
           y = -Inf, yend = chq_lys_ratio, lty = 4) +
  annotate("text", x = .12, y = tat_lys_ratio + .07, label = "TATNSF", size = 6) +
  annotate("text", x = .20, y = chq_lys_ratio + .07, label = "chloroquine", size = 6) +
   annotate("text", x = .40, y = bref_lys_ratio - .07, label = "brefeldin", size = 6) +
  annotate("text", x = .25, y = chp_lys_ratio - .07, label = "chlorpromazine", size = 6) +
  annotate("point", x = chq_gamma_I_inh_frac, y = chq_lys_ratio, size = 5) +
  annotate("text", x = .80, y = 1.15, label = parse(text = "beta"), 
           color = param_pal_list$beta, size = 12) +
  annotate("text", x = .80, y = 2.15, label = parse(text = "gamma"), 
           color = param_pal_list$gamma, size = 12) +
# or annotate()
    scale_x_continuous(expand = c(0,0)) +
  theme(legend.position = "none")

# save plot
ggsave(out_png("bg_18h_I_v_inh_annot_4"),
       bg_18h_I_v_inh_annot_4, width = 6, height = 4)

bg_18h_I_v_inh_annot_4
```

Interesting that the effect of secretion-inhibiting drugs on I at 18h can't be reached with this parameter set - suggests a significant missing component adjusting effects. Perhaps suggests 

This means the difference in parameters between control and drug treatment has to be more complex than just a change in secretion rate


##### Poster version

```{r}
bg_18h_I_v_inh_annot_4_poster <-Y_18h %>%
  filter(param_altered %in% c("beta", "gamma")) %>%
  ggplot(aes(inh_frac*100, I_norm, col = param_altered)) +
  geom_line(linewidth = 2) +
    scale_color_manual(values = c(param_pal_list$beta, param_pal_list$gamma)) +
  theme_poster +
  xlab(TeX(r'(% inhibition)')) +
  ylab(TeX(r'($I_{18h}$)')) +
  geom_hline(yintercept = bref_lys_ratio, lty = 3) + # annotate X level with brefeldin
  geom_hline(yintercept = chp_lys_ratio) +   # annotate X level with chlorpromazine
  geom_hline(yintercept = tat_lys_ratio, lty = 2) +
    annotate("segment", x = -Inf, xend = chq_gamma_I_inh_frac*100, 
           y = chq_lys_ratio, yend = chq_lys_ratio, lty = 4) +
  annotate("segment", x = chq_gamma_I_inh_frac*100, xend = chq_gamma_I_inh_frac*100, 
           y = -Inf, yend = chq_lys_ratio, lty = 4) +
  annotate("text", x = 12, y = tat_lys_ratio + .07, label = "TATNSF", size = 6) +
  annotate("text", x = 20, y = chq_lys_ratio + .07, label = "chloroquine", size = 6) +
   annotate("text", x = 40, y = bref_lys_ratio - .07, label = "brefeldin", size = 6) +
  annotate("text", x = 25, y = chp_lys_ratio - .07, label = "chlorpromazine", size = 6) +
  annotate("point", x = chq_gamma_I_inh_frac*100, y = chq_lys_ratio, size = 5) +
  annotate("text", x = 80, y = 1.15, label = parse(text = "beta"), 
           color = param_pal_list$beta, size = 12) +
  annotate("text", x = 80, y = 2.15, label = parse(text = "gamma"), 
           color = param_pal_list$gamma, size = 12) +
# or annotate()
    scale_x_continuous(expand = c(0,0)) +
  theme(legend.position = "none")

# save plot
ggsave(out_png("bg_18h_I_v_inh_annot_4_poster"),
       bg_18h_I_v_inh_annot_4_poster, width = 6, height = 4)

bg_18h_I_v_inh_annot_4_poster
```

#### X,I

```{r}
bg_18h_combo_v_inh_annot <- bg_18h_X_v_inh_annot_4 / bg_18h_I_v_inh_annot_4

# save plot
ggsave(out_png("bg_18h_combo_v_inh_annot"),
       bg_18h_combo_v_inh_annot, width = 6, height = 8)

bg_18h_combo_v_inh_annot
```


## Combined 18/72h plots

```{r}
inhfrac_abgd_xi_grid_combo <- Y_combo %>%
  mutate(t = factor(t)) %>%
  filter(param_altered != "tau") %>%
  mutate(param_color = case_when(t == "72" ~ param_pal[1],
                                 param_altered == "alpha" ~ param_pal_list$alpha,
                                 param_altered == "beta" ~ param_pal_list$beta,
                                 param_altered == "gamma" ~ param_pal_list$gamma,
                                 param_altered == "delta" ~ param_pal_list$delta)) %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "X", "I"),
         species = factor(species, levels = c("X", "I"))) %>%
  #mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  ggplot(aes(inh_frac, amount, color = param_altered, lty = t, alpha = t)) +
        geom_hline(yintercept = 1, linewidth = 0.5) +
  geom_line(linewidth = 2) +
  facet_grid(species ~ param_altered, scales = "free", labeller = label_parsed) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme_paper +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 1, by = 0.25)) +
  scale_color_manual(values = unlist(param_pal_list)) +
  scale_alpha_discrete(range=c(0.5, 1)) +
  theme(legend.position = "none",
        panel.spacing = unit(1.5, "lines"))
  
# save plot
ggsave(out_png("inhfrac_abgd_xi_grid_combo"),
       inhfrac_abgd_xi_grid_combo, width = 12, height = 6)

inhfrac_abgd_xi_grid_combo
```


## Time courses for 90% inhibition

```{r}
base_df
```


```{r}
Y_df %>%
  filter(inh_frac == .99) %>%
  select(t, X, I, param_altered)
```


```{r}
Y_df_90pct <- Y_df %>%
  filter(inh_frac == 0.9) %>%
  select(t, X, I, param_altered) %>%
  rbind(base_df) %>%
  mutate(linetype = case_when(#param_altered == "beta" ~ "3",
                              param_altered == "tau" ~ "2",
                              TRUE ~ "1"),
         linesize = ifelse(param_altered == "none", 4, 2))

#levels(Y_df_90pct$param_altered) <- c("none", "alpha", "beta", "gamma", "delta", "kappa", "tau")
```

### Extracellular

```{r}
X_tc_inh_90 <- Y_df %>%
  filter(inh_frac == 0.9) %>%
  ggplot(aes(t, X, color = param_altered)) +
  geom_line(data = base_df, color = "black", linewidth = 5) +
  geom_line(linewidth = 2) +
  geom_point(data = hornig_basal, color = "black", size = 8) +
  scale_color_manual(values = unlist(param_pal_list), label = color_labels) +
  scale_linetype_manual(values = c("solid", "dashed")) + 
  scale_x_continuous(breaks = seq(0, 72, by = 24), expand = c(0,0)) +
  theme_paper +
  #guides(color = guide_legend(title = "Parameter inhibited (90%)",
  #                            nrow = 2, byrow = TRUE), lty = "none") +
  xlab("Time (h)") +
  ylab("X (ng/mL)") +
  #theme(legend.position = c(0.3, 0.8)) +
  theme(legend.position = "none") +
  annotate("text", x = 30, y = 0.6, label = parse(text = "alpha"), color = param_pal_list$alpha, size = 12) +
  annotate("text", x = 70, y = 0.7, label = parse(text = "beta"), color = param_pal_list$beta, size = 12) +
  annotate("text", x = 25, y = 3.5, label = parse(text = "gamma"), color = param_pal_list$gamma, size = 12) +
  annotate("text", x = 55, y = 5.5, label = parse(text = "delta"), color = param_pal_list$delta, size = 12) +
  annotate("text", x = 60, y = 2.2, label = parse(text = "tau"), color = param_pal_list$tau, size = 12)

# save plot
ggsave(out_png("X_tc_inh_90"),
       X_tc_inh_90, width = 6, height = 4)

X_tc_inh_90
```

### Intracellular

```{r}
I_tc_inh_90 <- Y_df %>%
  filter(inh_frac == 0.9) %>%
  mutate(linetype = ifelse(param_altered == "tau", "2", "1")) %>%
  ggplot(aes(t, I, color = param_altered)) +
  geom_line(data = base_df, aes(t, I), color = "black", linewidth = 5) +
  geom_line(aes(lty = linetype), linewidth = 2) +
  scale_color_manual(values = unlist(param_pal_list), label = color_labels) +
  scale_linetype_manual(values = c("solid", "twodash", "dotdash", "dotted")) + 
  theme_paper +
  guides(color = guide_legend(title = "Parameter inhibited (90%)",
                              nrow = 2, byrow = TRUE), lty = "none") +
  xlab("Time (h)") +
  ylab("I (#/cell)") +
  #theme(legend.position = c(0.7, 0.65))
    theme(legend.position = "none") +
  annotate("text", x = 50, y = 2e5, label = parse(text = "alpha"), color = param_pal_list$alpha, size = 12) +
  annotate("text", x = 50, y = 1.3e6, label = parse(text = "beta"), color = param_pal_list$beta, size = 12) +
  annotate("text", x = 50, y = 2e6, label = parse(text = "gamma"), color = param_pal_list$gamma, size = 12) +
  annotate("text", x = 62, y = 6.2e5, label = parse(text = "delta"), color = param_pal_list$delta, size = 12) +
  annotate("text", x = 67, y = 6e5, label = parse(text = "tau"), color = param_pal_list$tau, size = 12) +
    scale_y_continuous(labels = scientific) +
  scale_x_continuous(breaks = seq(0, 72, by = 24), expand = c(0,0))

# save plot
ggsave(out_png("I_tc_inh_90"),
       I_tc_inh_90, width = 6, height = 4)

I_tc_inh_90
```

### Combined X/I

```{r}
combo_tc_inh_90 <- X_tc_inh_90 / I_tc_inh_90

# save plot
ggsave(out_png("combo_tc_inh_90"),
       combo_tc_inh_90, width = 6, height = 8)

combo_tc_inh_90
```




## Sensitivity table

```{r}
sens_df_pt9 <- Y_combo %>% 
  filter(inh_frac == 0.9) %>%
  select(param_altered, t, X, I) %>%
  #mutate(t = paste0(t, "h")) %>%
  pivot_longer(cols = X:I, names_to = "out_var", values_to = "abs_val") %>%
  mutate(out_var = paste0(out_var, "_", t, "h")) %>%
  select(-t) %>%
  mutate(ref_val = case_when(
    out_var == "X_18h" ~ X0_18h,
    out_var == "I_18h" ~ I0_18h,
    out_var == "X_72h" ~ X0_72h,
    out_var == "I_72h" ~ I0_72h
  )) %>%
  mutate(norm_val = abs_val / ref_val,
         abs_sens = (abs_val - ref_val) / ref_val,
         abs_sens_pct = abs_sens * 100, 
         rel_sens = abs_sens / -.9) %>%
  separate_wider_delim(out_var, "_", names = c("out_var", "out_time"))

sens_df_pt9
```

### Chemical inhibition at 18h

```{r}
sens_df_pt9_18h_df <- sens_df_pt9 %>%
  filter(out_time == "18h") %>%
  mutate(out_var = paste(out_var, out_time, sep = "_")) %>%
  select(param_altered, out_var, abs_sens) %>%
  pivot_wider(names_from = "out_var", values_from = "abs_sens")

sens_df_pt9_18h_mat <- sens_df_pt9_18h_df %>%
  select(-param_altered) %>%
  as.matrix()

sens_df_pt9_18h_mat
```

```{r}
sens_cheminh_18h_heatmap <- Heatmap(sens_df_pt9_18h_mat,
                              heatmap_height = unit(3, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = "Fractional change at 18h",
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 20, fontfamily = "serif"),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in")
                              ),
                              cluster_rows = FALSE,
                              cluster_columns = FALSE,
                              column_names_rot = 45,
                              row_names_side = "left",
                              row_labels = expression(alpha, beta, gamma, delta, tau),
                              row_names_gp = gpar(fontsize = 20, fontfamily = "serif"),
                              column_names_gp = gpar(fontsize = 16, fontfamily = "serif"),
                              column_labels = c(TeX(r'($X_{18h}$)'), 
                                                TeX(r'($I_{18h}$)')),
                              col = pg_color_func,
                              rect_gp = gpar(col = "black", lwd = 2),
                              cell_fun = function(i, j, x, y, width, height, fill) {
                                  grid.text(sprintf("%.2f", sens_df_pt9_18h_mat[j, i]), 
                                          x, y, gp = gpar(fontsize = 16,
                                                          #fontface = "bold",
                                                          col = ifelse(
                                                abs(sens_df_pt9_18h_mat[j,i]) < 0.6,
                                                            "black", "white")))
                                })

#local_sens_heatmap

draw(sens_cheminh_18h_heatmap, heatmap_legend_side = "top")
```

### Chemical inhibition at 72h

```{r}
sens_df_pt9_72h_df <- sens_df_pt9 %>%
  filter(out_time == "72h") %>%
  mutate(out_var = paste(out_var, out_time, sep = "_")) %>%
  select(param_altered, out_var, abs_sens) %>%
  pivot_wider(names_from = "out_var", values_from = "abs_sens")

sens_df_pt9_72h_mat <- sens_df_pt9_72h_df %>%
  select(-param_altered) %>%
  as.matrix()

sens_df_pt9_72h_mat
```

```{r}
sens_cheminh_72h_heatmap <- Heatmap(sens_df_pt9_72h_mat,
                              heatmap_height = unit(3, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = "Fractional change at 72h",
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 20, fontfamily = "serif"),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in")
                              ),
                              cluster_rows = FALSE,
                              cluster_columns = FALSE,
                              column_names_rot = 45,
                              row_names_side = "left",
                              row_labels = expression(alpha, beta, gamma, delta, tau),
                              row_names_gp = gpar(fontsize = 20, fontfamily = "serif"),
                              column_names_gp = gpar(fontsize = 16, fontfamily = "serif"),
                              column_labels = c(TeX(r'($X_{72h}$)'), 
                                                TeX(r'($I_{72h}$)')),
                              col = pg_color_func,
                              rect_gp = gpar(col = "black", lwd = 2),
                              cell_fun = function(i, j, x, y, width, height, fill) {
                                  grid.text(sprintf("%.2f", sens_df_pt9_72h_mat[j, i]), 
                                          x, y, gp = gpar(fontsize = 16,
                                                          #fontface = "bold",
                                                          col = ifelse(
                                                abs(sens_df_pt9_72h_mat[j,i]) < 0.6,
                                                            "black", "white")))
                                })

#local_sens_heatmap

draw(sens_cheminh_72h_heatmap, heatmap_legend_side = "top")
```

### Chemical inhibition at both times

```{r}
sens_df_pt9_mat <- sens_df_pt9 %>%
  mutate(out_var = paste(out_var, out_time, sep = "_")) %>%
  select(param_altered, out_var, abs_sens) %>%
  pivot_wider(names_from = "out_var", values_from = "abs_sens") %>%
  select(-param_altered) %>%
  as.matrix()

sens_df_pt9_mat <- sens_df_pt9_mat[, c("X_18h", "X_72h", "I_18h", "I_72h")]

sens_df_pt9_mat
```

```{r}
sens_cheminh_heatmap <- Heatmap(sens_df_pt9_mat,
                              heatmap_height = unit(3, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = "Fractional change",
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 20, fontfamily = "serif"),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in"),
                                at = c(-2.5, -1.25, 0, 1.25, 2.5)
                              ),
                              cluster_rows = FALSE,
                              cluster_columns = FALSE,
                              column_names_rot = 45,
                              row_names_side = "left",
                              row_labels = expression(alpha, beta, gamma, delta, tau),
                              row_names_gp = gpar(fontsize = 20, fontfamily = "serif"),
                              column_names_gp = gpar(fontsize = 16, fontfamily = "serif"),
                              column_labels = c(TeX(r'($X_{18h}$)'), TeX(r'($X_{72h}$)'), 
                                                TeX(r'($I_{18h}$)'), TeX(r'($I_{72h}$)')),
                              col = pg_color_func_broad,
                              rect_gp = gpar(col = "black", lwd = 2),
                              cell_fun = function(i, j, x, y, width, height, fill) {
                                  grid.text(sprintf("%.2f", sens_df_pt9_mat[j, i]), 
                                          x, y, gp = gpar(fontsize = 16,
                                                          #fontface = "bold",
                                                          col = ifelse(
                                                abs(sens_df_pt9_mat[j,i]) < 1.5,
                                                            "black", "white")))
                                })

#local_sens_heatmap

draw(sens_cheminh_heatmap, heatmap_legend_side = "top")
```

#### save heatmap

```{r}
png(file = paste0(outfile_prefix, "sens_cheminh_heatmap.png"), 
    width=5, height=4, units="in", res=100)
draw(sens_cheminh_heatmap, heatmap_legend_side = "top")
dev.off()
```

### Revised color scale

```{r}
sens_df_pt9_logfc_mat <- sens_df_pt9 %>%
  mutate(out_var = paste(out_var, out_time, sep = "_"),
         lognorm = log2(norm_val)) %>%
  select(param_altered, out_var, lognorm) %>%
  pivot_wider(names_from = "out_var", values_from = "lognorm") %>%
  select(-param_altered) %>%
  as.matrix()

sens_df_pt9_logfc_mat <- sens_df_pt9_logfc_mat[, c("X_18h", "X_72h", "I_18h", "I_72h")]

sens_df_pt9_logfc_mat
```

```{r}
# define purple -> green color palette
# endpoints taken from RColorBrewer "PRGn"
pg_color_func_broad <- colorRamp2(c(-3.5, 0, 3.5), c("#762A83", "white", "#1B7837"))

sens_cheminh_heatmap_log <- Heatmap(sens_df_pt9_logfc_mat,
                              heatmap_height = unit(3, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = TeX(r'($\log_2$ fold change)'),
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 20, fontfamily = "serif"),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in"),
                                at = c(-3, -1.5, 0, 1.5, 3)
                              ),
                              cluster_rows = FALSE,
                              cluster_columns = FALSE,
                              column_names_rot = 45,
                              row_names_side = "left",
                              row_labels = expression(alpha, beta, gamma, delta, tau),
                              row_names_gp = gpar(fontsize = 20, fontfamily = "serif"),
                              column_names_gp = gpar(fontsize = 16, fontfamily = "serif"),
                              column_labels = c(TeX(r'($X_{18h}$)'), TeX(r'($X_{72h}$)'), 
                                                TeX(r'($I_{18h}$)'), TeX(r'($I_{72h}$)')),
                              col = pg_color_func_broad,
                              rect_gp = gpar(col = "black", lwd = 2),
                              cell_fun = function(i, j, x, y, width, height, fill) {
                                  grid.text(sprintf("%.2f", sens_df_pt9_mat[j, i]), 
                                          x, y, gp = gpar(fontsize = 16,
                                                          #fontface = "bold",
                                                          col = ifelse(
                                                abs(sens_df_pt9_logfc_mat[j,i]) < 2,
                                                            "black", "white")))
                                })

#local_sens_heatmap

draw(sens_cheminh_heatmap_log, heatmap_legend_side = "top")

```


### Revised color and values

```{r}
sens_cheminh_heatmap_log_v2 <- Heatmap(sens_df_pt9_logfc_mat,
                              heatmap_height = unit(3, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = TeX(r'($\log_2$ fold change)'),
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 20, fontfamily = "serif"),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in"),
                                at = c(-3, -1.5, 0, 1.5, 3)
                              ),
                              cluster_rows = FALSE,
                              cluster_columns = FALSE,
                              column_names_rot = 45,
                              row_names_side = "left",
                              row_labels = expression(alpha, beta, gamma, delta, tau),
                              row_names_gp = gpar(fontsize = 20, fontfamily = "serif"),
                              column_names_gp = gpar(fontsize = 16, fontfamily = "serif"),
                              column_labels = c(TeX(r'($X_{18h}$)'), TeX(r'($X_{72h}$)'), 
                                                TeX(r'($I_{18h}$)'), TeX(r'($I_{72h}$)')),
                              col = pg_color_func_broad,
                              rect_gp = gpar(col = "black", lwd = 2),
                              cell_fun = function(i, j, x, y, width, height, fill) {
                                  grid.text(sprintf("%.2f", sens_df_pt9_logfc_mat[j, i]), 
                                          x, y, gp = gpar(fontsize = 16,
                                                          #fontface = "bold",
                                                          col = ifelse(
                                                abs(sens_df_pt9_logfc_mat[j,i]) < 2,
                                                            "black", "white")))
                                })

#local_sens_heatmap

draw(sens_cheminh_heatmap_log_v2, heatmap_legend_side = "top")

```

#### save heatmap

```{r}
png(file = paste0(outfile_prefix, "sens_cheminh_heatmap_log_v2"), 
    width=5, height=4, units="in", res=100)
draw(sens_cheminh_heatmap_log_v2, heatmap_legend_side = "top")
dev.off()
```
