---
title: "Local sensitivity analysis of constitutive sFLT1 secretion"
author: "Amy Gill"
date: "2023-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file imports data from sensitivity analysis of extracellular sFLT1 (X) and intracellular sFLT1 (I) at 72h performed 9/20/23 using the Matlab script "sFlt-model/04_sensitivity/sens_analysis.m" with run_mode = "local_sens_const".

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/04_sensitivity/local_sens/"
dir.create(outfile_prefix, recursive = TRUE)
```



## Load and wrangle simulation output files

### Load files

```{r}
# import file 
sens_file <- readMat("../../../saved-data/04_sensitivity/local_sens/local_sens_const.mat")

# inspect available variables
names(sens_file)
```

### Parameter names

```{r}
param_names <- unlist(sens_file$param.names)
param_names
```

### Extract outputs to list

```{r}
sens_file_outlist <- sens_file$out[,,1]
names(sens_file_outlist) <- str_replace_all(names(sens_file_outlist), "\\.", "_")
names(sens_file_outlist)
```

### Convert sensitivies to matrix

```{r}
# identify columns to show in heatmap (sensitivities except AUMCs)
ind <- which(str_detect(names(sens_file_outlist), "sens")
             & !str_detect(names(sens_file_outlist), "AUMC"))

# convert list of sensitivities to matrix
sens_mat <- as.matrix(as.data.frame(sens_file_outlist[ind]))

# transpose matrix so rows are output variables and columns are parameters varied
sens_mat <- t(sens_mat)

# add or adjust row/col annotations
colnames(sens_mat) <- param_names
rownames(sens_mat) <- str_remove(rownames(sens_mat), "sens_")

# inspect
sens_mat
```

## Heatmap v1 (x = output variable, y = parameter adjusted)

### Generate heatmap

```{r}
local_sens_heatmap_v1 <- Heatmap(t(sens_mat),
                              heatmap_height = unit(3, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = "Relative Sensitivity",
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 18),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in")
                              ),
                              cluster_rows = FALSE,
                              column_names_rot = 45,
                              row_names_side = "left",
                              row_labels = expression(alpha, beta, gamma, delta, tau),
                              row_names_gp = gpar(fontsize = 18),
                              column_names_gp = gpar(fontsize = 14),
                              column_labels = c(TeX(r'($X_{72h}$)'), 
                                                TeX(r'($I_{72h}$)'), 
                                             TeX(r'(${AUC}_{72h}(X)$)'), 
                                             TeX(r'(${AUC}_{72h}(I)$)'), 
                                             TeX(r'($I_{SS}$)'),
                                             TeX(r'($X_{SS}$)'),
                                             TeX(r'($T_{50}$ to $X_{SS}$)')),
                              col = pg_color_func,
                              rect_gp = gpar(col = "black", lwd = 2))

#local_sens_heatmap

draw(local_sens_heatmap_v1, heatmap_legend_side = "top")
```

### Save heatmap

Good to know - ggsave doesn't work because ComplexHeatmap doesn't use ggplot, but builds on a common ancestor

```{r}
png(file = paste0(outfile_prefix, "local_sens_heatmap_v1.png"), width=5, height=4, units="in", res=100)
draw(local_sens_heatmap_v1, heatmap_legend_side = "top")
dev.off()
```

## Heatmap v2 - minimum info (72h)


```{r}
sens_mat_small <- sens_mat[c("I_72", "X_72", "T50_X"),]

local_sens_heatmap_v2 <- Heatmap(t(sens_mat_small),
                              heatmap_height = unit(3, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = "Relative Sensitivity",
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 20, fontfamily = "serif"),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in")
                              ),
                              cluster_rows = FALSE,
                              column_names_rot = 45,
                              row_names_side = "left",
                              row_labels = expression(alpha, beta, gamma, delta, tau),
                              row_names_gp = gpar(fontsize = 20, fontfamily = "serif"),
                              column_names_gp = gpar(fontsize = 16, fontfamily = "serif"),
                              column_labels = c(TeX(r'($I_{72h}$)'), 
                                                TeX(r'($X_{72h}$)'), 
                                             TeX(r'($T_{50}(X_{SS})$)')),
                              col = pg_color_func,
                              rect_gp = gpar(col = "black", lwd = 2))

#local_sens_heatmap

draw(local_sens_heatmap_v2, heatmap_legend_side = "top")
```

## Heatmap v3 - annotated cells for v2


```{r}
local_sens_heatmap_v3 <- Heatmap(t(sens_mat_small),
                              heatmap_height = unit(3, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = "Relative Sensitivity",
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 20, fontfamily = "serif"),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in")
                              ),
                              cluster_rows = FALSE,
                              cluster_columns = FALSE,
                              column_names_rot = 45,
                              row_names_side = "left",
                              row_labels = expression(alpha, beta, gamma, delta, tau),
                              row_names_gp = gpar(fontsize = 20, fontfamily = "serif"),
                              column_names_gp = gpar(fontsize = 16, fontfamily = "serif"),
                              column_labels = c(TeX(r'($I_{72h}$)'), 
                                                TeX(r'($X_{72h}$)'), 
                                             TeX(r'($T_{50}(X_{SS})$)')),
                              col = pg_color_func,
                              rect_gp = gpar(col = "black", lwd = 2),
                              cell_fun = function(i, j, x, y, width, height, fill) {
                                  grid.text(sprintf("%.2f", sens_mat_small[i, j]), 
                                          x, y, gp = gpar(fontsize = 16,
                                                          #fontface = "bold",
                                                          col = ifelse(
                                                            abs(sens_mat_small[i,j]) < 0.6,
                                                            "black", "white")))
                                })

#local_sens_heatmap

draw(local_sens_heatmap_v3, heatmap_legend_side = "top")
```



We performed local univariate sensitivity analysis by starting from the median parameter set, individually increasing each parameter by 10%, and calculating the relative change in several output variables. 

(2 ways of saying the same thing.)

An increase in the production rate constant $\alpha$ causes a linear increase in both intracellular and extracellular sFLT1. (linear? directly proportional? 1:1? )

Increasing the secretion rate constant $\beta$ causes a moderate rise in extracellular sFLT1 but a mild decrease in intracellular sFLT1, while increasing the intracellular degradation rate constant $\gamma$ causes a decline in both intracellular and extracellular levels. The extracellular degradation rate constant $\delta$ has a strong negative effect on extracellular sFLT1 but has no impact on intracellular protein. Notably, $\delta$ is the only parameter which affects the time to half-steady-state ($T_{50}$) for extracellular sFLT1, with higher $\delta$ associated with a faster time to steady state [consistent with known effects of increasing clearance - ref a PK textbook?].




### Save heatmaps

Good to know - ggsave doesn't work because ComplexHeatmap doesn't use ggplot, but builds on a common ancestor

```{r}
png(file = paste0(outfile_prefix, "local_sens_heatmap_v2_serif.png"), width=5, height=4, units="in", res=100)
draw(local_sens_heatmap_v2, heatmap_legend_side = "top")
dev.off()
```


```{r}
png(file = paste0(outfile_prefix, "local_sens_heatmap_v3_serif.png"), width=5, height=4, units="in", res=100)
draw(local_sens_heatmap_v3, heatmap_legend_side = "top")
dev.off()
```

