---
title: "Untitled"
author: "Amy Gill"
date: "2023-06-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/04_sensitivity/explore_c2/"
dir.create(outfile_prefix, recursive = TRUE)
```


## Import data

### Import .mat files

```{r}
param_file_name <- "../../../saved-data/04_sensitivity/explore_c2/input.mat"
alt_file_name <- "../../../saved-data/04_sensitivity/explore_c2/explore_c2.mat"

# read files and extract data
param_file <- readMat(param_file_name)
alt_file <- readMat(alt_file_name)
```

```{r}
names(param_file)
```


```{r}
names(alt_file)
```

### Species names and parameter names

```{r}
# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")
param_names <- unlist(alt_file$param.names)
```

### Extract time points

```{r}
time_vec <- alt_file$sim.time[1,]
```

### Extract time courses

```{r}
# format data frames of simulated species values for time courses
make_sim_Y_df <- function(Y_mat){
  
  # flatten matrix to data frame of time course data
  # (this isn't very "R-style", but it works, so tell Hadley to chill)
  for (i in 1:dim(Y_mat)[3]) {
    if (i == 1) {    # first run: initialize output data frame
      # extract matrix for first parameter set to a data frame
      Y_temp <- as.data.frame(Y_mat[,,1])
      names(Y_temp) <- species_names
      Y_temp$run_y <- 1
        
      # initialize full Y data frame to the time course from this first run
      Y_df <- Y_temp
      
    } else {    # later runs: append to data frame
      Y_temp <- as.data.frame(Y_mat[,,i])
      names(Y_temp) <- species_names
      Y_temp$run_y <- i
      Y_df <- rbind(Y_df, Y_temp)
    }
  }
  
  Y_df
  
}

# extract simulated time courses
Y_df <- make_sim_Y_df(alt_file$Y.all)

Y_df <- Y_df %>%
  mutate(t = rep(time_vec, dim(alt_file$Y.all)[3])) %>%
  select(t, run = run_y, everything())
```

```{r}
Y_df
```
### Extract parameter values

```{r}
param_val_df <- as.data.frame(t(alt_file$p.mod.all))

names(param_val_df) <- param_names

param_val_df$run <- 1:nrow(param_val_df)

param_val_df <- param_val_df %>%
  mutate(ab = alpha*beta,
         bpg = beta + gamma)

param_val_df
```

```{r}
param_val_df %>%
  select(run, beta) %>%
  mutate(beta = factor(round(beta, 3))) %>%
  right_join(Y_df)
```


### Extract outputs to list

```{r}
alt_file_outlist <- alt_file$out[,,1]
names(alt_file_outlist) <- str_replace_all(names(alt_file_outlist), "\\.", "_")
names(alt_file_outlist)
```


## Plot

### X versus t colored by beta

```{r}
c2_fixed_X_v_t <- param_val_df %>%
  select(run, beta) %>%
  mutate(beta = factor(round(beta, 3))) %>%
  right_join(Y_df) %>%
  ggplot(aes(t, X, color = beta)) +
  geom_line(size = 2) +
  scale_x_continuous(breaks = seq(0, 72, by = 24), expand = c(0,0)) +
  scale_color_manual(name = TeX(r'($\beta$ $(h^{-1})$)'), values = viridis(7)) +
  theme_paper +
  xlab("Time (h)") +
  ylab("X (ng/mL)") #+
#  ggtitle(TeX(r'($c_2 = \beta + \gamma = .191$ $h^{-1}$)'))

# save plot
ggsave(out_png("c2_fixed_X_v_t"),
       c2_fixed_X_v_t, width = 6, height = 4)

c2_fixed_X_v_t
```


### X versus t colored by c1

```{r}
c2_fixed_X_v_t_by_c1 <- param_val_df %>%
  select(run, ab) %>%
  mutate(ab = factor(round(ab))) %>%
  right_join(Y_df) %>%
  ggplot(aes(t, X, color = ab)) +
  geom_line(size = 2) +
  scale_color_manual(name = TeX(r'($c_1 = \alpha \beta$ (#/cell/$h^2$))'), values = viridis(7)) +
  scale_x_continuous(breaks = seq(0, 72, by = 24), expand = c(0,0)) +
  theme_paper +
  xlab("Time (h)") +
  ylab("X (ng/mL)") #+
#  ggtitle(TeX(r'($c_2 = \beta + \gamma = .191$ $h^{-1}$)'))

# save plot
ggsave(out_png("c2_fixed_X_v_t_by_c1"),
       c2_fixed_X_v_t_by_c1, width = 6, height = 4)

c2_fixed_X_v_t_by_c1
```

### X at 72h versus beta

For a given c2, X is proportional to beta

```{r}
c2_fixed_x72_v_beta <- param_val_df %>%
  select(run, beta) %>%
  mutate(X_72 = alt_file_outlist$X_72) %>%
  ggplot(aes(beta, X_72)) +
  geom_point(size = 5) +
#  geom_line() +
  xlab(TeX(r'($\beta$ $(h^{-1})$)')) +
  ylab(TeX(r'($\X_{72h}$ (ng/mL))')) +
  theme_paper #+
#  ggtitle(TeX(r'($c_2 = \beta + \gamma = .191$ $h^{-1}$)'))

# save plot
ggsave(out_png("c2_fixed_x72_v_beta"),
       c2_fixed_x72_v_beta, width = 6, height = 4)

c2_fixed_x72_v_beta
```

### X at 72h versus gamma


```{r}
c2_fixed_x72_v_gamma <- param_val_df %>%
  select(run, gamma) %>%
  mutate(X_72 = alt_file_outlist$X_72) %>%
  ggplot(aes(gamma, X_72)) +
  geom_point(size = 5) +
#  geom_line() +
  xlab(TeX(r'($\gamma$ $(h^{-1})$)')) +
  ylab(TeX(r'($\X_{72h}$ (ng/mL))')) +
  theme_paper #+
#  ggtitle(TeX(r'($c_2 = \beta + \gamma = .191$ $h^{-1}$)'))

# save plot
ggsave(out_png("c2_fixed_x72_v_gamma"),
       c2_fixed_x72_v_gamma, width = 6, height = 4)

c2_fixed_x72_v_gamma
```

### X at 72h versus c1



```{r}
c2_fixed_x72_v_c1 <- param_val_df %>%
  select(run, ab) %>%
  mutate(X_72 = alt_file_outlist$X_72) %>%
  ggplot(aes(ab, X_72)) +
  geom_point(size = 5) +
  scale_x_continuous(breaks = seq(0,25000, by = 6250)) +
#  geom_line() +
  xlab(TeX(r'($c_1 = \alpha \beta$ (#/cell/$h^2$))')) +
  ylab(TeX(r'($\X_{72h}$ (ng/mL))')) +
  theme_paper #+
#  ggtitle(TeX(r'($c_2 = \beta + \gamma = .191$ $h^{-1}$)'))

# save plot
ggsave(out_png("c2_fixed_x72_v_c1"),
       c2_fixed_x72_v_c1, width = 6, height = 4)

c2_fixed_x72_v_c1
```

### X at 72h v nontrivial params

```{r}
c2_fixed_x72_v_params <- c2_fixed_x72_v_c1 + c2_fixed_x72_v_beta + c2_fixed_x72_v_gamma + 
  plot_layout(nrow = 1)

ggsave(out_png("c2_fixed_x72_v_params"),
       c2_fixed_x72_v_params, width = 12, height = 4)

c2_fixed_x72_v_params
```



### I versus t by c1

```{r}
c2_fixed_I_v_t_by_c1 <- param_val_df %>%
  select(run, ab) %>%
  mutate(ab = factor(round(ab))) %>%
  right_join(Y_df) %>%
  ggplot(aes(t, I, color = ab)) +
  geom_line(size = 2, alpha = 0.9) +
  scale_color_manual(name = TeX(r'($c_1 = \alpha \beta$ (#/cell/$h^2$))'), values = viridis(7)) +
  scale_x_continuous(breaks = seq(0, 72, by = 24), expand = c(0,0)) +
  scale_y_continuous(labels = scientific, limits = c(0,1e6)) +
  theme_paper +
  xlab("Time (h)") +
  ylab("I (#/cell)") +
  theme(legend.position = "right")

# save plot
ggsave(out_png("c2_fixed_I_v_t_by_c1"),
       c2_fixed_I_v_t_by_c1, width = 6, height = 4)

c2_fixed_I_v_t_by_c1
```
### X,I v t by c1 combined


```{r}
c2_fixed_IX_v_t_by_c1 <- (c2_fixed_X_v_t_by_c1 + theme(legend.position = "none")) / 
  (c2_fixed_I_v_t_by_c1 + theme(legend.position = "bottom"))

# save plot
ggsave(out_png("c2_fixed_IX_v_t_by_c1"),
       c2_fixed_IX_v_t_by_c1, width = 6.5, height = 8)

c2_fixed_IX_v_t_by_c1
```