---
title: "sFLT1 modeling global sensitivity analysis - Orders of magnitude in base parameters"
author: "Amy Gill"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## REVIEWED

This file imports data from sensitivity analysis of extracellular sFLT1 (X) and intracellular sFLT1 (I) at 18h and 72h performed 1/15/24 using the Matlab script "sFlt-model/dde_model/sens_analysis.m" with options:
- run_mode = "sens_logscale_sqrt10"
- genetic_chemical = "genetic"
- param_set = "hjk_opt_median_230821"

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/04_sensitivity/powers_sqrt_10/"
dir.create(outfile_prefix, recursive = TRUE)
```

## Load and wrangle files

### Simulation input file

```{r}
# import file of input parameters
param_file <- readMat("../../../saved-data/04_sensitivity/powers_sqrt_10/input.mat")

# inspect available variables
names(param_file)
```

### Simulation output file

```{r}
# import file 
alt_file <- readMat("../../../saved-data/04_sensitivity/powers_sqrt_10/sens_logscale_sqrt10_genetic.mat")

# inspect available variables
names(alt_file)
```

### Species names and parameter names

```{r}
# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")
param_names <- unlist(alt_file$param.names)
```

### Extract outputs to list

```{r}
alt_file_outlist <- alt_file$out[,,1]
names(alt_file_outlist) <- str_replace_all(names(alt_file_outlist), "\\.", "_")
names(alt_file_outlist)
```
```{r}
names(alt_file)
```



### Extract time points

```{r}
time_vec <- alt_file$sim.time[1,]
```

### Extract baseline time courses for X, I

```{r}
base_df <- data.frame(t = time_vec,
                      X = alt_file$base.sim.y[which(species_names == "X"),],
                      I = alt_file$base.sim.y[which(species_names == "I"),],
                      param_altered = "none")

base_df
```

### Extract time courses

```{r}
# format data frames of simulated species values for time courses
make_sim_Y_df <- function(Y_mat){
  
  # flatten matrix to data frame of time course data
  # (this isn't very "R-style", but it works, so tell Hadley to chill)
  for (i in 1:dim(Y_mat)[3]) {
    if (i == 1) {    # first run: initialize output data frame
      # extract matrix for first parameter set to a data frame
      Y_temp <- as.data.frame(Y_mat[,,1])
      names(Y_temp) <- species_names
      Y_temp$run_y <- 1
        
      # initialize full Y data frame to the time course from this first run
      Y_df <- Y_temp
      
    } else {    # later runs: append to data frame
      Y_temp <- as.data.frame(Y_mat[,,i])
      names(Y_temp) <- species_names
      Y_temp$run_y <- i
      Y_df <- rbind(Y_df, Y_temp)
    }
  }
  
  Y_df
  
}

# extract simulated time courses
Y_df <- make_sim_Y_df(alt_file$Y.all)
```

```{r}
Y_df
```

```{r}
num_params <- length(param_names)
num_factors <- length(alt_file$factors.to.try)
num_runs <- num_params * num_factors

Y_df <- Y_df %>%
  mutate(t = rep(time_vec, num_runs)) %>%
  select(t, run = run_y, everything())

Y_df 
```

### Extract parameter values

```{r}
# initial parameter values
param_file$p.init
```

```{r}
param_val_df <- as.data.frame(t(alt_file$p.mod.all))

names(param_val_df) <- param_names

param_val_df$run <- 1:nrow(param_val_df)

param_val_df
```


### Extract parameter alteration metadata

```{r}
# extract parameter altered
param_altered <- unlist(lapply(1:length(param_names), function(x){
  rep(param_names[x],9)
}))

# extract inhibition fraction
param_meta_df <- data.frame(run = as.numeric(1:num_runs),
                           param_altered = factor(param_altered,
                                                  levels = c("alpha", "beta", "gamma",
                                                             "delta", "tau")), 
                           scaling_numeric = rep(alt_file$factors.to.try, num_params)) %>%
  mutate(scaling_factor = factor(round(scaling_numeric, 3))) %>%
  left_join(param_val_df, by = "run") %>%
  mutate(ab = alpha*beta,
         bpg = beta + gamma,
         I_bar_calc = alpha / bpg,
         X_bar_calc = ab / (delta * bpg))

param_meta_df
```

### Merge time course data and metadata

```{r}
Y_df <- Y_df %>%
  left_join(param_meta_df)
```

```{r}
Y_df
```


### Extract 72h values from time courses

```{r}
# all species at 72h in all conditions
Y_72h <- Y_df %>%
  filter(t == 72)

# baseline X at 18h
X0_72h <- base_df %>%
  filter(t == 72) %>%
  pull(X)

# baseline I at 18h
I0_72h <- base_df %>%
  filter(t == 72) %>%
  pull(I)

# calculate X,I normalized to baseline 18h
Y_72h <- Y_72h %>%
  mutate(X_norm = X / X0_72h,
         I_norm = I / I0_72h)

Y_72h
```

### Extract 18h values from time courses

```{r}
# all species at 18h in all conditions
Y_18h <- Y_df %>%
  filter(t == 18)

# baseline X at 18h
X0_18h <- base_df %>%
  filter(t == 18) %>%
  pull(X)

# baseline I at 18h
I0_18h <- base_df %>%
  filter(t == 18) %>%
  pull(I)

# calculate X,I normalized to baseline 18h
Y_18h <- Y_18h %>%
  mutate(X_norm = X / X0_18h,
         I_norm = I / I0_18h)

Y_18h
```

### Extract 0h values from time courses

```{r}
# all species at 18h in all conditions
Y_0h <- Y_df %>%
  filter(t == 0)

# baseline X at 18h
X0_0h <- base_df %>%
  filter(t == 0) %>%
  pull(X)

# baseline I at 18h
I0_0h <- base_df %>%
  filter(t == 0) %>%
  pull(I)

# calculate X,I normalized to baseline 18h
Y_0h <- Y_0h %>%
  mutate(X_norm = X / X0_0h,
         I_norm = I / I0_0h)

Y_0h
```

### Hornig experimental data

```{r}
# load experimental data from Hornig
load("../../../saved-data/hornig.rda")

hornig_basal <- hornig %>%
  filter(condition == "basal") %>%
  select(t = nominal_time, X = Sx_ng_ml)

hornig_basal
```





## Base plots

### Base X

```{r}
base_df %>%
  ggplot(aes(t, X)) +
  geom_line() +
  xlab("Time (h)") +
  ylab("Extracellular sFLT1 (X, ng/mL)") +
  theme1
```

```{r}
X_norm_val_72 <- base_df %>%
  filter(t == 72) %>%
  pull(X)

X_norm_val_72  # ng/mL
```


### Base I

```{r}
base_df %>%
  ggplot(aes(t, I)) +
  geom_line() +
  xlab("Time (h)") +
  ylab("Intracellular sFLT1 (I, #/cell)") +
  ylim(c(0, 1e6)) +
  theme1
```

```{r}
I_norm_val_72 <- base_df %>%
  filter(t == 72) %>%
  pull(I)

I_norm_val_72  # #/cell
```

### Plot X time course colored by scaling factor

#### Faceted in 2 rows

```{r}
X_facet_scale_sqrt10 <- Y_df %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_wrap(~ param_altered, scales = "free", labeller = label_parsed) +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_facet_scale_sqrt10"),
       X_facet_scale_sqrt10, width = 8, height = 4)

X_facet_scale_sqrt10
```

#### Faceted in 1 row

```{r}
X_facet_scale_sqrt10_1row <- Y_df %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_wrap(~ param_altered, scales = "free", labeller = label_parsed, nrow = 1) +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_facet_scale_sqrt10_1row"),
       X_facet_scale_sqrt10_1row, width = 16, height = 3)

X_facet_scale_sqrt10_1row
```


#### Grid (forces same y scale)

```{r}
X_facet_scale_sqrt10_1row_samescale <- Y_df %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_grid(. ~ param_altered, labeller = label_parsed) +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_facet_scale_sqrt10_1row_samescale"),
       X_facet_scale_sqrt10_1row_samescale, width = 16, height = 3)

X_facet_scale_sqrt10_1row_samescale
```

#### Grid with y log-scaled

```{r}
X_facet_scale_sqrt10_1row_samelog <- X_facet_scale_sqrt10_1row_samescale +
  scale_y_log10(, labels = function(x) sprintf("%g", x))

# save plot
ggsave(out_png("X_facet_scale_sqrt10_1row_samelog"),
       X_facet_scale_sqrt10_1row_samelog, width = 16, height = 3)

X_facet_scale_sqrt10_1row_samelog
```

### X time course: beta

Template for separately plotting any parameters

#### Powers of sqrt(10)

```{r}
X_beta_scale_sqrt10 <- Y_df %>%
  filter(param_altered == "beta") %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 2) +
  theme_paper +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  ggtitle(TeX(r'(Scaling $\beta$)')) +
  guides(color = guide_legend(title = "Scaling\nfactor")) +
  scale_color_manual(values = viridis(9)) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_beta_scale_sqrt10"),
       X_beta_scale_sqrt10, width = 6, height = 4)

X_beta_scale_sqrt10
```

#### Powers of 10

```{r}
X_beta_scale_10 <- Y_df %>%
  filter(param_altered == "beta" & scaling_numeric %in% 10^(-2:2)) %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 2) +
  theme_paper +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  ggtitle(TeX(r'(Scaling $\beta$)')) +
  guides(color = guide_legend(title = "Scaling\nfactor")) +
  scale_color_manual(values = cb_pal) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_beta_scale_10"),
       X_beta_scale_10, width = 6, height = 4)

X_beta_scale_10
```

### Explore X at 72h

#### Raw values: Matrix with parameter altered x scaling factor

```{r}
Y_72h
```


```{r}
X_72h_df <- Y_72h %>%
  select(param_altered, scaling_numeric, X) %>%
  pivot_wider(names_from = param_altered, values_from = X) %>%
  mutate(scaling_factor = factor(round(scaling_numeric, 3)))

X_72h_df

X_72h_mat <- as.matrix(select(X_72h_df, alpha:tau))

rownames(X_72h_mat) <- X_72h_df$scaling_factor

X_72h_mat
```

#### Normalized values: Matrix with parameter altered x scaling factor, div original X_72h


```{r}
X_72h_mat / X_norm_val_72
```


### X (norm) at 72h versus scaling factor

#### Faceted in 2 rows

```{r}
X_72h_norm_df <- X_72h_df %>%
  pivot_longer(alpha:tau, names_to = "param_altered", values_to = "raw_value") %>%
  mutate(norm_value = raw_value / X0_72h,
         param_altered = factor(param_altered, levels = c("alpha", "beta", "gamma", "delta", "tau"))) 

X_72h_norm_v_scale_factor <- X_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_wrap(~ param_altered, scales = "free", labeller = "label_parsed") +
    ylab(TeX(r'($X_{72h}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("X_72h_norm_v_scale_factor"),
       X_72h_norm_v_scale_factor, width = 9, height = 5)

X_72h_norm_v_scale_factor
```


#### Faceted in 1 row

```{r}
X_72h_norm_v_scale_1row <- X_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_wrap(~ param_altered, scales = "free", labeller = "label_parsed", nrow = 1) +
    ylab(TeX(r'($X_{72h}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("X_72h_norm_v_scale_1row"),
       X_72h_norm_v_scale_1row, width = 16, height = 3)

X_72h_norm_v_scale_1row
```


#### Grid (forces same y scale)

```{r}
X_72h_norm_v_scale_1row_samescale <- X_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_grid(. ~ param_altered, labeller = "label_parsed") +
    ylab(TeX(r'($X_{72h}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("X_72h_norm_v_scale_1row_samescale"),
       X_72h_norm_v_scale_1row_samescale, width = 16, height = 3)

X_72h_norm_v_scale_1row_samescale
```

#### Grid with y log-scaled

```{r}
X_72h_norm_v_scale_1row_samelog <- X_72h_norm_v_scale_1row_samescale +
  scale_y_log10(, labels = function(x) sprintf("%g", x))

# save plot
ggsave(out_png("X_72h_norm_v_scale_1row_samelog"),
       X_72h_norm_v_scale_1row_samelog, width = 16, height = 3)

X_72h_norm_v_scale_1row_samelog
```

## NEW 2024-01-16

### Plot I time course (show steady state)

#### Faceted in 2 rows

```{r}
I_facet_scale_sqrt10 <- Y_df %>%
  ggplot(aes(t, I, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_wrap(~ param_altered, scales = "free", labeller = label_parsed) +
  ylab("I (#/cell)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_facet_scale_sqrt10"),
       I_facet_scale_sqrt10, width = 8, height = 4)

I_facet_scale_sqrt10
```

#### Faceted in 1 row

```{r}
I_facet_scale_sqrt10_1row <- Y_df %>%
  ggplot(aes(t, I, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_wrap(~ param_altered, scales = "free", labeller = label_parsed, nrow = 1) +
  ylab("I (#/cell)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_facet_scale_sqrt10_1row"),
       I_facet_scale_sqrt10_1row, width = 16, height = 3)

I_facet_scale_sqrt10_1row
```


#### Grid (forces same y scale)

```{r}
I_facet_scale_sqrt10_1row_samescale <- Y_df %>%
  ggplot(aes(t, I, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_grid(. ~ param_altered, labeller = label_parsed) +
  ylab("I (#/cell)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_facet_scale_sqrt10_1row_samescale"),
       I_facet_scale_sqrt10_1row_samescale, width = 16, height = 3)

I_facet_scale_sqrt10_1row_samescale
```

#### Grid with y log-scaled

```{r}
I_facet_scale_sqrt10_1row_samelog <- I_facet_scale_sqrt10_1row_samescale +
  scale_y_log10(, labels = function(x) sprintf("%g", x))

# save plot
ggsave(out_png("I_facet_scale_sqrt10_1row_samelog"),
       I_facet_scale_sqrt10_1row_samelog, width = 16, height = 3)

I_facet_scale_sqrt10_1row_samelog
```



#### I at 72h (steady state)

```{r}
I_72h_df <- Y_72h %>%
  select(param_altered, scaling_factor, I, scaling_numeric) %>%
  pivot_wider(names_from = param_altered, values_from = I)

I_72h_mat <- as.matrix(select(I_72h_df, alpha:tau))

rownames(I_72h_mat) <- I_72h_df$scaling_factor

I_72h_df
```

#### Normalized values

```{r}
I_72h_norm_df <- I_72h_df %>%
  pivot_longer(alpha:tau, names_to = "param_altered", values_to = "raw_value") %>%
  mutate(norm_value = raw_value / I0_72h,
         param_altered = factor(param_altered, levels = c("alpha", "beta", "gamma", "delta", "tau"))) 

I_72h_norm_df
```


### I (norm) at 72h (ss) versus scaling factor

#### Faceted in 2 rows

```{r}
I_72h_norm_v_scale_factor <- I_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_wrap(~ param_altered, scales = "free", labeller = "label_parsed") +
    ylab(TeX(r'($I_{SS}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("I_72h_norm_v_scale_factor"),
       I_72h_norm_v_scale_factor, width = 9, height = 5)

I_72h_norm_v_scale_factor
```


#### Faceted in 1 row

```{r}
I_72h_norm_v_scale_1row <- I_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_wrap(~ param_altered, scales = "free", labeller = "label_parsed", nrow = 1) +
    ylab(TeX(r'($I_{SS}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("I_72h_norm_v_scale_1row"),
       I_72h_norm_v_scale_1row, width = 16, height = 3)

I_72h_norm_v_scale_1row
```


#### Grid (forces same y scale)

```{r}
I_72h_norm_v_scale_1row_samescale <- I_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_grid(. ~ param_altered, labeller = "label_parsed") +
    ylab(TeX(r'($I_{SS}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("I_72h_norm_v_scale_1row_samescale"),
       I_72h_norm_v_scale_1row_samescale, width = 16, height = 3)

I_72h_norm_v_scale_1row_samescale
```

#### Grid with y log-scaled

```{r}
I_72h_norm_v_scale_1row_samelog <- I_72h_norm_v_scale_1row_samescale +
  scale_y_log10(, labels = function(x) sprintf("%g", x))

# save plot
ggsave(out_png("I_72h_norm_v_scale_1row_samelog"),
       I_72h_norm_v_scale_1row_samelog, width = 16, height = 3)

I_72h_norm_v_scale_1row_samelog
```




## NOT REVIEWED



## -- WORKSPACE 5/18 --




### Heatmap v1 (x = parameter adjusted, y = scaling factor, color = X (ng/mL))


```{r}
Heatmap(X_72h_mat,
        heatmap_height = unit(4, "in"),
        heatmap_width = unit(4, "in"),
        column_names_gp = gpar(fontsize = 24),
        row_names_gp = gpar(fontsize = 14),
        cluster_columns = FALSE,
        cluster_rows = FALSE,
        column_names_rot = 0,
        column_names_side = "top",
        column_labels = expression(alpha, beta, gamma, delta, tau),
        rect_gp = gpar(col = "black", lwd = 2),
        col = viridis(10))
```











### Heatmap v2 (x = parameter adjusted, y = scaling factor, color = X_norm (fold change from baseline parameter values)))

```{r}
X_norm_72h_df <- Y_72h %>%
  select(param_altered, scaling_factor, X_norm) %>%
  pivot_wider(names_from = param_altered, values_from = X_norm)

X_norm_72h_mat <- as.matrix(X_norm_72h_df[,-1])

rownames(X_norm_72h_mat) <- X_norm_72h_df$scaling_factor

X_norm_72h_mat
```



```{r}
Y_72h
```












```{r}
local_sens_heatmap_v1 <- Heatmap(sens_mat,
                              heatmap_height = unit(4, "in"),
                              heatmap_width = unit(4, "in"),
                              heatmap_legend_param = list(
                                title = "Relative Sensitivity",
                                direction = "horizontal",
                                title_position = "topcenter",
                                title_gp = gpar(fontsize = 18),
                                #grid_height = unit(0.5, "in"),
                                legend_width = unit(2.25, "in")
                              ),
                              cluster_columns = FALSE,
                              column_names_rot = 0,
                              #column_names_side = "top",
                              column_labels = expression(alpha, beta, gamma, delta, tau),
                              column_names_gp = gpar(fontsize = 18),
                              row_names_gp = gpar(fontsize = 14),
                              row_labels = c("X at 72h", "I at 72h", 
                                             TeX(r'(${AUC}_{72h}$ for X)'), 
                                             TeX(r'(${AUC}_{72h}$ for I)'), 
                                             TeX(r'($\bar{I}$)'),
                                             TeX(r'($\bar{X}$)'),
                                             TeX(r'($T_{50}$ to $\bar{X}$)')),
                              col = pg_color_func,
                              rect_gp = gpar(col = "black", lwd = 2))

#local_sens_heatmap

draw(local_sens_heatmap_v1, heatmap_legend_side = "top")

```

##


## --STORAGE--

## 72h plots: fraction of initial amount 

### abgd, facet grid, X and I


```{r}
Y_72h
```


```{r}
Y_72h %>%
  filter(param_altered == "alpha") %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount")
```






```{r}
inhfrac_abgd_xi_grid_72h <- Y_72h %>%
  filter(!param_altered %in% c("kappa", "tau")) %>%
  mutate(param_color = case_when(param_altered == "alpha" ~ param_pal_list$alpha,
                                 param_altered == "beta" ~ param_pal_list$beta,
                                 param_altered == "gamma" ~ param_pal_list$gamma,
                                 param_altered == "delta" ~ param_pal_list$delta)) %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  #mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  ggplot(aes(scaling_factor, amount, color = param_altered)) +
  geom_point(size = 2) +
  facet_grid(species ~ param_altered, scales = "free", labeller = label_parsed) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme1 +
  scale_color_manual(values = unlist(param_pal)) +
  theme(legend.position = "none")
  
inhfrac_abgd_xi_grid_72h
```

```{r}
# save plot
ggsave(out_png("inhfrac_abgd_xi_grid_72h"),
       inhfrac_abgd_xi_grid_72h, width = 12, height = 6)
```

### abgd, shared axes, X and I

```{r}
inhfrac_abgd_xi_shared_72h <- Y_72h %>%
  filter(!param_altered %in% c("kappa", "tau")) %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "Extracellular (X)", "Intracellular (I)")) %>%
  #mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  ggplot(aes(inh_frac, amount, col = param_altered)) +
  geom_line(size = 2) +
  facet_wrap(~ species) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme1 +
  scale_color_manual(values = unlist(param_pal), label = color_labels) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Parameter inhibited"))

inhfrac_abgd_xi_shared_72h
```

```{r}
# save plot
ggsave(out_png("inhfrac_abgd_xi_shared_72h"),
       inhfrac_abgd_xi_shared_72h, width = 8, height = 4)
```


## 18h plots: fraction of initial amount 

### All params, faceted, X only

```{r}
Y_18h %>%
  ggplot(aes(inh_frac, X_norm)) +
  geom_point() +
  facet_wrap(~param_altered) +
  theme1 +
  xlab("Fraction inhibition") +
  ylab("Extracellular sFLT1 \n(normalized)")
```

### abgd, facet grid, X and I

```{r}
inhfrac_abgd_xi_grid_18h <- Y_18h %>%
  filter(!param_altered %in% c("kappa", "tau")) %>%
  mutate(param_color = case_when(param_altered == "alpha" ~ param_pal$alpha,
                                 param_altered == "beta" ~ param_pal$beta,
                                 param_altered == "gamma" ~ param_pal$gamma,
                                 param_altered == "delta" ~ param_pal$delta)) %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "Extracellular (X)", "Intracellular (I)")) %>%
  #mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  ggplot(aes(inh_frac, amount, color = param_altered)) +
  geom_line(size = 2) +
  facet_grid(species ~ param_altered, scales = "free", labeller = label_parsed) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme1 +
  scale_color_manual(values = unlist(param_pal)) +
  theme(legend.position = "none")
  
inhfrac_abgd_xi_grid_18h
```

```{r}
# save plot
ggsave(out_png("inhfrac_abgd_xi_grid_18h"),
       inhfrac_abgd_xi_grid_18h, width = 12, height = 6)
```

### abgd, shared axes, X and I

```{r}
inhfrac_abgd_xi_shared_18h <- Y_18h %>%
  filter(!param_altered %in% c("kappa", "tau")) %>%
  pivot_longer(cols = c(X_norm, I_norm),
               names_to = "species", values_to = "amount") %>%
  mutate(species = ifelse(species == "X_norm", "Extracellular (X)", "Intracellular (I)")) %>%
  #mutate(species = ifelse(species == "X_norm", "X", "I")) %>%
  ggplot(aes(inh_frac, amount, col = param_altered)) +
  geom_line(size = 2) +
  facet_wrap(~ species) +
  xlab("Fraction inhibition") +
  ylab("Normalized sFLT1") +
  theme1 +
  scale_color_manual(values = unlist(param_pal), label = color_labels) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Parameter inhibited"))

inhfrac_abgd_xi_shared
```

```{r}
# save plot
ggsave(out_png("inhfrac_abgd_xi_shared_18h"),
       inhfrac_abgd_xi_shared_18h, width = 8, height = 4)
```

### Beta only

#### Extracellular

```{r}
Y_18h %>%
  filter(param_altered == "beta") %>%
  ggplot(aes(inh_frac, X_norm)) +
  geom_line(size = 2, col = param_pal$beta) +
  theme1 +
  xlab(TeX(r'(Fraction inhibition of $\beta$)')) +
  ylab("Extracellular sFLT1 at 18h \n(Normalized to control)") +
  #geom_hline(yintercept = bref_cm_ratio) + # annotate X level with brefeldin
  #geom_hline(yintercept = chp_cm_ratio) +   # annotate X level with chlorpromazine
# or annotate()
  annotate("segment", x = 0, xend = bref_beta_X_inh_frac, 
           y = bref_cm_ratio, yend = bref_cm_ratio, lty = 2) +
  annotate("segment", x = bref_beta_X_inh_frac, xend = bref_beta_X_inh_frac, 
           y = 0, yend = bref_cm_ratio, lty = 2) +
  annotate("segment", x = 0, xend = chp_beta_X_inh_frac, 
           y = chp_cm_ratio, yend = chp_cm_ratio, lty = 1) +
  annotate("segment", x = chp_beta_X_inh_frac, xend = chp_beta_X_inh_frac, 
           y = 0, yend = chp_cm_ratio, lty = 1) +
  annotate("text", x = 0.4, y = bref_cm_ratio + .05, label = "brefeldin", size = 6) +
  annotate("text", x = 0.2, y = chp_cm_ratio + .05, label = "chlorpromazine", size = 6) +
  annotate("point", x = bref_beta_X_inh_frac, y = bref_cm_ratio, size = 5) +
  annotate("point", x = chp_beta_X_inh_frac, y = chp_cm_ratio, size = 5)


```

#### Intracellular

```{r}
Y_18h %>%
  filter(param_altered == "beta") %>%
  ggplot(aes(inh_frac, I_norm)) +
  geom_line(size = 2, col = param_pal$beta) +
  theme1 +
  xlab(TeX(r'(Fraction inhibition of $\beta$)')) +
  ylab("Intracellular sFLT1 at 18h \n(Normalized to control)") +
  geom_hline(yintercept = bref_lys_ratio, lty = 2) + # annotate X level with brefeldin
  geom_hline(yintercept = chp_lys_ratio) +   # annotate X level with chlorpromazine
# or annotate()
  annotate("text", x = 0.4, y = bref_lys_ratio + .05, label = "brefeldin", size = 6) +
  annotate("text", x = 0.2, y = chp_lys_ratio + .05, label = "chlorpromazine", size = 6) 
```

Interesting that the effect of drugs on I at 18h can't be reached with this parameter set

This means the difference in secretion between control and drug treatment has to be 

```{r}

```




```{r}
pull_mean_ratio <- function(treatment, var){
  mean_trt_ctrl_ratios %>%
    filter(treatment_type == treatment & measured_var == var) %>%
    pull(mean_ratio)
}

bref_cm_ratio <- pull_mean_ratio("Brefeldin", "sFlt1_CM")
bref_lys_ratio <- pull_mean_ratio("Brefeldin", "sFlt1_lysate")
chp_cm_ratio <- pull_mean_ratio("Chlorpromazine", "sFlt1_CM")
chp_lys_ratio <- pull_mean_ratio("Chlorpromazine", "sFlt1_lysate")
chq_cm_ratio <- pull_mean_ratio("Chloroquine", "sFlt1_CM")
chq_lys_ratio <- pull_mean_ratio("Chloroquine", "sFlt1_lysate")
```

```{r}
Y_18h_beta <- Y_18h %>%
  filter(param_altered == "beta")

# find beta inhibition value that maps to brefeldin effect on extracellular sFLT1
bref_beta_X_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$X_norm - bref_cm_ratio))]

# find beta inhibition value that maps to brefeldin effect on intracellular sFLT1
bref_beta_I_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$I_norm - bref_lys_ratio))]

# find beta inhibition value that maps to chlorpromazine effect on extracellular sFLT1
chp_beta_X_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$X_norm - chp_cm_ratio))]

# find beta inhibition value that maps to chlorpromazine effect on intracellular sFLT1
chp_beta_I_inh_frac <- Y_18h_beta$inh_frac[which.min(abs(Y_18h_beta$I_norm - chp_lys_ratio))]
```

```{r}
bref_beta_X_inh_frac
bref_beta_I_inh_frac
chp_beta_X_inh_frac
chp_beta_I_inh_frac
```

## Time courses for 90% inhibition

```{r}
names(base_df)
```



```{r}
Y_df_90pct <- Y_df %>%
  filter(inh_frac == 0.9) %>%
  select(t, X, I, param_altered) %>%
  rbind(base_df) %>%
  mutate(linetype = case_when(#param_altered == "beta" ~ "3",
                              param_altered == "tau" ~ "2",
                              TRUE ~ "1"),
         linesize = ifelse(param_altered == "none", 4, 2))

#levels(Y_df_90pct$param_altered) <- c("none", "alpha", "beta", "gamma", "delta", "kappa", "tau")
```

```{r}
Y_df_90pct %>%
  ggplot(aes(t, X, color = param_altered, lty = linetype)) +
  geom_line(size = 2) +
  scale_color_manual(values = unlist(param_pal), label = color_labels) +
  scale_linetype_manual(values = c("solid", "dotted", "twodash")) + 
  theme1 +
  guides(color = guide_legend(title = "Parameter inhibited (90%)",
                              nrow = 2, byrow = TRUE),
         lty = "none") +
  xlab("Time (h)") +
  ylab("Extracellular sFLT1") +
  theme(legend.position = c(0.3, 0.8))
```

### Intracellular

```{r}
I_tc_inh_90 <- Y_df %>%
  filter(inh_frac == 0.9) %>%
  mutate(linetype = case_when(param_altered == "kappa" ~ "2", 
                              param_altered == "tau" ~ "3",
                              TRUE ~ "1")) %>%
  ggplot(aes(t, I, color = param_altered)) +
  geom_line(data = base_df, aes(t, I), color = "black", size = 6) +
  geom_line(aes(lty = linetype), size = 3) +
  scale_color_manual(values = unlist(param_pal), label = color_labels) +
  scale_linetype_manual(values = c("solid", "twodash", "dotdash", "dotted")) + 
  theme1 +
  guides(color = guide_legend(title = "Parameter inhibited (90%)",
                              nrow = 2, byrow = TRUE),
         lty = "none") +
  xlab("Time since inhibitor (h)") +
  ylab("Intracellular sFLT1 (#/cell)") +
  theme(legend.position = c(0.7, 0.65))

I_tc_inh_90
```

```{r}
# save plot
ggsave(out_png("I_tc_inh_90"),
       I_tc_inh_90, width = 6, height = 4)
```

### Extracellular

```{r}
X_tc_inh_90 <- Y_df %>%
  filter(inh_frac == 0.9) %>%
  mutate(linetype = case_when(param_altered == "tau" ~ "2",
                              TRUE ~ "1")) %>%
  ggplot(aes(t, X, color = param_altered)) +
  geom_line(data = base_df, color = "black", size = 6) +
  geom_line(aes(lty = linetype), size = 3) +
  geom_point(data = hornig_basal, color = "black", size = 8) +
  scale_color_manual(values = unlist(param_pal), label = color_labels) +
  scale_linetype_manual(values = c("solid", "dashed")) + 
  theme1 +
  #guides(color = guide_legend(title = "Parameter inhibited (90%)",
  #                            nrow = 2, byrow = TRUE), lty = "none") +
  xlab("Time (h)") +
  ylab("Extracellular sFLT1 (ng/mL)") +
  #theme(legend.position = c(0.3, 0.8)) +
  theme(legend.position = "none") +
  annotate("text", x = 35, y = 0.8, label = parse(text = "alpha"), color = param_pal$alpha, size = 12) +
  annotate("text", x = 70, y = 1, label = parse(text = "beta"), color = param_pal$beta, size = 12) +
  annotate("text", x = 35, y = 4.8, label = parse(text = "gamma"), color = param_pal$gamma, size = 12) +
  annotate("text", x = 45, y = 3.5, label = parse(text = "delta"), color = param_pal$delta, size = 12) +
  annotate("text", x = 63, y = 3, label = parse(text = "kappa"), color = param_pal$kappa, size = 12) +
  annotate("text", x = 68, y = 3.1, label = parse(text = "tau"), color = param_pal$tau, size = 12) +
  annotate("text", x = 60, y = 1.6, label = "none", color = param_pal$none, size = 8)

X_tc_inh_90
```

```{r}
# save plot
ggsave(out_png("X_tc_inh_90"),
       X_tc_inh_90, width = 6, height = 4)
```


