---
title: "sFLT1 modeling global sensitivity analysis - Orders of magnitude in base parameters"
author: "Amy Gill"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## REVIEWED

This file imports data from sensitivity analysis of extracellular sFLT1 (X) and intracellular sFLT1 (I) at 18h and 72h performed 1/15/24 using the Matlab script "sFlt-model/dde_model/sens_analysis.m" with options:
- run_mode = "sens_logscale_sqrt10"
- genetic_chemical = "genetic"
- param_set = "hjk_opt_median_230821"

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../saved-figs/04_sensitivity/powers_sqrt_10/"
dir.create(outfile_prefix, recursive = TRUE)
```

## Load and wrangle files

### Simulation input file

```{r}
# import file of input parameters
param_file <- readMat("../../saved-data/04_sensitivity/powers_sqrt_10/input.mat")

# inspect available variables
names(param_file)
```

### Simulation output file

```{r}
# import file 
alt_file <- readMat("../../saved-data/04_sensitivity/powers_sqrt_10/sens_logscale_sqrt10_genetic.mat")

# inspect available variables
names(alt_file)
```

### Species names and parameter names

```{r}
# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")
param_names <- unlist(alt_file$param.names)
```

### Extract outputs to list

```{r}
alt_file_outlist <- alt_file$out[,,1]
names(alt_file_outlist) <- str_replace_all(names(alt_file_outlist), "\\.", "_")
names(alt_file_outlist)
```
```{r}
names(alt_file)
```



### Extract time points

```{r}
time_vec <- alt_file$sim.time[1,]
```

### Extract baseline time courses for X, I

```{r}
base_df <- data.frame(t = time_vec,
                      X = alt_file$base.sim.y[which(species_names == "X"),],
                      I = alt_file$base.sim.y[which(species_names == "I"),],
                      param_altered = "none")

base_df
```

### Extract time courses

```{r}
# format data frames of simulated species values for time courses
make_sim_Y_df <- function(Y_mat){
  
  # flatten matrix to data frame of time course data
  # (this isn't very "R-style", but it works, so tell Hadley to chill)
  for (i in 1:dim(Y_mat)[3]) {
    if (i == 1) {    # first run: initialize output data frame
      # extract matrix for first parameter set to a data frame
      Y_temp <- as.data.frame(Y_mat[,,1])
      names(Y_temp) <- species_names
      Y_temp$run_y <- 1
        
      # initialize full Y data frame to the time course from this first run
      Y_df <- Y_temp
      
    } else {    # later runs: append to data frame
      Y_temp <- as.data.frame(Y_mat[,,i])
      names(Y_temp) <- species_names
      Y_temp$run_y <- i
      Y_df <- rbind(Y_df, Y_temp)
    }
  }
  
  Y_df
  
}

# extract simulated time courses
Y_df <- make_sim_Y_df(alt_file$Y.all)
```

```{r}
Y_df
```

```{r}
num_params <- length(param_names)
num_factors <- length(alt_file$factors.to.try)
num_runs <- num_params * num_factors

Y_df <- Y_df %>%
  mutate(t = rep(time_vec, num_runs)) %>%
  select(t, run = run_y, everything())

Y_df 
```

### Extract parameter values

```{r}
# initial parameter values
param_file$p.init
```

```{r}
param_val_df <- as.data.frame(t(alt_file$p.mod.all))

names(param_val_df) <- param_names

param_val_df$run <- 1:nrow(param_val_df)

param_val_df
```


### Extract parameter alteration metadata

```{r}
# extract parameter altered
param_altered <- unlist(lapply(1:length(param_names), function(x){
  rep(param_names[x],9)
}))

# extract inhibition fraction
param_meta_df <- data.frame(run = as.numeric(1:num_runs),
                           param_altered = factor(param_altered,
                                                  levels = c("alpha", "beta", "gamma",
                                                             "delta", "tau")), 
                           scaling_numeric = rep(alt_file$factors.to.try, num_params)) %>%
  mutate(scaling_factor = factor(round(scaling_numeric, 3))) %>%
  left_join(param_val_df, by = "run") %>%
  mutate(ab = alpha*beta,
         bpg = beta + gamma,
         I_bar_calc = alpha / bpg,
         X_bar_calc = ab / (delta * bpg))

param_meta_df
```

### Merge time course data and metadata

```{r}
Y_df <- Y_df %>%
  left_join(param_meta_df)
```

```{r}
Y_df
```


### Extract 72h values from time courses

```{r}
# all species at 72h in all conditions
Y_72h <- Y_df %>%
  filter(t == 72)

# baseline X at 18h
X0_72h <- base_df %>%
  filter(t == 72) %>%
  pull(X)

# baseline I at 18h
I0_72h <- base_df %>%
  filter(t == 72) %>%
  pull(I)

# calculate X,I normalized to baseline 18h
Y_72h <- Y_72h %>%
  mutate(X_norm = X / X0_72h,
         I_norm = I / I0_72h)

Y_72h
```

### Extract 18h values from time courses

```{r}
# all species at 18h in all conditions
Y_18h <- Y_df %>%
  filter(t == 18)

# baseline X at 18h
X0_18h <- base_df %>%
  filter(t == 18) %>%
  pull(X)

# baseline I at 18h
I0_18h <- base_df %>%
  filter(t == 18) %>%
  pull(I)

# calculate X,I normalized to baseline 18h
Y_18h <- Y_18h %>%
  mutate(X_norm = X / X0_18h,
         I_norm = I / I0_18h)

Y_18h
```

### Extract 0h values from time courses

```{r}
# all species at 18h in all conditions
Y_0h <- Y_df %>%
  filter(t == 0)

# baseline X at 18h
X0_0h <- base_df %>%
  filter(t == 0) %>%
  pull(X)

# baseline I at 18h
I0_0h <- base_df %>%
  filter(t == 0) %>%
  pull(I)

# calculate X,I normalized to baseline 18h
Y_0h <- Y_0h %>%
  mutate(X_norm = X / X0_0h,
         I_norm = I / I0_0h)

Y_0h
```

### Hornig experimental data

```{r}
# load experimental data from Hornig
load("../../saved-data/hornig.rda")

hornig_basal <- hornig %>%
  filter(condition == "basal") %>%
  select(t = nominal_time, X = Sx_ng_ml)

hornig_basal
```





## Base plots

### Base X

```{r}
base_df %>%
  ggplot(aes(t, X)) +
  geom_line() +
  xlab("Time (h)") +
  ylab("Extracellular sFLT1 (X, ng/mL)") +
  theme1
```

```{r}
X_norm_val_72 <- base_df %>%
  filter(t == 72) %>%
  pull(X)

X_norm_val_72  # ng/mL
```


### Base I

```{r}
base_df %>%
  ggplot(aes(t, I)) +
  geom_line() +
  xlab("Time (h)") +
  ylab("Intracellular sFLT1 (I, #/cell)") +
  ylim(c(0, 1e6)) +
  theme1
```

```{r}
I_norm_val_72 <- base_df %>%
  filter(t == 72) %>%
  pull(I)

I_norm_val_72  # #/cell
```

### Plot X time course colored by scaling factor

#### Faceted in 2 rows

```{r}
X_facet_scale_sqrt10 <- Y_df %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_wrap(~ param_altered, scales = "free", labeller = label_parsed) +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_facet_scale_sqrt10"),
       X_facet_scale_sqrt10, width = 8, height = 4)

X_facet_scale_sqrt10
```

#### Faceted in 1 row

```{r}
X_facet_scale_sqrt10_1row <- Y_df %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_wrap(~ param_altered, scales = "free", labeller = label_parsed, nrow = 1) +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_facet_scale_sqrt10_1row"),
       X_facet_scale_sqrt10_1row, width = 16, height = 3)

X_facet_scale_sqrt10_1row
```


#### Grid (forces same y scale)

```{r}
X_facet_scale_sqrt10_1row_samescale <- Y_df %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_grid(. ~ param_altered, labeller = label_parsed) +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_facet_scale_sqrt10_1row_samescale"),
       X_facet_scale_sqrt10_1row_samescale, width = 16, height = 3)

X_facet_scale_sqrt10_1row_samescale
```

#### Grid with y log-scaled

```{r}
X_facet_scale_sqrt10_1row_samelog <- X_facet_scale_sqrt10_1row_samescale +
  scale_y_log10(, labels = function(x) sprintf("%g", x))

# save plot
ggsave(out_png("X_facet_scale_sqrt10_1row_samelog"),
       X_facet_scale_sqrt10_1row_samelog, width = 16, height = 3)

X_facet_scale_sqrt10_1row_samelog
```

### X time course: beta

Template for separately plotting any parameters

#### Powers of sqrt(10)

```{r}
X_beta_scale_sqrt10 <- Y_df %>%
  filter(param_altered == "beta") %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 2) +
  theme_paper +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  ggtitle(TeX(r'(Scaling $\beta$)')) +
  guides(color = guide_legend(title = "Scaling\nfactor")) +
  scale_color_manual(values = viridis(9)) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_beta_scale_sqrt10"),
       X_beta_scale_sqrt10, width = 6, height = 4)

X_beta_scale_sqrt10
```

#### Powers of 10

```{r}
X_beta_scale_10 <- Y_df %>%
  filter(param_altered == "beta" & scaling_numeric %in% 10^(-2:2)) %>%
  ggplot(aes(t, X, color = scaling_factor)) +
  geom_line(size = 2) +
  theme_paper +
  ylab("X (ng/mL)") +
  xlab("Time (h)") +
  ggtitle(TeX(r'(Scaling $\beta$)')) +
  guides(color = guide_legend(title = "Scaling\nfactor")) +
  scale_color_manual(values = cb_pal) +
  theme(legend.position = "right")

# save plot
ggsave(out_png("X_beta_scale_10"),
       X_beta_scale_10, width = 6, height = 4)

X_beta_scale_10
```

### Explore X at 72h

#### Raw values: Matrix with parameter altered x scaling factor

```{r}
Y_72h
```


```{r}
X_72h_df <- Y_72h %>%
  select(param_altered, scaling_numeric, X) %>%
  pivot_wider(names_from = param_altered, values_from = X) %>%
  mutate(scaling_factor = factor(round(scaling_numeric, 3)))

X_72h_df

X_72h_mat <- as.matrix(select(X_72h_df, alpha:tau))

rownames(X_72h_mat) <- X_72h_df$scaling_factor

X_72h_mat
```

#### Normalized values: Matrix with parameter altered x scaling factor, div original X_72h


```{r}
X_72h_mat / X_norm_val_72
```


### X (norm) at 72h versus scaling factor

#### Faceted in 2 rows

```{r}
X_72h_norm_df <- X_72h_df %>%
  pivot_longer(alpha:tau, names_to = "param_altered", values_to = "raw_value") %>%
  mutate(norm_value = raw_value / X0_72h,
         param_altered = factor(param_altered, levels = c("alpha", "beta", "gamma", "delta", "tau"))) 

X_72h_norm_v_scale_factor <- X_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_wrap(~ param_altered, scales = "free", labeller = "label_parsed") +
    ylab(TeX(r'($X_{72h}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("X_72h_norm_v_scale_factor"),
       X_72h_norm_v_scale_factor, width = 9, height = 5)

X_72h_norm_v_scale_factor
```


#### Faceted in 1 row

```{r}
X_72h_norm_v_scale_1row <- X_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_wrap(~ param_altered, scales = "free", labeller = "label_parsed", nrow = 1) +
    ylab(TeX(r'($X_{72h}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("X_72h_norm_v_scale_1row"),
       X_72h_norm_v_scale_1row, width = 16, height = 3)

X_72h_norm_v_scale_1row
```


#### Grid (forces same y scale)

```{r}
X_72h_norm_v_scale_1row_samescale <- X_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_grid(. ~ param_altered, labeller = "label_parsed") +
    ylab(TeX(r'($X_{72h}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("X_72h_norm_v_scale_1row_samescale"),
       X_72h_norm_v_scale_1row_samescale, width = 16, height = 3)

X_72h_norm_v_scale_1row_samescale
```

#### Grid with y log-scaled

```{r}
X_72h_norm_v_scale_1row_samelog <- X_72h_norm_v_scale_1row_samescale +
  scale_y_log10(, labels = function(x) sprintf("%g", x))

# save plot
ggsave(out_png("X_72h_norm_v_scale_1row_samelog"),
       X_72h_norm_v_scale_1row_samelog, width = 16, height = 3)

X_72h_norm_v_scale_1row_samelog
```

### Plot I time course (show steady state)

#### Faceted in 2 rows

```{r}
I_facet_scale_sqrt10 <- Y_df %>%
  ggplot(aes(t, I, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_wrap(~ param_altered, scales = "free", labeller = label_parsed) +
  ylab("I (#/cell)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_facet_scale_sqrt10"),
       I_facet_scale_sqrt10, width = 8, height = 4)

I_facet_scale_sqrt10
```

#### Faceted in 1 row

```{r}
I_facet_scale_sqrt10_1row <- Y_df %>%
  ggplot(aes(t, I, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_wrap(~ param_altered, scales = "free", labeller = label_parsed, nrow = 1) +
  ylab("I (#/cell)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_facet_scale_sqrt10_1row"),
       I_facet_scale_sqrt10_1row, width = 16, height = 3)

I_facet_scale_sqrt10_1row
```


#### Grid (forces same y scale)

```{r}
I_facet_scale_sqrt10_1row_samescale <- Y_df %>%
  ggplot(aes(t, I, color = scaling_factor)) +
  geom_line(size = 1) +
  facet_grid(. ~ param_altered, labeller = label_parsed) +
  ylab("I (#/cell)") +
  xlab("Time (h)") +
  scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right")

# save plot
ggsave(out_png("I_facet_scale_sqrt10_1row_samescale"),
       I_facet_scale_sqrt10_1row_samescale, width = 16, height = 3)

I_facet_scale_sqrt10_1row_samescale
```

#### Grid with y log-scaled

```{r}
I_facet_scale_sqrt10_1row_samelog <- I_facet_scale_sqrt10_1row_samescale +
  scale_y_log10(, labels = function(x) sprintf("%g", x))

# save plot
ggsave(out_png("I_facet_scale_sqrt10_1row_samelog"),
       I_facet_scale_sqrt10_1row_samelog, width = 16, height = 3)

I_facet_scale_sqrt10_1row_samelog
```



#### I at 72h (steady state)

```{r}
I_72h_df <- Y_72h %>%
  select(param_altered, scaling_factor, I, scaling_numeric) %>%
  pivot_wider(names_from = param_altered, values_from = I)

I_72h_mat <- as.matrix(select(I_72h_df, alpha:tau))

rownames(I_72h_mat) <- I_72h_df$scaling_factor

I_72h_df
```

#### Normalized values

```{r}
I_72h_norm_df <- I_72h_df %>%
  pivot_longer(alpha:tau, names_to = "param_altered", values_to = "raw_value") %>%
  mutate(norm_value = raw_value / I0_72h,
         param_altered = factor(param_altered, levels = c("alpha", "beta", "gamma", "delta", "tau"))) 

I_72h_norm_df
```


### I (norm) at 72h (ss) versus scaling factor

#### Faceted in 2 rows

```{r}
I_72h_norm_v_scale_factor <- I_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_wrap(~ param_altered, scales = "free", labeller = "label_parsed") +
    ylab(TeX(r'($I_{SS}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("I_72h_norm_v_scale_factor"),
       I_72h_norm_v_scale_factor, width = 9, height = 5)

I_72h_norm_v_scale_factor
```


#### Faceted in 1 row

```{r}
I_72h_norm_v_scale_1row <- I_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_wrap(~ param_altered, scales = "free", labeller = "label_parsed", nrow = 1) +
    ylab(TeX(r'($I_{SS}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("I_72h_norm_v_scale_1row"),
       I_72h_norm_v_scale_1row, width = 16, height = 3)

I_72h_norm_v_scale_1row
```


#### Grid (forces same y scale)

```{r}
I_72h_norm_v_scale_1row_samescale <- I_72h_norm_df %>%
  ggplot(aes(scaling_numeric, norm_value, color = scaling_factor)) +
  geom_point(size = 2) +
  facet_grid(. ~ param_altered, labeller = "label_parsed") +
    ylab(TeX(r'($I_{SS}$ (Normalized))')) +
  xlab("Scaling factor") +
  scale_x_log10(breaks = c(1e-2, 1, 1e2), labels = function(x) sprintf("%g", x)) +
  scale_color_manual(values = viridis(9)) +
  labs(color = "Scaling\nfactor") +
  theme_paper +
  theme(legend.position = "right",
        panel.spacing=unit(1.5,"lines"))

# save plot
ggsave(out_png("I_72h_norm_v_scale_1row_samescale"),
       I_72h_norm_v_scale_1row_samescale, width = 16, height = 3)

I_72h_norm_v_scale_1row_samescale
```

#### Grid with y log-scaled

```{r}
I_72h_norm_v_scale_1row_samelog <- I_72h_norm_v_scale_1row_samescale +
  scale_y_log10(, labels = function(x) sprintf("%g", x))

# save plot
ggsave(out_png("I_72h_norm_v_scale_1row_samelog"),
       I_72h_norm_v_scale_1row_samelog, width = 16, height = 3)

I_72h_norm_v_scale_1row_samelog
```


### Heatmap (x = parameter adjusted, y = scaling factor, color = X (ng/mL))


```{r}
Heatmap(X_72h_mat,
        heatmap_height = unit(4, "in"),
        heatmap_width = unit(4, "in"),
        column_names_gp = gpar(fontsize = 24),
        row_names_gp = gpar(fontsize = 14),
        cluster_columns = FALSE,
        cluster_rows = FALSE,
        column_names_rot = 0,
        column_names_side = "top",
        column_labels = expression(alpha, beta, gamma, delta, tau),
        rect_gp = gpar(col = "black", lwd = 2),
        col = viridis(10))
```



