---
title: "sFLT1 ODE modeling - Import simulation data"
author: "Amy Gill"
date: "2023-08-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file analyzes data from optimization of the 2 equation sFLT1 DDE model  (adding delay to secretion and intracellular degradation) to 1 pulse-chase and 2 constitutive secretion (Kinghorn, Hornig) time course datasets.

The simulation was originally performed 8/17/23 using the Matlab script "02_dde_model/Matlab/dde_driver.m" with following options: 

run_mode = "base_on_off";

This simulation uses 4 species (I, I_deg, X, X_deg) and 4 parameters ($\alpha=10,000$, $\beta=0.1$, $\gamma=0.1$, $\delta = 0.1$) for a period of 6 * T50_X on and 6 * T50_X off.

## Setup

### Packages, helpers, palettes, themes

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")

# define theme for exporting at a legible size (default)
theme_big <- theme1 +
  theme(legend.position = "bottom",
        axis.title.y = element_text(angle = 0, vjust = 0.5, size = 16),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 10))
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../saved-figs/02_dde_model/general/"
dir.create(outfile_prefix, recursive = TRUE)
```

## Import data

### Load files

```{r}
# specify file names
param_file_name <- "../../saved-data/02_dde_model/on_off_input.mat"
opt_file_name <- "../../saved-data/02_dde_model/on_off_sim.mat"

# import file data
param_file <- readMat(param_file_name)
opt_file <- readMat(opt_file_name)
```

### Extract parameter values

Not needed in current workflow - keeping for reference

```{r}
p <- param_file$p[,,1]

#p_df <- as.data.frame(p) %>%
#  pivot_longer(everything(), names_to = "parameter", values_to = "value")

p_list <- lapply(1:length(p), function(x){
  p[[x]][1,1]
})

names(p_list) <- names(p)
```

### Extract species and parameter names, key numbers

```{r}
# extract parameter names
param_names <- names(p)

# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")

# extract number of parameter sets
num_param_sets <- param_file$num.param.sets[1,1]

# extract number of time units simulated in each phase
num_time_units <- param_file$num.time.units[1,1]
```

### Calculate steady states

Not needed in current workflow because normalized in Matlab - keep for reference

```{r}
I_ss <- p_list$alpha / (p_list$beta + p_list$gamma)
X_ss <- p_list$beta * I_ss / p_list$delta

t50_I_ss <- log(2) / (p_list$beta + p_list$gamma)
t50_X_ss <- log(2) / p_list$delta
```

### Extract time course

```{r}
# extract time course data
# at t=0, num_time_units time units "on", then num_time_units time units "off"
Y_df <- data.frame(t = opt_file$pc.sim.time[1,] + num_time_units, 
                   I_norm = t(opt_file$pc.sim.I.norm),
                   X_norm = t(opt_file$pc.sim.X.norm))

# note "off" versus "on" period
Y_df <- Y_df %>%
  mutate(alpha_frac = ifelse(t <= num_time_units, 1, 0),
         order = 1,
         t_round = round(t, 3))

# note that the first two time points are intended to be before the "on" phase (draw base line)
Y_df$alpha_frac[1:2] <- 0

# note that the "on" point should be plotted after the "off" point at t=0
Y_df$order[3] <- 2

# shorten the time interval displayed before the "on" phase
Y_df$t[1] <- -2


# revise duplicated point where production switches off
ind <- which(Y_df$t_round == 12 & duplicated(Y_df$t_round))
Y_df$alpha_frac[ind] <- 0
Y_df$order[ind] <- 2


Y_df <- Y_df %>%
  arrange(t, order)
  
Y_df
```


## Plot time courses

### Plot alpha

```{r}
dde_sol_alpha_v_t <- Y_df %>%
  ggplot(aes(t, alpha_frac)) +
  #geom_vline(xintercept = 0, size = 0.5, lty = 2) +
  #geom_vline(xintercept = num_time_units, size = 0.5, lty = 2) +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_hline(yintercept = 1, size = 1, lty = 2, color = cb_pal[2]) +
  annotate("label", x = 17, y = 1, label = TeX(r'($\alpha_{max}$)'), 
           size = 8, color = cb_pal[2], family = "serif") +
  geom_line(color = cb_pal[param$alpha], size = 2) +
  theme_big +
  ylab(TeX(r'($\frac{\alpha}{\alpha_{max}}$)')) +
  xlab("Time (normalized)") +
  scale_x_continuous(limits = c(-2, 20), expand = c(0,0),
                     breaks = seq(0, 20, by = 2))

dde_sol_alpha_v_t
```


### Plot I

```{r}
dde_sol_I_v_t <- Y_df %>%
  ggplot(aes(t, I_norm)) +
  geom_vline(xintercept = 0, size = 1, lty = 2, color = cb_pal[param$alpha]) +
  geom_vline(xintercept = num_time_units, size = 1, lty = 2, color = cb_pal[param$alpha]) +
  geom_hline(yintercept = 1, size = 1, lty = 2, color = cb_pal[2]) +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_line(color = cb_pal[1], size = 1.5) +
  annotate("label", x = 17, y = 1, label = TeX(r'($I_{SS}$)'), 
           size = 8, color = cb_pal[2], family = "serif") +
  theme_big +
  ylab(TeX(r'($\frac{I}{I_{SS}}$)')) +
  xlab("Time (normalized)") +
  scale_y_continuous(breaks = seq(0, 2, by = 0.5)) +
  scale_x_continuous(limits = c(-2, 20), expand = c(0,0),
                     breaks = seq(0, 20, by = 2))
dde_sol_I_v_t
```

### Plot X

```{r}
dde_sol_X_v_t <- Y_df %>%
  ggplot(aes(t, X_norm)) +
  geom_vline(xintercept = 0, size = 1, lty = 2, color = cb_pal[param$alpha]) +
  geom_vline(xintercept = num_time_units, size = 1, lty = 2, color = cb_pal[param$alpha]) +
  geom_hline(yintercept = 1, size = 1, lty = 2, color = cb_pal[2]) +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_line(color = cb_pal[1], size = 1.5) +
  annotate("label", x = 17, y = 1, label = TeX(r'($X_{SS}$)'), 
           size = 8, color = cb_pal[2], family = "serif") +
  theme_big +
  ylab(TeX(r'($\frac{X}{X_{SS}}$)')) +
  xlab("Time (normalized)") +
  scale_y_continuous(breaks = seq(0, 1.5, by = 0.25)) +
  scale_x_continuous(limits = c(-2, 20), expand = c(0,0),
                     breaks = seq(0, 20, by = 2))

dde_sol_X_v_t
```

### Combined time courses

```{r}
dde_sol_timecourses <- (dde_sol_alpha_v_t +
  theme(axis.title.x = element_blank()) +
  scale_y_continuous(limits = c(0,1.2), breaks = seq(0,1,by=0.25))) /
(dde_sol_I_v_t +
  theme(axis.title.x = element_blank())) / 
  dde_sol_X_v_t

ggsave(out_png("dde_sol_timecourses"),
       dde_sol_timecourses, width = 4, height = 8)

dde_sol_timecourses
```

### Add reference points to time course plots

```{r}
t_annot <- seq(0,20,by=2)

Y_annot <- Y_df %>%
  filter(t_round %in% t_annot & !duplicated(t_round)) %>%
  mutate(viridis = factor(seq_along(1:length(t_annot))))

dde_sol_I_annot <- dde_sol_I_v_t +
  geom_point(data = Y_annot, aes(fill = viridis),
             shape = 21, col = "black", size = 5, show.legend = FALSE) +
  scale_fill_manual(values = viridis(length(t_annot)))
dde_sol_X_annot <- dde_sol_X_v_t +
  geom_point(data = Y_annot, aes(fill = viridis),
             shape = 21, col = "black", size = 5, show.legend = FALSE) +
  scale_fill_manual(values = viridis(length(t_annot)))

dde_sol_timecourses_annot <- (dde_sol_alpha_v_t +
  theme(axis.title.x = element_blank()) +
  scale_y_continuous(limits = c(0,1.2), breaks = seq(0,1,by=0.25)) +
  scale_x_continuous(limits = c(-2, 21), expand = c(0,0),
                     breaks = seq(0, 20, by = 2))) /
(dde_sol_I_annot +
  theme(axis.title.x = element_blank()) +
  scale_x_continuous(limits = c(-2, 21), expand = c(0,0),
                     breaks = seq(0, num_time_units*2, by = 2))) / 
  (dde_sol_X_annot +
  scale_x_continuous(limits = c(-2, 21), expand = c(0,0),
                     breaks = seq(0, 20, by = 2)))

ggsave(out_png("dde_sol_timecourses_annot"),
       dde_sol_timecourses_annot, width = 4, height = 8)

dde_sol_timecourses_annot
```

## Plot phase portrait

```{r}
dde_sol_phase_portrait <- Y_df %>%
  ggplot(aes(X_norm, I_norm)) +
  geom_abline(slope = 1, intercept = 0, lty = 2, size = 1.5, color = cb_pal[2]) +
  geom_path(size = 1) +
  #geom_path(data = filter(Y_df, t < 0.5), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 0.5), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 1.55), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 2.3), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 4.6), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 6.6), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 8.8), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 11), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 12.6), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 13.4), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 15), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 17), arrow = arrow()) +
  annotate("label", x = 0.4, y = 0.4, label = TeX(r'(I = \frac{\beta}{\delta}X)'), 
           color = cb_pal[2], size = 5, family = "serif") +
  annotate("label", x = 0.3, y = 1.2, label = "Increasing\nsFLT1", family = "serif") +
  annotate("label", x = 0.8, y = 0.3, label = "Decreasing\nsFLT1", family = "serif") +
  geom_point(data = Y_annot, aes(fill = viridis), alpha = 0.9,
             shape = 21, col = "black", size = 5, show.legend = FALSE) +
  scale_fill_manual(values = viridis(length(t_annot))) +
  theme_big +
  scale_x_continuous(breaks = seq(0, 1.2, by = 0.2)) +
  xlab(TeX(r'($\frac{X}{X_{SS}}$)')) +
  ylab(TeX(r'($\frac{I}{I_{SS}}$)'))

ggsave(out_png("dde_sol_phase_portrait"),
       dde_sol_phase_portrait, width = 4, height = 4)

dde_sol_phase_portrait
```


