---
title: "sFLT1 ODE modeling - Import simulation data"
author: "Amy Gill"
date: "2023-08-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file analyzes data from the general solution of the base (5 parameter) 2 equation sFLT1 DDE model at various values of maturation delay $\tau$.

The simulations were performed 8/18/23 using the Matlab script "sFlt-model/02_dde_model/base_dde_driver.m" with following options: 

run_mode = "vary_tau_on_off";
alpha = 10,000, beta = 0.1, gamma = 0.1, delta = 0.1,
taus = (1e-6:0.25:2) * t50_X;
num_time_points = 12;    # duration of pulse and chase in terms of X half-lives 

## Setup

### Packages, helpers, palettes, themes

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")

# define theme for exporting at a legible size (default)
theme_big <- theme1 +
  theme(legend.position = "bottom",
        axis.title.y = element_text(angle = 0, vjust = 0.5, size = 16),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 10))
```


### Set output directory

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/02_dde_model/toy_taus/"
dir.create(outfile_prefix, recursive = TRUE)
```

## Import data

### Load files

```{r}
# specify file names
param_file_name <- "../../../saved-data/02_dde_model/on_off_taus_input.mat"
sim_file_name <- "../../../saved-data/02_dde_model/on_off_sim_taus.mat"

# import file data
param_file <- readMat(param_file_name)
sim_file <- readMat(sim_file_name)
```

### Extract species and parameter names, key numbers

```{r}
# extract parameter names
param_names <- names(param_file$p[,,1])

# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")

# extract number of parameter sets
num_param_sets <- length(param_file$taus)

# extract conversion factor to move X concentrations (ng/mL) to corresponding #/cell
conv_factor_numcell_to_ngml <- param_file$conv.factor.ngml[1,1]

# extract number of time units
num_time_units <- param_file$num.time.units
```

### Extract initial parameter values

```{r}
# extract the sets of parameter values
p_vec <- unlist(param_file$p[,,1])

init_df <- data.frame(run = 1:num_param_sets,
                      alpha = p_vec["alpha"],
                      beta = p_vec["beta"],
                      gamma = p_vec["gamma"],
                      delta = p_vec["delta"],
                      tau = t(param_file$taus))

init_df <- init_df %>%
  mutate(t50_X = log(2)/delta,
         tau_raw = round(tau/t50_X,3),
         tau_raw_fac = factor(tau_raw))

init_df
```


```{r}
init_df_small <- init_df %>%
  filter(tau_raw %in% seq(0, 2, by = 0.5))

init_df_small
```



### Extract time points

```{r}
t_vec <- sim_file$pc.sim.time[1,] + c(num_time_units)
```

### Extract time courses

```{r}
I_df <- data.frame(sim_file$I.norm.all) %>%
  mutate(t = t_vec) %>%
  pivot_longer(X1:X9, names_to = "run", values_to = "I_norm") %>%
  mutate(run = as.integer(str_remove(run, "X")))

X_df <- data.frame(sim_file$X.norm.all) %>%
  mutate(t = t_vec) %>%
  pivot_longer(X1:X9, names_to = "run", values_to = "X_norm") %>%
  mutate(run = as.integer(str_remove(run, "X")))

make_tc_df <- function(in_df){
  
  bind_rows(lapply(1:num_param_sets, function(i){
  
  temp_df <- in_df %>%
    filter(run == i) %>%
    mutate(alpha_frac = ifelse(t <= c(num_time_units), 1, 0),
           order = 1)

  # note that the first two time points are intended to be before the "on" phase (draw base line)
  temp_df$alpha_frac[1:2] <- 0

  # note that the "on" point should be plotted after the "off" point at t=0
  temp_df$order[3] <- 2

  # shorten the time interval displayed before the "on" phase
  temp_df$t[1] <- -2

  # revise duplicated point where production switches off
  ind <- which(temp_df$t == c(num_time_units) & duplicated(temp_df$t))
  temp_df$alpha_frac[ind] <- 0
  temp_df$order[ind] <- 2
  
  temp_df <- temp_df %>%
    arrange(t, order)
  
  temp_df
}),
.id = "column_label")}

I_df <- make_tc_df(I_df)
X_df <- make_tc_df(X_df)

full_df <- I_df %>%
  left_join(X_df, by = c("column_label", "run", "t", "order", "alpha_frac")) %>%
  select(run, t, order, alpha_frac, I_norm, X_norm) %>%
  left_join(init_df, by = "run") %>%
  mutate(secr_flux_norm = beta * I_norm,
         int_degr_flux_norm = gamma * I_norm,
         ext_degr_flux_norm = delta * X_norm)

full_df
```


```{r}
full_df_small <- full_df %>%
  filter(tau_raw %in% seq(0, 2, by = 0.5))
```


## Plot time courses

### Plot alpha

```{r}
dde_sol_alpha_v_t <- full_df %>%
  ggplot(aes(t, alpha_frac)) +
  #geom_vline(xintercept = 0, size = 0.5, lty = 2) +
  #geom_vline(xintercept = num_time_units, size = 0.5, lty = 2) +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_hline(yintercept = 1, size = 1, lty = 2, color = cb_pal[2]) +
  annotate("label", x = 17, y = 1, label = TeX(r'($\alpha_{max}$)'), 
           size = 8, color = cb_pal[2], family = "serif") +
  geom_line(color = cb_pal[param$alpha], size = 1.5) +
  theme_paper +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, size = 16)) +
  ylab(TeX(r'($\frac{\alpha}{\alpha_{max}}$)')) +
  xlab("Time (normalized)") +
  scale_x_continuous(limits = c(-2, 20), expand = c(0,0),
                     breaks = seq(0, 20, by = 2)) +
  scale_y_continuous(expand = c(0.05, 0.05))

dde_sol_alpha_v_t
```


### Plot I

```{r}
dde_sol_I_v_t <- full_df_small %>%
  ggplot(aes(t, I_norm, color = factor(tau_raw))) +
  geom_vline(xintercept = 0, size = 1, lty = 2, color = cb_pal[param$alpha]) +
  geom_vline(xintercept = num_time_units, size = 1, lty = 2, color = cb_pal[param$alpha]) +
  geom_hline(yintercept = 1, size = 1, lty = 2, color = cb_pal[2]) +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_line(size = 1.5) +
  annotate("label", x = 17, y = 1, label = TeX(r'($I_{SS}$)'), 
           size = 8, color = cb_pal[2], family = "serif") +
  theme_paper +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, size = 16)) +
  ylab(TeX(r'($\frac{I}{I_{SS}}$)')) +
  xlab("Time (normalized)") +
  scale_y_continuous(breaks = seq(0, 3, by = 0.5)) +
  scale_x_continuous(limits = c(-2, 20), expand = c(0,0),
                     breaks = seq(0, 20, by = 2)) +
  scale_color_manual(values = viridis(nrow(init_df_small))) +
  labs(color = TeX(r'($\tau$)'))

dde_sol_I_v_t
```

### Plot X

```{r}
dde_sol_X_v_t <- full_df_small %>%
  ggplot(aes(t, X_norm, color = tau_raw_fac)) +
  geom_vline(xintercept = 0, size = 1, lty = 2, color = cb_pal[param$alpha]) +
  geom_vline(xintercept = num_time_units, size = 1, lty = 2, color = cb_pal[param$alpha]) +
  geom_hline(yintercept = 1, size = 1, lty = 2, color = cb_pal[2]) +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_line(size = 1.5) +
  annotate("label", x = 17, y = 1, label = TeX(r'($X_{SS}$)'), 
           size = 8, color = cb_pal[2], family = "serif") +
  theme_paper +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, size = 16)) +
  ylab(TeX(r'($\frac{X}{X_{SS}}$)')) +
  xlab("Time (normalized)") +
  scale_y_continuous(breaks = seq(0, 3, by = 0.5)) +
  scale_x_continuous(limits = c(-2, 20), expand = c(0,0),
                     breaks = seq(0, 20, by = 2)) +
  scale_color_manual(values = viridis(nrow(init_df_small))) +
  labs(color = TeX(r'($\tau$)'))

dde_sol_X_v_t
```

### Combined time courses

```{r}
dde_sol_timecourses <- (dde_sol_alpha_v_t +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24)) +
  scale_y_continuous(limits = c(0,1.1), breaks = seq(0,1,by=0.25))) /
  (dde_sol_I_v_t +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24),
        legend.position = "none")) / 
  (dde_sol_X_v_t +
 # plot_layout(guides = "collect") &
  theme(legend.position = "bottom",
        legend.title = element_text(size = 28),
        legend.text = element_text(size = 20),
        axis.title.y = element_text(size = 24)))

ggsave(out_png("dde_sol_timecourses"),
       dde_sol_timecourses, width = 6, height = 8)

dde_sol_timecourses
```



## Plot phase portraits





```{r}
dde_sol_phases <- full_df_small %>%
  ggplot(aes(X_norm, I_norm, color = tau_raw_fac)) +
  geom_abline(slope = 1, intercept = 0, lty = 2, size = 1.5, color = cb_pal[2]) +
  geom_path(size = 1.5) +
  theme_paper +
  scale_color_manual(values = viridis(nrow(init_df_small))) +
  labs(color = TeX(r'($\tau$)')) +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 28),
        legend.text = element_text(size = 20),
        axis.title.y = element_text(size = 24, angle = 0)) +
  ylab(TeX(r'($\frac{I}{I_{SS}}$)')) +
  xlab(TeX(r'($\frac{X}{X_{SS}}$)')) +
  geom_path(data = filter(full_df, tau_raw_fac == 2, t < 2.5), 
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 1.5, t < 2.25),
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 1, t < 1.9),
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 0.5, t < 1.6),
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 0, t < 1.8),
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 2, t < 12.6),
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 1.5, t < 13.2),
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 1, t < 13.4),
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 0.5, t < 15),
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  geom_path(data = filter(full_df, tau_raw_fac == 0, t < 18), 
            arrow = arrow(), size = 1.5, show.legend = FALSE) +
  annotate("label", x = 1.65, y = 1.65, label = TeX(r'(I = \frac{\beta}{\delta}X)'), 
           color = cb_pal[2], size = 5, family = "serif") 

ggsave(out_png("dde_sol_phases"),
       dde_sol_phases, width = 6, height = 8)

dde_sol_phases
```




### OLD

```{r}
dde_sol_phase_portrait <- Y_df %>%
  ggplot(aes(X_norm, I_norm)) +
  geom_abline(slope = 1, intercept = 0, lty = 2, size = 1.5, color = cb_pal[2]) +
  geom_path(size = 1) +
  #geom_path(data = filter(Y_df, t < 0.5), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 0.5), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 1.55), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 2.3), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 4.6), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 6.6), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 8.8), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 11), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 12.6), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 13.4), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 15), arrow = arrow()) +
  geom_path(data = filter(Y_df, t < 17), arrow = arrow()) +
  annotate("label", x = 0.4, y = 0.4, label = TeX(r'(I = \frac{\beta}{\delta}X)'), 
           color = cb_pal[2], size = 5, family = "serif") +
  annotate("label", x = 0.3, y = 1.2, label = "Increasing\nsFLT1", family = "serif") +
  annotate("label", x = 0.8, y = 0.3, label = "Decreasing\nsFLT1", family = "serif") +
  geom_point(data = Y_annot, aes(fill = viridis), alpha = 0.9,
             shape = 21, col = "black", size = 5, show.legend = FALSE) +
  scale_fill_manual(values = viridis(length(t_annot))) +
  theme_big +
  scale_x_continuous(breaks = seq(0, 1.2, by = 0.2)) +
  xlab(TeX(r'($\frac{X}{X_{SS}}$)')) +
  ylab(TeX(r'($\frac{I}{I_{SS}}$)'))

ggsave(out_png("dde_sol_phase_portrait"),
       dde_sol_phase_portrait, width = 4, height = 4)

dde_sol_phase_portrait
```
