---
title: "Preprocessing: Optimization of DDEs to sFLT1 datasets from Hornig, Pulse-chase and Kinghorn"
author: "Amy Gill"
date: "2023-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file analyzes data from optimization of the 2 equation sFLT1 DDE model to the pulse-chase and constitutive time course datasets. These data were imported using the script "sFlt-model/analysis-R/draft/02_dde_model/dde_opt_import.Rmd".

The optimization was performed 8/21/23 using the Matlab script "sFlt-model/02_dde_model/base_dde_driver.m" with following options: 

run_mode = "base_model_opt";
jung_cost_option = "omit_last_Sx_relative";
hornig_cost_option = "relative_omit3h";
kinghorn_cost_option = "all_pts_relative";
num_param_sets = 1000;    (# per candidate model)

Optimizing to summary statistics only (mean at each time point) for Kinghorn optimization.

This optimization uses 4 species (I, I_deg, X, X_deg) and optimizes 5 parameters ($\alpha$, $\beta$, $\gamma$, $\delta$, $\tau$) by calculating relative cost in comparison to all of the following datasets:

- Jung X (normalized to 8h)
- Jung I (normalized to 0h)
- Hornig basal X concentration (absolute, ng/mL -> #/cell)
- Kinghorn X (normalized to 24h)
- Kinghorn I (normalized to 0h)

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/02_dde_model/dde_opt_v2/"
dir.create(outfile_prefix, recursive = TRUE)

# specify decimal resolution
options(digits = 4)
```

### Import data

#### Optimization runs

```{r}
# load optimization results from 8/21
# (if file is not present, run "dde_opt_import.Rmd" to create it)
load("../../../saved-data/02_dde_model/dde_opt_sims.rda")
```

### Add fill color columns to experimental data

```{r}
jung <- jung %>%
  mutate(fill_col = factor("Jung", levels = c("Jung", "Kinghorn", "Hornig")))

kinghorn <- kinghorn %>%
  mutate(fill_col = factor("Kinghorn", levels = c("Jung", "Kinghorn", "Hornig")))

hornig <- hornig %>%
  mutate(fill_col = factor("Hornig", levels = c("Jung", "Kinghorn", "Hornig")))
```

## Analysis before preprocessing

### Create formatted data frames for analysis

```{r}
# param_df: a 4-column data frame
# run = run number 1:1000
# parameter = factor, parameter definition(alpha, beta, gamma, delta, epsilon, kappa, tau)
# init_opt = factor, INIT = initial guess, OPT = optimized value
# value = parameter value
param_df <- meta_df %>%
  select(run:OPT_tau) %>%
  pivot_longer(INIT_alpha:OPT_tau, names_to = "parameter", values_to = "value") %>%
  separate(parameter, into = c("init_opt", "parameter"), sep = "_") %>%
  mutate(init_opt = ifelse(init_opt == "INIT", "initial", "optimal"),
                           init_opt = factor(init_opt),
         parameter = factor(parameter,
                            levels = c("alpha", "beta", "gamma", "delta", "tau"))) 

# add cost to param_df
param_df <- meta_df %>%
  select(run, cost = total_cost) %>%
  right_join(param_df, by = "run")

# extract parameter names
param_names <- levels(param_df$parameter)
```


### Cost distribution

#### Density plot

```{r}
# create function to plot density function of cost
plot_cost_distr_density <- function(df){
  df %>%
    ggplot(aes(total_cost, ..scaled..)) +
    geom_density(fill = cb_pal[3]) +
    scale_x_log10() +
    #scale_x_continuous(trans = "log10", breaks = 10^(-2:5)) +
    theme1 +
    xlab("Sum of squared difference") +
    ylab("Scaled density")
}

# plot cost distribution
combo_cost_distr_unfiltered <- plot_cost_distr_density(meta_df) +
    ggtitle("Cost function values for optimal parameters - all fits")
  
# save plot
ggsave(out_png("combo_cost_distr_unfiltered"),
       combo_cost_distr_unfiltered, width = 9, height = 6)
  
combo_cost_distr_unfiltered
```

#### Violin plot

```{r}
# create function to make violin plot of cost distribution
plot_cost_violin <- function(df){
  df %>%
    ggplot(aes(factor(" "), total_cost)) +
    geom_violin(scale = "width", fill = cb_pal[2]) +   # not trimming b/c hard min value 
    geom_boxplot(width = 0.3, fill = "white") +
    scale_y_continuous(trans = "log10", breaks = 10^(-2:5)) +
    xlab("") +
    ylab("Sum of squared difference") +
    coord_flip() +
    theme1
}

cost_violin <- plot_cost_violin(meta_df) +
  ggtitle("Violin plot of cost function values for optimal parameters")

cost_violin
```

### Raw time courses

Individual time courses are colored by the sum of squared differences (SSD) between simulated and observed measurements of both I and X at the experimental time points.

#### Kinghorn X

Some fits have interesting oscillations. For future analysis, impose a cost filter to refine to top fits.

```{r}
constitutive_full_df %>%
  ggplot(aes(t, X_frac)) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.25) +
  geom_point(data = kinghorn, aes(time, X_frac_24h), size = 5, 
               shape = 21, color = "white", fill = cb_pal[2]) +
    xlab("Time (h)") +
    ylab("X (Normalized to 24h)") +
    labs(color = "log2(Cost)") +
    theme1
```


#### Kinghorn I

Some fits oscillate, others appear roughly constant at the steady state value.

```{r}
constitutive_opt_I_timecourse_unfiltered <- constitutive_full_df %>%
  ggplot(aes(t, I_fc)) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.25) +
  geom_point(data = kinghorn, aes(time, I_fc_0h), size = 5, 
               shape = 21, color = "white", fill = cb_pal[2]) +
    xlab("Time (h)") +
    ylab("I (Normalized to 0h)") +
    labs(color = "log2(Cost)") +
    theme1

ggsave(out_png("constitutive_opt_I_timecourse_unfiltered"),
       constitutive_opt_I_timecourse_unfiltered, width = 8, height = 6)

constitutive_opt_I_timecourse_unfiltered
```

##### "Poster version"

```{r}
kinghorn_opt_I_timecourse_poster <- constitutive_full_df %>%
  ggplot(aes(t, I_fc)) +
    geom_line(aes(group = run, color = log2(total_cost)), size = 1,  alpha = 0.25) +
  geom_point(data = kinghorn, aes(time, I_fc_0h, fill = fill_col), size = 5, 
               shape = 21, color = "white") +
  scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
  xlab("Time (h)") +
  scale_x_continuous(limits = c(-1, 72.2), breaks = seq(0, 72, by = 12), expand = c(0,0.3)) +
    ylab(TeX(r'($\frac{I}{I_{0h}}$)')) +
    theme_poster

# save plot
ggsave(out_png("time_courses/kinghorn_opt_I_timecourse_poster_unfiltered"),
       kinghorn_opt_I_timecourse_poster, width = 6, height = 4)

kinghorn_opt_I_timecourse_poster
```

#### Constitutive X

It is difficult to make out the properties of the low-cost fits due to several fits with very large errors. Like the I time courses, there is a mix of oscillating and monotonically increasing X time courses.

```{r}
constitutive_plot_X_timecourse_ngml <- function(df){
  df %>%
    ggplot(aes(t, X)) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.25) +
    geom_point(data = hornig, aes(nominal_time, X_ng_ml), size = 5, 
               shape = 21, color = "white", fill = cb_pal[2]) +
    xlab("Time (h)") +
    ylab("X (ng/mL)") +
    labs(color = "log2(Cost)") +
    theme1
}

constitutive_opt_X_timecourse_ngml_unfiltered <- constitutive_plot_X_timecourse_ngml(constitutive_full_df) +
  ggtitle("Constitutive extracellular sFLT1 (X) - All fits")

ggsave(out_png("constitutive_opt_X_timecourse_ngml_unfiltered"),
       constitutive_opt_X_timecourse_ngml_unfiltered, width = 8, height = 6)

constitutive_opt_X_timecourse_ngml_unfiltered
```

##### "Poster version"

```{r}
const_opt_X_timecourse_combo_poster <- constitutive_full_df %>%
    ggplot(aes(t, X / X_24)) +
    geom_line(aes(group = run, color = log2(total_cost)), size = 1, alpha = 0.25) +
    geom_point(data = hornig, aes(nominal_time, X_norm_24, fill = fill_col), 
               size = 5, shape = 21, color = "white") +
#    geom_line(data = hornig, aes(nominal_time, X_norm_24), linewidth = 1, color = cb_pal[2]) 
  geom_point(data = kinghorn, aes(time, X_frac_24h, fill = fill_col), 
             size = 5, shape = 21, color = "white", show.legend = FALSE) +
#  geom_line(data = kinghorn, aes(time, X_frac_24h), linewidth = 1, color = cb_pal[7]) +
  scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
  xlab("Time (h)") +
    theme_poster +
  scale_x_continuous(limits = c(-1, 73.5), breaks = seq(0, 72, by = 12), expand = c(0,0)) +
  scale_y_continuous(name = TeX(r'($\frac{X}{X_{24h}}$)'),
                     sec.axis = sec_axis(trans = ~.*hornig_X_numcell_24,
                                         name = "X\n(#/cell)")) +
  theme(axis.title.y.right = element_text(angle = 0, vjust = 0.5, size = 16))

ggsave(out_png("time_courses/const_opt_X_timecourse_combo_poster_unfiltered"),
       const_opt_X_timecourse_combo_poster, width = 6, height = 4)

const_opt_X_timecourse_combo_poster
```

#### Pulse-chase X

```{r}
pulsechase_plot_X_frac_timecourse <- function(df){
  df %>%
    ggplot(aes(t, X_frac)) +
    annotate("segment", x = -0.7, xend = -0.17, y = 1.05, yend = 0.65, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 1.14, yend = 1.14, 
             lty = 2, size = 0.5) +
    annotate("rect", xmin = 3, xmax = 5, ymin = 1.13, ymax = 1.17, fill = "white", color = "white") +
    annotate("text", x = -0.65, y = 1.15, label = "P", size = 5.5, angle = 0, fontface = "bold") +
    annotate("text", x = 4, y = 1.15, label = "chase", size = 5) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    geom_point(data = jung, aes(time, media_frac_max, fill = "temp"),
               size = 5, shape = 21, color = "white",
               show.legend = TRUE) +
    scale_fill_manual(labels = "X", values = cb_pal[2]) +

    ylim(c(-0.05, 1.2)) +
    xlim(c(-1, 10)) +
    #scale_color_viridis_b() +
    xlab("Time (h)") +
    ylab("X (Normalized to 8h)") +
    labs(color = "log2(Cost)",
         fill = "Observed data") +
    theme1 +
    theme(plot.title = element_text(size = 16))
}

pulsechase_opt_X_timecourse_unfiltered <- pulsechase_plot_X_frac_timecourse(pulsechase_full_df) +
  ggtitle("Pulse-chase extracellular sFLT1 (X) - All fits")

# save plot
ggsave(out_png("pulsechase_opt_X_timecourse_unfiltered"),
       pulsechase_opt_X_timecourse_unfiltered, width = 8, height = 6)

pulsechase_opt_X_timecourse_unfiltered
```


##### "Poster version"

```{r}
jung_opt_X_timecourse_poster <- pulsechase_full_df %>%
  ggplot(aes(t, X_frac)) +
    annotate("rect", xmin = -20/60, xmax = 0, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    #annotate("rect", xmin = 0, xmax = Inf, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.1, size = 1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    #annotate("text", x = -0.75, y = 0.55, label = "Start pulse", angle = 90, size = 6) +
    #annotate("text", x = 0.35, y = 0.75, label = "Start chase", angle = 90, size = 6) +
    annotate("segment", x = -0.7, xend = -0.17, y = 7.3, yend = 5, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 10, yend = 10, 
             lty = 2, size = 0.5) +
    annotate("rect", xmin = 5.5, xmax = 8.5, ymin = 9.95, ymax = 10.05, fill = "white", color = "white") +
    annotate("text", x = -0.8, y = 7.7, label = "P", size = 5, angle = 0, fontface = "bold") +
    annotate("text", x = 7, y = 10, label = "chase", size = 5.5) +
    geom_point(data = jung, aes(time, media_frac_max, fill = fill_col),
               size = 5, shape = 21, color = "white",
               show.legend = TRUE) +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
    labs(fill = "Experimental data") +    
    #scale_y_continuous(limits = c(-0.05, 1.2), breaks = seq(0, 1, by = 0.25)) +
    scale_x_continuous(limits = c(-1, 10), breaks = seq(0, 10, by = 2), expand = c(0,.1)) +
    #scale_color_viridis_b() +
    xlab("Time (h)") +
    ylab(TeX(r'($\frac{X}{X_{8h}}$)')) +
  theme_poster


# save plot
ggsave(out_png("time_courses/jung_opt_X_timecourse_poster"),
       jung_opt_X_timecourse_poster, width = 6, height = 4)

jung_opt_X_timecourse_poster
```


#### Pulse-chase I

```{r}
pulsechase_plot_I_fc_timecourse <- function(df){
  df %>%
    ggplot(aes(t, I_fc)) +
    annotate("rect", xmin = -20/60, xmax = 0, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    geom_point(data = jung, aes(time, lysate_fc, fill = "temp"),
               size = 5, shape = 21, color = "white", show.legend = TRUE) +
    scale_fill_manual(labels = "I", values = cb_pal[1]) +
    #scale_color_viridis_b() +
    xlab("Time (h)") +
    ylab("I (Normalized to 0h)") +
    labs(color = "log2(Cost)",
         fill = "") +
    #ylim(c(-0.05, 5)) +
    xlim(c(-1, 10)) +
    theme1 +
    theme(plot.title = element_text(size = 16))
}

pulsechase_opt_I_timecourse_unfiltered <- pulsechase_plot_I_fc_timecourse(pulsechase_full_df) +
    ggtitle("Pulse-chase intracellular sFlt1 (I) - All fits") +
    annotate("segment", x = -0.7, xend = -0.17, y = 1.05, yend = 0.65, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 1.2, yend = 1.2, 
             lty = 2, size = 0.5) +
    annotate("rect", xmin = 3, xmax = 5, ymin = 1.18, ymax = 1.21, fill = "white", color = "white") +
    annotate("text", x = -0.65, y = 1.15, label = "P", size = 5.5, angle = 0, fontface = "bold") +
    annotate("text", x = 4, y = 1.21, label = "chase", size = 5)
    #annotate("text", x = -0.65, y = 0.25, label = "Start pulse", angle = 90, size = 6) +
    #annotate("text", x = 0.25, y = 0.5, label = "Start chase", angle = 90, size = 6)

# save plot
ggsave(out_png("pulsechase_opt_I_timecourse_unfiltered"),
       pulsechase_opt_I_timecourse_unfiltered, width = 8, height = 6)

pulsechase_opt_I_timecourse_unfiltered
```

##### "Poster version"

```{r}
jung_opt_I_timecourse_poster <- pulsechase_full_df %>%
  ggplot(aes(t, I_fc)) +
    annotate("rect", xmin = -20/60, xmax = 0, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.1, size = 1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    annotate("segment", x = -0.7, xend = -0.2, y = 1.1, yend = 0.8, size = 0.5) +
    annotate("segment", x = 0.05, xend = 10.15, y = -0.05, yend = -0.05, lty = 2, size = 0.5) +
    annotate("rect", xmin = 0.5, xmax = 3.5, ymin = -.05, ymax = -.03, fill = "white", color = "white") +
    annotate("text", x = -0.7, y = 1.15, label = "P", size = 5, angle = 0, fontface = "bold") +
    annotate("text", x = 2, y = -0.04, label = "chase", size = 5.5) +
    geom_point(data = jung, aes(time, lysate_fc, fill = fill_col),
               size = 5, shape = 21, color = "white", show.legend = TRUE) +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
    xlab("Time (h)") +
    ylab(TeX(r'($\frac{I}{I_{0h}}$)')) +
    scale_y_continuous(limits = c(-0.05, 1.2), breaks = seq(0, 1, by = 0.25)) +
    scale_x_continuous(limits = c(-1, 10.2), breaks = seq(0, 10, by = 2), expand = c(0,0)) +
    theme_poster +
  guides(color=guide_colorbar(barwidth=10))

# save plot
ggsave(out_png("time_courses/jung_opt_I_timecourse_poster_unfiltered"),
       jung_opt_I_timecourse_poster, width = 6, height = 4)

jung_opt_I_timecourse_poster
```

### Combined figure

```{r}
p1 <- jung_opt_X_timecourse_poster
p2 <- jung_opt_I_timecourse_poster
p3 <- const_opt_X_timecourse_combo_poster
p4 <- kinghorn_opt_I_timecourse_poster

dde_combo_plot_poster <- (p1 + p3) / (p2 + p4) +
  plot_layout(guides = "collect") &
  theme(plot.title = element_blank(), 
        legend.position = "bottom",
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(angle = 0, vjust = 0.5, size = 20),
        axis.title.y.right = element_text(angle = 0, vjust = 0.5, size = 20),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 14)) &
  guides(color = guide_colorbar(barwidth = 10))

ggsave(out_png("time_courses/dde_combo_plot_poster_unfiltered"),
       dde_combo_plot_poster, width = 9, height = 6)

dde_combo_plot_poster
```


## Preprocessing

### Define runs to remove

#### Runs that returned NaN cost 

```{r}
runs_to_remove_na <- meta_df %>%
  filter(is.nan(c(total_cost))) %>%
  pull(run)

length(runs_to_remove_na)
```

#### Negative species - Constitutive

```{r}
# identify runs where either species goes below 0
runs_neg_constitutive <- constitutive_full_df %>%
  filter(X < 0 | I < 0) %>%
  pull(run) %>%
  unique()

length(runs_neg_constitutive)
```

#### Negative species - Pulse-chase

```{r}
# identify runs where either species goes below 0
runs_neg_pulsechase <- pulsechase_full_df %>%
  filter(X < 0 | I < 0) %>%
  pull(run) %>%
  unique()

length(runs_neg_pulsechase)
```

### Remove runs

```{r}
# remove runs from constitutive simulated time course
constitutive_sims <- constitutive_full_df %>%
  filter(! run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive) %>%
  mutate(t = round(t,3))

# remove runs from pulse-chase simulated time course
pulsechase_sims <- pulsechase_full_df %>%
  filter(!run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive) %>%
  mutate(t = round(t,3))

# remove runs from meta_base_df
meta_df_filtered <- meta_df %>%
    filter(!run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive)

# remove runs from opt_df
opt_df_filtered <- opt_df %>%
  filter(!run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive)
  
# remove prefix from parameter names in opt_df_filtered
names(opt_df_filtered) <- str_remove(names(opt_df_filtered), "^OPT_")

# remove runs from param_df
param_df_filtered <- param_df %>%
  filter(!run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive)

length(unique(param_df_filtered$run))
```


### Cost filtering

The best fitting parameter set had a cost of 1.108. We refined our simulations to fits with a cost within 10% of the minimum cost set.

```{r}
min(meta_df_filtered$total_cost)

cost_cutoff <- 1.1*min(meta_df_filtered$total_cost)

sum(meta_df_filtered$total_cost < cost_cutoff)
```

```{r}
# list low-cost runs from meta_df_filtered
runs_to_keep <- meta_df_filtered %>%
  filter(total_cost < cost_cutoff) %>%
  pull(run)

# remove runs from constitutive simulated time course
constitutive_sims <- constitutive_sims %>%
  filter(run %in% runs_to_keep)

# remove runs from pulse-chase simulated time course
pulsechase_sims <- pulsechase_sims %>%
  filter(run %in% runs_to_keep)

# remove runs from meta_df
meta_df_filtered <- meta_df_filtered %>%
  filter(run %in% runs_to_keep)

# remove runs from opt_df
opt_df_filtered <- opt_df_filtered %>%
  filter(run %in% runs_to_keep)

# remove runs from param_df
param_df_filtered <- param_df_filtered %>%
  filter(run %in% runs_to_keep)

# create in wide format - separate column for each param for both init and opt
param_df_wide <- param_df_filtered %>%
  filter(!(parameter %in% c("epsilon", "kappa"))) %>%
  mutate(param_init_opt = paste0(ifelse(init_opt == "initial", "INIT_", "OPT_"), parameter)) %>%
  select(run, param_init_opt, value) %>%
  pivot_wider(names_from = "param_init_opt", values_from = "value")

# init-opt split but params in shared column
param_df_init_opt_cols <- param_df_filtered %>%
  pivot_wider(names_from = "init_opt", values_from = "value")

# params split but init-opt in shared column
param_df_param_cols <- param_df_filtered %>%
  pivot_wider(names_from = "parameter", values_from = "value") %>%
  mutate(alpha_beta = alpha*beta,
         beta_plus_gamma = beta + gamma)
```

#### Plot raw distribution with line for filtering cutoff

```{r}
p1 <- plot_cost_distr_density(meta_df) +
  geom_vline(xintercept = 1.108*1.1, color = cb_pal[2])

p1
```


#### Plot filtered distribution

```{r}
p2 <- plot_cost_distr_density(meta_df_filtered) +
  xlim(c(1.1, 1.108*1.1)) +
  geom_vline(xintercept = 1.108*1.1, color = cb_pal[2])

p2
```

#### Plot cost distribution with inset of filtered distribution

```{r}
cost_density_inset <- p1 + 
  annotate("rect", xmin = 4.5, xmax = 5000, ymin = 0.15, ymax = 0.95, color = cb_pal[1], fill = NA) +
  inset_element(p2, left = 0.2, bottom = 0.2, right = 0.9, top = 0.9)

ggsave(out_png("cost_density_inset"),
       cost_density_inset, width = 6, height = 4)

cost_density_inset
```


#### Cost versus run number

```{r}
# visually inspect cost vs run number
# (densities and boxplots were surprisingly uninformative bc of tight distribution)
meta_df_filtered %>%
  ggplot(aes(as.numeric(run), total_cost)) +
  geom_point(alpha = 0.3) +
  scale_x_continuous(n.breaks = 11) +
  theme1 +
  xlab("Run number") +
  ylab("Sum of squared difference") +
  ggtitle("Cost of fit by run")
```

```{r}

```


## Save objects

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix <- "../../../saved-data/02_dde_model/"
dir.create(outfile_prefix, recursive = TRUE)

save(param_df, param_df_filtered, meta_df_filtered, opt_df_filtered, param_df_wide,
     param_df_init_opt_cols, param_df_param_cols, constitutive_sims, pulsechase_sims,
     file = paste0(outfile_prefix, "dde_opt_sims_processed.rda"))
```