---
title: "Preprocessing: Optimization of DDEs to sFLT1 datasets from Hornig, Pulse-chase and Kinghorn"
author: "Amy Gill"
date: "2023-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file analyzes data from optimization of the 2 equation sFLT1 DDE model to the pulse-chase and constitutive time course datasets. These data were imported using the script "03_candidate_models/R/candidate_models_import.Rmd".

The optimization was originally performed 8/23/23 using the Matlab script "03_candidate_models/Matlab/dde_candidate_models.m" with following options: 

run_mode = "candidate_model_opt";
jung_cost_option = "omit_last_Sx_relative";
hornig_cost_option = "relative_omit3h";
kinghorn_cost_option = "all_pts_relative";
num_param_sets = 100;    (# per candidate model)

Optimizing to summary statistics only (mean at each time point) for Kinghorn optimization.

This optimization uses 4 species (I, I_deg, X, X_deg) and optimizes 7 parameters ($\alpha$, $\beta$, $\gamma$, $\delta$, $\epsilon$, $\kappa$, $\tau$) by calculating absolute cost in comparison to all of the following datasets:

- Jung X (normalized to 8h)
- Jung I (normalized to 0h)
- Hornig basal X concentration (absolute, ng/mL -> #/cell)
- Kinghorn X (normalized to 24h)
- Kinghorn I (normalized to 0h)

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../saved-figs/03_candidate_models/"
dir.create(outfile_prefix, recursive = TRUE)

# specify decimal resolution
options(digits = 4)
```

### Import data

#### Optimization runs

```{r}
# load optimization results from 8/23
# (if file is not present, run "candidate_models_import.Rmd" to create it)
load("../../saved-data/03_candidate_models/pulsechase_constitutive_opt_sims.rda")
```

## Analysis before preprocessing

### Create formatted data frames for analysis

```{r}
# param_df: a 4-column data frame
# run = run number 1:1000
# parameter = factor, parameter definition(alpha, beta, gamma, delta, epsilon, kappa, tau)
# init_opt = factor, INIT = initial guess, OPT = optimized value
# value = parameter value
param_df <- meta_df %>%
  select(run:OPT_tau) %>%
  pivot_longer(INIT_alpha:OPT_tau, names_to = "parameter", values_to = "value") %>%
  separate(parameter, into = c("init_opt", "parameter"), sep = "_") %>%
  mutate(init_opt = ifelse(init_opt == "INIT", "initial", "optimal"),
                           init_opt = factor(init_opt),
         parameter = factor(parameter,
                            levels = c("alpha", "beta", "gamma", "delta", 
                                       "epsilon", "kappa", "tau"))) 

# add cost to param_df
param_df <- meta_df %>%
  select(run, cost = total_cost) %>%
  right_join(param_df, by = "run")

# extract parameter names
param_names <- levels(param_df$parameter)
```


### Cost distribution

#### Density plot

```{r}
# create function to plot density function of cost
plot_cost_distr_density <- function(df){
  df %>%
    ggplot(aes(total_cost, ..scaled..)) +
    geom_density(fill = cb_pal[3]) +
    scale_x_log10() +
    #scale_x_continuous(trans = "log10", breaks = 10^(-2:5)) +
    theme1 +
    xlab("Sum of squared difference") +
    ylab("Scaled density")
}

# plot cost distribution
combo_cost_distr_unfiltered <- plot_cost_distr_density(meta_df) +
    ggtitle("Cost function values for optimal parameters - all fits")
  
# save plot
ggsave(out_png("combo_cost_distr_unfiltered"),
       combo_cost_distr_unfiltered, width = 9, height = 6)
  
combo_cost_distr_unfiltered
```

#### Violin plot

```{r}
# create function to make violin plot of cost distribution
plot_cost_violin <- function(df){
  df %>%
    ggplot(aes(factor(" "), total_cost)) +
    geom_violin(scale = "width", fill = cb_pal[2]) +   # not trimming b/c hard min value 
    geom_boxplot(width = 0.3, fill = "white") +
    scale_y_continuous(trans = "log10", breaks = 10^(-2:5)) +
    xlab("") +
    ylab("Sum of squared difference") +
    coord_flip() +
    theme1
}

cost_violin <- plot_cost_violin(meta_df) +
  ggtitle("Violin plot of cost function values for optimal parameters")

cost_violin
```

### Raw time courses

Individual time courses are colored by the sum of squared differences (SSD) between simulated and observed measurements of both I and X at the experimental time points.

#### Kinghorn X

Some fits have interesting oscillations. For future analysis, impose a cost filter to refine to top fits.

```{r}
constitutive_full_df %>%
  ggplot(aes(t, X_frac)) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.25) +
  geom_point(data = kinghorn, aes(time, X_frac_24h), size = 5, 
               shape = 21, color = "white", fill = cb_pal[2]) +
    xlab("Time (h)") +
    ylab("X (Normalized to 24h)") +
    labs(color = "log2(Cost)") +
    theme1
```


#### Kinghorn I

Some fits oscillate, others appear roughly constant at the steady state value.

```{r}
constitutive_opt_I_timecourse_unfiltered <- constitutive_full_df %>%
  ggplot(aes(t, I_fc)) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.25) +
  geom_point(data = kinghorn, aes(time, I_fc_0h), size = 5, 
               shape = 21, color = "white", fill = cb_pal[2]) +
    xlab("Time (h)") +
    ylab("I (Normalized to 0h)") +
    labs(color = "log2(Cost)") +
    theme1

ggsave(out_png("constitutive_opt_I_timecourse_unfiltered"),
       constitutive_opt_I_timecourse_unfiltered, width = 8, height = 6)

constitutive_opt_I_timecourse_unfiltered
```

#### Constitutive X

It is difficult to make out the properties of the low-cost fits due to several fits with very large errors. Like the I time courses, there is a mix of oscillating and monotonically increasing X time courses.

```{r}
constitutive_plot_X_timecourse_ngml <- function(df){
  df %>%
    ggplot(aes(t, X)) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.25) +
    geom_point(data = hornig, aes(nominal_time, X_ng_ml), size = 5, 
               shape = 21, color = "white", fill = cb_pal[2]) +
    xlab("Time (h)") +
    ylab("X (ng/mL)") +
    labs(color = "log2(Cost)") +
    theme1
}

constitutive_opt_X_timecourse_ngml_unfiltered <- constitutive_plot_X_timecourse_ngml(constitutive_full_df) +
  ggtitle("Constitutive extracellular sFLT1 (X) - All fits")

ggsave(out_png("constitutive_opt_X_timecourse_ngml_unfiltered"),
       constitutive_opt_X_timecourse_ngml_unfiltered, width = 8, height = 6)

constitutive_opt_X_timecourse_ngml_unfiltered
```

#### Pulse-chase X

```{r}
pulsechase_plot_X_frac_timecourse <- function(df){
  df %>%
    ggplot(aes(t, X_frac)) +
    annotate("segment", x = -0.7, xend = -0.17, y = 1.05, yend = 0.65, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 1.14, yend = 1.14, 
             lty = 2, size = 0.5) +
    annotate("rect", xmin = 3, xmax = 5, ymin = 1.13, ymax = 1.17, fill = "white", color = "white") +
    annotate("text", x = -0.65, y = 1.15, label = "P", size = 5.5, angle = 0, fontface = "bold") +
    annotate("text", x = 4, y = 1.15, label = "chase", size = 5) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    geom_point(data = jung, aes(time, media_frac_max, fill = "temp"),
               size = 5, shape = 21, color = "white",
               show.legend = TRUE) +
    scale_fill_manual(labels = "X", values = cb_pal[2]) +

    ylim(c(-0.05, 1.2)) +
    xlim(c(-1, 10)) +
    #scale_color_viridis_b() +
    xlab("Time (h)") +
    ylab("X (Normalized to 8h)") +
    labs(color = "log2(Cost)",
         fill = "Observed data") +
    theme1 +
    theme(plot.title = element_text(size = 16))
}

pulsechase_opt_X_timecourse_unfiltered <- pulsechase_plot_X_frac_timecourse(pulsechase_full_df) +
  ggtitle("Pulse-chase extracellular sFLT1 (X) - All fits")

# save plot
ggsave(out_png("pulsechase_opt_X_timecourse_unfiltered"),
       pulsechase_opt_X_timecourse_unfiltered, width = 8, height = 6)

pulsechase_opt_X_timecourse_unfiltered
```


#### Pulse-chase I

```{r}
pulsechase_plot_I_fc_timecourse <- function(df){
  df %>%
    ggplot(aes(t, I_fc)) +
    annotate("rect", xmin = -20/60, xmax = 0, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    geom_line(aes(group = run, color = log2(total_cost)), alpha = 0.1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    geom_point(data = jung, aes(time, lysate_fc, fill = "temp"),
               size = 5, shape = 21, color = "white", show.legend = TRUE) +
    scale_fill_manual(labels = "I", values = cb_pal[1]) +
    #scale_color_viridis_b() +
    xlab("Time (h)") +
    ylab("I (Normalized to 0h)") +
    labs(color = "log2(Cost)",
         fill = "") +
    #ylim(c(-0.05, 5)) +
    xlim(c(-1, 10)) +
    theme1 +
    theme(plot.title = element_text(size = 16))
}

pulsechase_opt_I_timecourse_unfiltered <- pulsechase_plot_I_fc_timecourse(pulsechase_full_df) +
    ggtitle("Pulse-chase intracellular sFlt1 (I) - All fits") +
    annotate("segment", x = -0.7, xend = -0.17, y = 1.05, yend = 0.65, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 1.2, yend = 1.2, 
             lty = 2, size = 0.5) +
    annotate("rect", xmin = 3, xmax = 5, ymin = 1.18, ymax = 1.21, fill = "white", color = "white") +
    annotate("text", x = -0.65, y = 1.15, label = "P", size = 5.5, angle = 0, fontface = "bold") +
    annotate("text", x = 4, y = 1.2, label = "chase", size = 5)
    #annotate("text", x = -0.65, y = 0.25, label = "Start pulse", angle = 90, size = 6) +
    #annotate("text", x = 0.25, y = 0.5, label = "Start chase", angle = 90, size = 6)

# save plot
ggsave(out_png("pulsechase_opt_I_timecourse_unfiltered"),
       pulsechase_opt_I_timecourse_unfiltered, width = 8, height = 6)

pulsechase_opt_I_timecourse_unfiltered
```

## Preprocessing

### Define runs to remove

#### Runs that returned NaN cost 

```{r}
runs_to_remove_na <- meta_df %>%
  filter(is.nan(c(total_cost))) %>%
  pull(run)

length(runs_to_remove_na)
```

#### Negative species - Constitutive

```{r}
# identify runs where either species goes below 0
runs_neg_constitutive <- constitutive_full_df %>%
  filter(X < 0 | I < 0) %>%
  pull(run) %>%
  unique()

length(runs_neg_constitutive)
```

#### Negative species - Pulse-chase

```{r}
# identify runs where either species goes below 0
runs_neg_pulsechase <- pulsechase_full_df %>%
  filter(X < 0 | I < 0) %>%
  pull(run) %>%
  unique()

length(runs_neg_pulsechase)
```

### Remove runs

```{r}
# remove runs from constitutive simulated time course
constitutive_sims <- constitutive_full_df %>%
  filter(! run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive) %>%
  mutate(t = round(t,3))

# remove runs from pulse-chase simulated time course
pulsechase_sims <- pulsechase_full_df %>%
  filter(!run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive) %>%
  mutate(t = round(t,3))

# remove runs from meta_base_df
meta_df_filtered <- meta_df %>%
    filter(!run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive)

# remove runs from opt_df
opt_df_filtered <- opt_df %>%
  filter(!run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive)
  
# remove prefix from parameter names in opt_df_filtered
names(opt_df_filtered) <- str_remove(names(opt_df_filtered), "^OPT_")

# remove runs from param_df
param_df_filtered <- param_df %>%
  filter(!run %in% runs_to_remove_na & !run %in% runs_neg_pulsechase & !run %in% runs_neg_constitutive)

length(unique(param_df_filtered$run))
```

## Save objects

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix <- "../../saved-data/03_candidate_models/"
dir.create(outfile_prefix, recursive = TRUE)

save(param_df, param_df_filtered, meta_df_filtered, opt_df_filtered, constitutive_sims, pulsechase_sims,
     file = paste0(outfile_prefix, "candidate_opt_sims_processed.rda"))
```