---
title: "sFLT1 secretion candidate model import"
author: "Amy Gill"
date: "2023-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file imports data from optimization of the 2 equation sFLT1 DDE model to the Jung and Hornig time course datasets. These data will be processed further using the script "03_candidate_models/candidate_opt_preprocessing.Rmd" and then analyzed with the script "03_candidate_models/R/candidate_model_workspace.Rmd".

The optimization was originally performed 8/23/23 using the Matlab script "03_candidate_models/Matlab/dde_candidate_models.m" with following options: 

run_mode = "candidate_model_opt";
jung_cost_option = "omit_last_Sx_relative";
hornig_cost_option = "relative_omit3h";
kinghorn_cost_option = "all_pts_relative";
num_param_sets = 100;    (# per candidate model)

Optimizing to summary statistics only (mean at each time point) for Kinghorn optimization.

This optimization uses 4 species (I, I_deg, X, X_deg) and optimizes 7 parameters ($\alpha$, $\beta$, $\gamma$, $\delta$, $\epsilon$, $\kappa$, $\tau$) by calculating absolute cost in comparison to all of the following datasets:

- Jung X (normalized to 8h)
- Jung I (normalized to 0h)
- Hornig basal X concentration (absolute, ng/mL -> #/cell)
- Kinghorn X (normalized to 24h)
- Kinghorn I (normalized to 0h)

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Adjust parameter mapping to color palette

```{r}
# define colorblind-friendly palette
# black, red, sky blue, orange, green, yellow, dark blue, pink
cb_pal <- c("#000000", "#D55E00",  "#56B4E9", "#E69F00", "#009E73", "#F0E442", "#0072B2",  "#CC79A7")

# map parameter name to index of desired color
param_ind <- list(alpha = 3, beta = 2, gamma = 4, delta = 5, kappa = 7, tau = 8, none = 1)

# store parameter names
param_names <- names(param_ind)

# create parameter palette (list to map parameter names to values)
param_pal <- list()
for (i in 1:length(param_names)) {
  param_pal[param_names[i]] <- cb_pal[unlist(param_ind[param_names[i]])]
}

# set labels for Greek letters in plots
color_breaks <- names(param_pal)
color_labels <- parse(text = color_breaks)
```

## Import data

### Load files

```{r}
# specify file names
param_file_name <- "../../saved-data/03_candidate_models/hjk_opt_candidates_input.mat"
opt_file_name <- "../../saved-data/03_candidate_models/hjk_opt_candidates_100.mat"
model_meta_file_name <- "../../saved-data/03_candidate_models/hjk_opt_candidates_meta.csv"

# import file data
param_file <- readMat(param_file_name)
opt_file <- readMat(opt_file_name)
model_meta <- read_csv(model_meta_file_name, col_names = FALSE, show_col_types = FALSE)
```

### Construct run metadata

```{r}
# format model metadata
names(model_meta) <- c("prod_decay", "mat_delay", "internalize", "aic_param_count")
model_meta$candidate_model <- 1:nrow(model_meta)

# extract number of candidate models
num_candidate_models <- param_file$num.candidate.models[1,1]

# extract number of parameter sets
num_param_sets <- param_file$num.param.sets[1,1]

# extract total number of runs
num_total_runs <- param_file$num.total.runs[1,1]

test_that("number of runs = candidates * parameter sets",
          {expect_equal(num_candidate_models * num_param_sets, num_total_runs)})

# construct run metadata
meta_df <- data.frame(run = 1:num_total_runs,
           param_set = rep(seq_along(1:num_param_sets), num_candidate_models),
           candidate_model = unlist(lapply(1:num_candidate_models,
                                           function(x){rep(x, num_param_sets)})))

# combine run and model metadata
meta_df <- meta_df %>%
  left_join(model_meta, by = "candidate_model")
```

```{r}
# extract conversion factor to move X concentrations (ng/mL) to corresponding #/cell
conv_factor_numcell_to_ngml <- param_file$conv.factor.ngml[1,1]
```


### Extract initial parameter values

```{r}
# extract the randomly generated parameter sets
# (cleaner to import the parameter matrix from param_mat, but this is equivalent)
init_df <- data.frame(alpha = param_file$initial.alpha,
                        beta = param_file$initial.beta,
                        gamma = param_file$initial.gamma,
                        delta = param_file$initial.delta,
                        epsilon = param_file$initial.epsilon,
                        kappa = param_file$initial.kappa,
                        tau = param_file$initial.tau)

# designate these values as initial (rather than optimized)
names(init_df) <- paste0("INIT_", names(init_df))

# add index to each parameter set
init_df <- init_df %>%
  mutate(param_set = 1:nrow(init_df))

# create tidy version of param_mat
init_df_tidy <- init_df %>%
  pivot_longer(INIT_alpha:INIT_tau, names_to = "parameter", values_to = "value") %>%
  mutate(parameter = factor(str_remove(parameter, "INIT_")))
```

### Extract species and parameter names

```{r}
# extract parameter names
param_names <- levels(init_df_tidy$parameter)

# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")
```

### Extract optimized parameter values

- Optimal parameter values returned by `lsqnonlin` for each optimization run
- Cost associated with this parameter set (sum of squares returned by `lsqnonlin` using the `resnorm` output argument)
- Exit flag for optimization

```{r}
opt_df <- as.data.frame(opt_file$optimal)

# add prefix to denote these are initial values
names(opt_df) <- paste0("OPT_", param_names)

# add run index and cost
  opt_df <- opt_df %>%
    mutate(run = 1:nrow(opt_df),
           cost_hornig_x_ssd = opt_file$hornig.X.cost.ssd,
           cost_jung_x_ssd = opt_file$jung.X.cost.ssd,
           cost_jung_i_ssd = opt_file$jung.I.cost.ssd,
           cost_kinghorn_x_ssd = opt_file$kinghorn.X.cost.ssd,
           cost_kinghorn_i_ssd = opt_file$kinghorn.I.cost.ssd,
           total_cost = cost_hornig_x_ssd + cost_jung_x_ssd + cost_jung_i_ssd +
           cost_kinghorn_x_ssd + cost_kinghorn_i_ssd,
           log2_cost = log2(total_cost))

opt_df
```

### Add parameters to run metadata

```{r}
meta_df <- meta_df %>%
  left_join(init_df, by = "param_set") %>%
  left_join(opt_df, by = "run")

meta_df
```

### Set unused parameters to NA based on candidate model properties

```{r}
meta_df <- meta_df %>%
  mutate(INIT_kappa = ifelse(prod_decay == "off", NA, INIT_kappa),
         INIT_tau = ifelse(mat_delay == "no", NA, INIT_tau),
         INIT_epsilon = ifelse(internalize == "no", NA, INIT_epsilon),
         OPT_kappa = ifelse(prod_decay == "off", NA, OPT_kappa),
         OPT_tau = ifelse(mat_delay == "no", NA, OPT_tau),
         OPT_epsilon = ifelse(internalize == "no", NA, OPT_epsilon))
```

### Extract time points

```{r}
# extract Jung time points
pulsechase_t_all <- opt_file$jung.T.all[,1]

# Extract Hornig time points
constitutive_t_all <- opt_file$hornig.T.all[,1] # just the first column, since all the same
```

### Extract time courses

```{r}
# format data frames of simulated species values for time courses
make_sim_Y_df <- function(Y_mat){
  # confirm matrix has correct dimensions - warn if not
  test_that("matrix has correct dimensions",{
    expect_equal(dim(Y_mat)[2], length(species_names))
    expect_equal(dim(Y_mat)[3], num_total_runs)
    # not checking dim 1 (num time points) bc differs by model
  })
  
  # flatten matrix to data frame of time course data
  # (this isn't very "R-style", but it works, so tell Hadley to chill)
  for (i in 1:dim(Y_mat)[3]) {
    if (i == 1) {    # first run: initialize output data frame
      # extract matrix for first parameter set to a data frame
      Y_temp <- as.data.frame(Y_mat[,,1])
      names(Y_temp) <- species_names
      Y_temp$run <- 1
        
      # initialize full Y data frame to the time course from this first run
      Y_df <- Y_temp
      
    } else {    # later runs: append to data frame
      Y_temp <- as.data.frame(Y_mat[,,i])
      names(Y_temp) <- species_names
      Y_temp$run <- i
      Y_df <- rbind(Y_df, Y_temp)
    }
  }
  
  
  Y_df <- Y_df %>%
    mutate(X_numcell = X / conv_factor_numcell_to_ngml) # convert units from ng/ml to #/cell
  
  Y_df
}

# extract simulated data for Jung and Hornig
pulsechase_Y_df <- make_sim_Y_df(opt_file$jung.Y.all)
constitutive_Y_df <- make_sim_Y_df(opt_file$hornig.Y.all)
```

```{r}
constitutive_Y_df
```


### Combine time points and time courses

```{r}
test_that("Jung times and time courses have same dimension",
          {expect_equal(length(pulsechase_t_all) * num_total_runs,
                        nrow(pulsechase_Y_df))})

test_that("Hornig times and time courses have same dimension",
          {expect_equal(length(constitutive_t_all) * num_total_runs,
                        nrow(constitutive_Y_df))})

pulsechase_tcs <- cbind(t = rep(pulsechase_t_all, num_total_runs), pulsechase_Y_df)
constitutive_tcs <- cbind(t = rep(constitutive_t_all, num_total_runs), constitutive_Y_df)
```

### Combine full dataset

```{r}
pulsechase_full_df <- meta_df %>%
  left_join(pulsechase_tcs, by = "run")

constitutive_full_df <- meta_df %>%
  left_join(constitutive_tcs, by = "run")

# clean up repeat data
rm(constitutive_Y_df, constitutive_tcs, pulsechase_tcs, pulsechase_Y_df)
```

## Add normalized values

### Jung normalization

```{r}
pulsechase_I_0h <- pulsechase_full_df %>%
  filter(t==0) %>%
  select(run, I_0h = I) %>%
  mutate(rownum = 1:nrow(.)) %>%
  filter(rownum %% 2 == 0) %>%   # remove duplicated rows (not sure why unique/distinct don't work)
  select(-rownum)

pulsechase_X_8h <- pulsechase_full_df %>%
  filter(t == 8) %>%
  select(run, X_8h = X) 

pulsechase_full_df <- pulsechase_full_df %>%
  left_join(pulsechase_I_0h, by = "run", "I_0h") %>%
  left_join(pulsechase_X_8h, by = "run", "X_8h") %>%
  mutate(I_fc = I / I_0h,
         X_frac = X / X_8h)
```

### Hornig normalization

```{r}
constitutive_I_0 <- constitutive_full_df %>%
  filter(t == 0) %>%
  select(run, I_0 = I)

constitutive_X_24 <- constitutive_full_df %>%
  filter(t == 24) %>%
  select(run, X_24 = X)
```

```{r}
constitutive_full_df <- constitutive_full_df %>%
  left_join(constitutive_I_0) %>%
  left_join(constitutive_X_24) %>%
  mutate(I_fc = I / I_0,
         X_frac = X / X_24)

constitutive_full_df
```

## Save objects

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix <- "../../saved-data/03_candidate_models/"
dir.create(outfile_prefix, recursive = TRUE)

save(pulsechase_full_df, constitutive_full_df, meta_df, opt_df, conv_factor_numcell_to_ngml,
     file = paste0(outfile_prefix, "pulsechase_constitutive_opt_sims.rda"))
```
