---
title: "Optimization to sFLT1 datasets from Hornig, Jung and Kinghorn"
author: "Amy Gill"
date: "2023-08-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file analyzes data from optimization of the 2 equation sFLT1 DDE model to the Jung and Hornig time course datasets. These data were imported using the script "sFlt-model/analysis-R/draft/03_candidate_models/candidate_models_import.Rmd".

The optimization was performed 8/23/23 using the Matlab script "sFlt-model/03_candidate_models/dde_candidate_models.m with following options: 

run_mode = "candidate_model_opt";
jung_cost_option = "omit_last_Sx_relative";
hornig_cost_option = "relative_omit3h";
kinghorn_cost_option = "all_pts_relative";
num_param_sets = 100;    (# per candidate model)

Optimizing to summary statistics only (mean at each time point) for Kinghorn optimization.

This optimization uses 4 species (I, I_deg, X, X_deg) and optimizes 7 parameters ($\alpha$, $\beta$, $\gamma$, $\delta$, $\epsilon$, $\kappa$, $\tau$) by calculating absolute cost in comparison to all of the following datasets:

- Jung X (normalized to 8h)
- Jung I (normalized to 0h)
- Hornig basal X concentration (absolute, ng/mL -> #/cell)
- Kinghorn X (normalized to 24h)
- Kinghorn I (normalized to 0h)

## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Define output directory

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../../saved-figs/03_candidate_models/"
dir.create(outfile_prefix, recursive = TRUE)
```


### Import data

#### Optimization runs

```{r}
# load optimization results from 2023-08-23
# (if file is not present, run "candidate_models_import.Rmd" to create it)
load("../../../saved-data/03_candidate_models/pulsechase_constitutive_opt_sims.rda")
```

```{r}
# load filtered optimization results
# (if file is not present, run "candidate_models_import.Rmd" to create it)
load("../../../saved-data/03_candidate_models/candidate_opt_sims_processed.rda")
```


### Specify number of observations for AIC calculation

Jung X: 0 (no), **2, 4, 6,** 8 (no - normalized to 1), 10 (omitted) - 3 points
Jung I: 0 (no - normalized to 1), **2, 4, 6, 8, 10** - 5 points
Hornig X: 0 (no), 3 (omitted - assume BLQ), **6, 9, 12, 24, 48, 72** - 6 points

After summarizing, number of obs from Kinghorn

Kinghorn X: **0, 1, 2, 4, 8, 12,** 24 (no) - 6 points
Kinghorn I: 0 (no), **1, 2, 4, 8, 12, 24** - 6 points

total: n = 26

```{r}
aic_num_obs <- 26
```

## Analysis before preprocessing

### Create formatted data frames for analysis

```{r}
# param_df: a 4-column data frame
# run = run number 1:1000
# parameter = factor, parameter definition(alpha, beta, gamma, delta, epsilon, kappa, tau)
# init_opt = factor, INIT = initial guess, OPT = optimized value
# value = parameter value
param_df <- meta_df %>%
  select(run:OPT_tau) %>%
  pivot_longer(INIT_alpha:OPT_tau, names_to = "parameter", values_to = "value") %>%
  separate(parameter, into = c("init_opt", "parameter"), sep = "_") %>%
  mutate(init_opt = ifelse(init_opt == "INIT", "initial", "optimal"),
                           init_opt = factor(init_opt),
         parameter = factor(parameter,
                            levels = c("alpha", "beta", "gamma", "delta", 
                                       "epsilon", "kappa", "tau"))) 

# add cost to param_df (why not just update param_df?)
param_df <- meta_df %>%
  select(run, cost = total_cost) %>%
  right_join(param_df, by = "run")

# extract parameter names
param_names <- levels(param_df$parameter)
```

### Cost distribution

#### Density plot

```{r}
# create function to plot density function of cost
plot_cost_distr_density <- function(df){
  df %>%
    ggplot(aes(total_cost, after_stat(scaled))) +
    geom_density(fill = cb_pal[3]) +
    #scale_x_log10() +
    scale_x_continuous(trans = "log10", breaks = 10^(-2:5)) +
    theme1 +
    xlab("Sum of squared difference") +
    ylab("Scaled density")
}

# plot cost distribution
combo_cost_distr_unfiltered <- plot_cost_distr_density(meta_df) +
    ggtitle("Cost function values for optimal parameters - all fits")
  
  
combo_cost_distr_unfiltered
```

```{r}
# save plot
ggsave(out_png("combo_cost_distr_unfiltered"),
       combo_cost_distr_unfiltered, width = 9, height = 6)
```

#### Violin plot

```{r}
# create function to make violin plot of cost distribution
plot_cost_violin <- function(df){
  df %>%
    ggplot(aes(factor(" "), total_cost)) +
    geom_violin(scale = "width", fill = cb_pal[2]) +   # not trimming b/c hard min value 
    geom_boxplot(width = 0.3, fill = "white") +
    scale_y_continuous(trans = "log10", breaks = 10^(-2:5)) +
    xlab("") +
    ylab("Sum of squared difference") +
    coord_flip() +
    theme1
}

cost_violin <- plot_cost_violin(meta_df) +
  ggtitle("Violin plot of cost function values for optimal parameters")

cost_violin
```

#### Faceted by model

```{r}
meta_df %>%
    ggplot(aes(total_cost, ..scaled..)) +
    geom_density(fill = cb_pal[3]) +
  facet_wrap(~ candidate_model) +
    #scale_x_log10() +
    scale_x_continuous(trans = "log10", breaks = 10^(-2:5)) +
    theme1 +
    xlab("Sum of squared difference") +
    ylab("Scaled density")
```

## AIC analysis

### Min cost per model

```{r}
model_min_costs <- meta_df %>%
  group_by(candidate_model) %>%
  summarize(model_min_cost = min(total_cost, na.rm=TRUE))

model_min_costs
```

### Calculate AIC

n = `aic_num_obs`
k = `aic_param_count`

AIC = n log (cost/n) + 2k

Because n/k < 40, use corrected AIC (AIC_c) = AIC + 2k*(k+1)/(n-k-1)

```{r}
# calculate AIC using the lowest cost fit
aic_min_df <- model_min_costs %>%
  left_join(meta_df) %>%
  filter(total_cost == model_min_cost) %>%
  mutate(candidate_model = factor(as.character(candidate_model)),
         aic = aic_num_obs * log(model_min_cost / aic_num_obs) + 2 * aic_param_count,
         aic_c = aic + 2 * aic_param_count * (aic_param_count + 1) / (aic_num_obs - aic_param_count - 1)) %>%
  arrange(aic_c) %>%
  mutate(candidate_model = factor(candidate_model)) %>%
  select(candidate_model, aic_param_count, aic, aic_c, prod_decay:internalize, cost_hornig_x_ssd:log2_cost)

aic_min_df
```

```{r}
aic_stats_df <- model_min_costs %>%
  mutate(cost_cutoff = model_min_cost * 1.1) %>%
  left_join(meta_df) %>%
  filter(total_cost < cost_cutoff) %>%  # keep only fits within 10% of minimum cost per model %>%
  mutate(candidate_model = factor(as.character(candidate_model)),
         aic = aic_num_obs * log(model_min_cost / aic_num_obs) + 2 * aic_param_count,
         aic_c = aic + 2 * aic_param_count * (aic_param_count + 1) / 
           (aic_num_obs - aic_param_count - 1)) %>%
  group_by(candidate_model) %>%
  summarize(avg_cost_top10pct = mean(total_cost),
            avg_aic = mean(aic),
            avg_aic_c = mean(aic_c)) %>%
  arrange(avg_aic_c) 

aic_stats_df <- meta_df %>%
  select(candidate_model, prod_decay, mat_delay, internalize) %>%
  mutate(candidate_model = factor(candidate_model)) %>%
  filter(!duplicated(candidate_model)) %>%
  right_join(aic_stats_df, by = "candidate_model") %>%
  arrange(avg_aic_c)

aic_stats_df
```

### AIC barplot

```{r}
aic_stats_df_reordered <- aic_min_df %>%
  mutate(candidate_model = fct_reorder(candidate_model, aic_c))

aic_stats_df_reordered %>%
  ggplot(aes(factor(candidate_model), aic_c)) +
  geom_bar(stat = "identity") +
  xlab("Candidate model") +
  ylab("Corrected AIC") +
  theme1
```
### Cost barplot

#### Standard

```{r}
aic_df_reordered <- aic_min_df %>%
  mutate(candidate_model = fct_reorder(candidate_model, aic_c))

aic_df_reordered %>%
  ggplot(aes(factor(candidate_model), total_cost)) +
  geom_bar(stat = "identity") +
  xlab("Candidate model") +
  ylab("Cost (Residual sum of squares)") +
  theme1
```

#### Stacked

Stacked barplot showing total cost (sum of squared differences) and cost distribution across (Hornig X, Kinghorn X, Kinghorn I, Jung X, Jung I)

```{r}
aic_df_reordered %>%
  pivot_longer(cost_hornig_x_ssd:cost_kinghorn_i_ssd, 
               names_to = "dataset", values_to = "cost") %>%
  ggplot(aes(factor(candidate_model), cost, fill = dataset)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("Candidate model") +
  ylab("Cost (Residual sum of squares)") +
  theme1 +
  scale_fill_manual(values = cb_pal[3:7], labels = c("Hornig X", "Jung I", "Jung X", "Kinghorn I", "Kinghorn X"))
```

### Heatmap attempt 1

```{r}
# heatmap - just repeats model component annotation
cost_mat <- t(as.matrix(select(aic_df_reordered, cost_hornig_x_ssd: cost_kinghorn_i_ssd)))

# create barplot annotations for candidate models
aic_anno <- anno_barplot(aic_df_reordered$aic_c)
cost_anno <- anno_barplot(t(cost_mat), gp = gpar(fill = cb_pal[7:3]), height = unit(6, "cm"))

# 
model_features <- aic_df_reordered %>%
  select(prod_decay, mat_delay, internalize)


# create annotation table for model components
model_anno <- HeatmapAnnotation(cost = cost_anno,
                                AIC_c = aic_anno,
                                df = data.frame(prod_decay = aic_df_reordered$prod_decay,
                                                mat_delay = aic_df_reordered$mat_delay,
                                                internalize = aic_df_reordered$internalize),
                                col = list(prod_decay = c("off" = cb_pal[6], 
                                                          "lin_decay" = cb_pal[3],
                                                          "exp_decay" = cb_pal[4],
                                                          "delay_off" = cb_pal[5]),
                                           mat_delay = c("no" = cb_pal[6], 
                                                         "yes" = cb_pal[7]),
                                           internalize = c("no" = cb_pal[6],
                                                                "yes" = cb_pal[2])),
                                gp = gpar(col = cb_pal[1]),
                                annotation_height = c(40,20,5,5,5))



Heatmap(cost_mat,
        col = viridis(100),
        cluster_columns = FALSE,
        bottom_annotation = model_anno)
```

## Analysis with only one prod_decay variant

```{r}
# identify candidate models that test only 2 production decays: none or exponential 
reduced_models <- meta_df %>%
  filter(prod_decay %in% c("off", "exp_decay")) %>%
  pull(candidate_model) %>%
  unique()

reduced_aic_min_df <- aic_min_df %>%
  filter(candidate_model %in% reduced_models)

reduced_aic_stats_df <- aic_stats_df %>%
  filter(candidate_model %in% reduced_models)

reduced_aic_stats_df_reordered <- aic_stats_df_reordered %>%
  filter(candidate_model %in% reduced_models)

reduced_aic_df_reordered <- aic_df_reordered %>%
    filter(candidate_model %in% reduced_models) %>%
  mutate(prod_decay = (prod_decay == "exp_decay"),
         mat_delay = (mat_delay == "yes"),
         internalize = (internalize == "yes"))

reduced_aic_df_reordered 
```


```{r}
reduced_aic_min_df
```


```{r}
reduced_aic_min_df %>% 
  select(internalize, prod_decay, mat_delay) %>%
  mutate(internalize = internalize == "yes",
         prod_decay = prod_decay == "exp_decay",
         mat_delay = mat_delay == "yes") %>%
  arrange(internalize, prod_decay, mat_delay) %>%
  mutate(label = paste0("M", row_number())) %>%
  select(label, everything())
```


```{r}
reduced_aic_df_v2 <- reduced_aic_min_df %>% 
  select(internalize, prod_decay, mat_delay) %>%
  arrange(internalize, desc(prod_decay), mat_delay) %>%
  mutate(label = paste0("M", row_number())) %>%
  select(label, everything()) %>%
  left_join(reduced_aic_min_df) %>%
  select(-candidate_model) %>%
  mutate(internalize = internalize == "yes",
         prod_decay = prod_decay == "exp_decay",
         mat_delay = mat_delay == "yes") %>%
  arrange(aic_c)

reduced_aic_df_v2
```


### AIC barplot

```{r}
reduced_aic_df_v2 %>%
  ggplot(aes(label, aic_c)) +
  geom_bar(stat = "identity") +
  xlab("Candidate model") +
  ylab("Corrected AIC") +
  theme1
```

### Cost barplot

#### Standard

```{r}
reduced_aic_df_v2 %>%
  ggplot(aes(label, total_cost)) +
  geom_bar(stat = "identity") +
  xlab("Candidate model") +
  ylab("Cost (Residual sum of squares)") +
  theme1
```

#### Stacked

Stacked barplot showing total cost (sum of squared differences) and cost distribution across (Hornig X, Kinghorn X, Kinghorn I, Jung X, Jung I)

```{r}
reduced_aic_df_v2 %>%
  pivot_longer(cost_hornig_x_ssd:cost_kinghorn_i_ssd, 
               names_to = "dataset", values_to = "cost") %>%
  ggplot(aes(label, cost, fill = dataset)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("Candidate model") +
  ylab("Cost (Residual sum of squares)") +
  theme1 +
  scale_fill_manual(name = "Cost (SSD)", values = cb_pal[3:7], 
                    labels = c("Hornig X", "Jung I", "Jung X", "Kinghorn I", "Kinghorn X"))
```

### Heatmap attempt 2


```{r}
component_mat <- reduced_aic_df_v2 %>%
  select(internalize:mat_delay) %>%
  as.matrix() %>%
  t()

colnames(component_mat) <- reduced_aic_df_v2$label

component_mat
```


```{r}
# heatmap - just repeats model component annotation
component_num_mat <- matrix(as.integer(component_mat), nrow = nrow(component_mat))

colnames(component_num_mat) <- reduced_aic_df_v2$label

component_num_mat
```


```{r}
# heatmap - just repeats model component annotation
cost_mat <- t(as.matrix(select(reduced_aic_df_v2, cost_hornig_x_ssd: cost_kinghorn_i_ssd)))

# create barplot annotations for candidate models
aic_anno <- anno_barplot(reduced_aic_df_v2$aic_c, baseline = -55)
cost_anno <- anno_barplot(t(cost_mat), gp = gpar(fill = cb_pal[7:3]), height = unit(6, "cm"))

# 
model_features <- reduced_aic_df_v2 %>%
  select(prod_decay, mat_delay, internalize)

# create annotation table for model components
model_anno <- HeatmapAnnotation(cost = cost_anno,
                                AIC_c = aic_anno,
                                components = t(component_mat),
                                gp = gpar(col = cb_pal[1]),
                                annotation_height = c(40,20,15))

ht_list <- Heatmap(cost_mat,
        col = viridis(100),
        cluster_columns = FALSE,
        bottom_annotation = model_anno)

ht_list
```




```{r}
model_anno_2 <- HeatmapAnnotation(cost = cost_anno,
                                AIC_c = aic_anno,
                                annotation_height = c(40,20),
                                annotation_label = gt_render(
                                  c("Cost<br>(SSD)",
                                    "AIC<sub>c</sub>"),
                                  gp = gpar(fontsize = 16)
                                ))

ht_list_2 <- Heatmap(component_num_mat, col = c("white", "gray30"),
                     row_labels = expression(epsilon, kappa, tau),
                     column_labels = colnames(component_num_mat),
                     column_names_rot = 45,
                     row_names_gp = gpar(fontsize = 24),
                     column_names_gp = gpar(fontsize = 20),
                     heatmap_legend_param = list(
                       labels = c("Yes", "No"),
                       title = "Process\nincluded",
                       border = cb_pal[1]),
        cluster_columns = FALSE,
        cluster_rows = FALSE,
        rect_gp = gpar(col = cb_pal[1], lwd = 2),
        bottom_annotation = model_anno_2)

lgd_anno <- Legend(labels = c("Hornig X", "Jung I", "Jung X", 
                              "Kinghorn I", "Kinghorn X"),
                   title = "Dataset", legend_gp = gpar(fill = cb_pal[3:7]))

lgd_heat <- Legend(labels = c("Yes", "No"), title = "Process\nIncluded",
                   legend_gp = gpar(fill = c("gray30", "white")),
                   border = cb_pal[1])

draw(ht_list_2, annotation_legend_list = list(lgd_heat, lgd_anno))
```

```{r}
g <- draw(ht_list_2, annotation_legend_list = list(lgd_heat, lgd_anno))
g
```


```{r}
png(out_png("gridplot"))
plot(g)
dev.off()
```

## -- BREAKPOINT --

### Cost vs number of parameters

```{r}
reduced_aic_df_v2 %>%
  select(label, aic_param_count, total_cost)
```



```{r}
set.seed(4)
a <- 15
model_cost_v_nparams <- reduced_aic_df_v2 %>%
  ggplot(aes(aic_param_count, total_cost)) +
  geom_point(size = 3) +
  
  # manually add epsilon arrows
  annotate("segment", x = 5.1, xend = 5.9,   # M1>M5 (epsilon)
           y = 1.59, yend = 1.575, col = "gray40",
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 6.05, xend = 6.95,   # M2>M6 (epsilon)
           y = 1.09, yend = 1.04, col = "gray40",
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 6.05, xend = 6.95,   # M3>M7 (epsilon)
           y = 1.24, yend = 1.23, col = "gray40",
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 7.05, xend = 7.95,   # M4>M8 (epsilon)
           y = 1.02, yend = 0.96, col = "gray40",
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  
  # manually add kappa arrows
  annotate("segment", x = 5.05, xend = 5.95,   # M1>M3 (kappa)
           y = 1.57, yend = 1.27, col = cb_pal[7],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 6.05, xend = 6.95,   # M2>M4 (kappa)
           y = 1.1, yend = 1.06, col = cb_pal[7],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 6.05, xend = 6.95,   # M5>M7 (kappa)
           y = 1.55, yend = 1.24, col = cb_pal[7],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 7.05, xend = 7.95,   # M6>M8 (kappa)
           y = 1.035, yend = 0.96, col = cb_pal[7],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  
  # manually add tau arrows
  annotate("segment", x = 5.05, xend = 5.95,   # M1>M2 (tau)
           y = 1.57, yend = 1.13, col = cb_pal[8],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 6.05, xend = 6.95,   # M3>M4 (tau)
           y = 1.24, yend = 1.06, col = cb_pal[8],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 6.05, xend = 6.95,   # M5>M6 (tau)
           y = 1.55, yend = 1.05, col = cb_pal[8],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("segment", x = 7.05, xend = 7.95,   # M7>M8 (tau)
           y = 1.21, yend = 0.96, col = cb_pal[8],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  
  # manually add text labels in desired spots
  annotate("text", x = 4.95, y = 1.53, label = "M1", size = 5) +
  annotate("text", x = 5.8, y = 1.09, label = "M2", size = 5) +
  annotate("text", x = 6.2, y = 1.28, label = "M3", size = 5) +
  annotate("text", x = 7.25, y = 1.23, label = "M4", size = 5) +
  annotate("text", x = 6.2, y = 1.59, label = "M5", size = 5) +
  annotate("text", x = 6.9, y = 0.98, label = "M6", size = 5) +
  annotate("text", x = 7.15, y = 1.08, label = "M7", size = 5) +
  annotate("text", x = 8.2, y = 0.95, label = "M8", size = 5) +

  # handmake legend
  annotate("rect", xmin = 6.8, xmax = 8.3, ymin = 1.35, ymax = 1.62,
           col = "black", fill = "white") +
  annotate("segment", x = 6.9, xend = 7.5, 
           y = 1.56, yend = 1.56, col = cb_pal[8],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("text", x = 7.9, y = 1.56, col = cb_pal[8],
           label = "+ tau", parse = "true", size = 8) +
  annotate("segment", x = 6.9, xend = 7.5, 
           y = 1.485, yend = 1.485, col = cb_pal[7],
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("text", x = 7.9, y = 1.485, col = cb_pal[7],
           label = "+ kappa", parse = "true", size = 8) +
  annotate("segment", x = 6.9, xend = 7.5, 
           y = 1.41, yend = 1.41, col = "gray40",
           arrow = arrow(angle = a, type = "closed"), size = 1) +
  annotate("text", x = 7.9, y = 1.41, col = "gray40",
           label = "+ epsilon", parse = "true", size = 8) +
  
  #geom_label_repel(aes(label = label), size = 5) +
  xlab("Number of parameters") +
  ylab("Cost (SSD)") +
  theme_big +
  theme(axis.title.y = element_text(angle = 90, vjust = 0.5, size = 16))


ggsave(out_png("model_cost_v_nparams"), model_cost_v_nparams, width = 4, height = 3)

model_cost_v_nparams
```

```{r}
reduced_aic_df_v2
```


## -- BREAKPOINT --

## Add M candidate model labels 

### Map candidate models to M labels

```{r}
model_label_map <- reduced_aic_min_df %>%
  arrange(internalize, desc(prod_decay), mat_delay) %>%
  mutate(label = paste0("M", row_number())) %>%
  select(label, candidate_model) 

model_label_map
```

### Insert model labels in simulation data frames

```{r}
pulsechase_sims_labeled <- pulsechase_sims %>%
  filter(candidate_model %in% reduced_models) %>%
  mutate(candidate_model = factor(candidate_model)) %>%
  left_join(model_label_map) %>%
  select(label, everything())

constitutive_sims_labeled <- constitutive_sims %>%
  filter(candidate_model %in% reduced_models) %>%
  mutate(candidate_model = factor(candidate_model)) %>%
  left_join(model_label_map) %>%
  select(label, everything())
```

## Time courses with experimental data

### Add fill color columns to experimental data

```{r}
jung <- jung %>%
  mutate(fill_col = factor("Jung", levels = c("Jung", "Kinghorn", "Hornig")))

kinghorn <- kinghorn %>%
  mutate(fill_col = factor("Kinghorn", levels = c("Jung", "Kinghorn", "Hornig")))

hornig <- hornig %>%
  mutate(fill_col = factor("Hornig", levels = c("Jung", "Kinghorn", "Hornig")))
```

### Jung X

```{r}
jung_plot_X_frac_timecourse <- function(df){
  df %>%
    ggplot(aes(t, X_frac)) +
    annotate("rect", xmin = -20/60, xmax = 0, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    #annotate("rect", xmin = 0, xmax = Inf, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    geom_line(aes(group = run), alpha = 0.1, size = 1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    #annotate("text", x = -0.75, y = 0.55, label = "Start pulse", angle = 90, size = 6) +
    #annotate("text", x = 0.35, y = 0.75, label = "Start chase", angle = 90, size = 6) +
    annotate("segment", x = -0.7, xend = -0.17, y = 1.05, yend = 0.65, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 1.14, yend = 1.14, 
             lty = 2, size = 0.5) +
    annotate("rect", xmin = 1.5, xmax = 4.5, ymin = 1.13, ymax = 1.17, fill = "white", color = "white") +
    annotate("text", x = -0.65, y = 1.15, label = "P", size = 5.5, angle = 0, fontface = "bold") +
    annotate("text", x = 3, y = 1.15, label = "chase", size = 5.5, angle = 0, fontface = "bold") +
    geom_point(data = jung, aes(time, media_frac_max, fill = fill_col),
               size = 5, shape = 21, color = "white",
               show.legend = TRUE) +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
    labs(fill = "Experimental data") +    
    scale_y_continuous(limits = c(-0.05, 1.2), breaks = seq(0, 1, by = 0.25)) +
    scale_x_continuous(limits = c(-1, 10), breaks = seq(0, 10, by = 2), expand = c(0,0)) +
    #scale_color_viridis_b() +
    xlab("Time (h)") +
    ylab(TeX(r'($\frac{X}{X_{8h}}$)')) +

    #ylab("X (Normalized to 8h)") +
    theme_big
}

pulsechase_X_tc_grid <- pulsechase_sims_labeled %>%
  jung_plot_X_frac_timecourse() +
  facet_wrap(~label, ncol = 2) +
  theme(legend.position = "none")

pulsechase_X_tc_grid

ggsave(out_png("time_courses/pulsechase_X_tc_grid"),
       pulsechase_X_tc_grid, width = 6, height = 8)
```


#### Alternate plot approach using patchwork

```{r eval=FALSE}
# alternate approach using patchwork - NOT CURRENTLY USED
plot_list <- lapply(1:8, function(x){
  
  # define model version to analyze
  model_label <- paste0("M", x)
  
  # plot sim v expt for all sims of that model version
  gg <- pulsechase_sims_labeled %>%
    filter(label == model_label) %>%
    jung_plot_X_frac_timecourse() +
    ggtitle(model_label) +
    theme(legend.position = "none")

  #save plot
  ggsave(out_png(paste0("time_courses/jung_opt_X_timecourse_big_", model_label)),
         gg, width = 4, height = 3)
  
  gg
})

wrap_plots(plot_list, ncol = 2)
```

### Jung I

```{r}
jung_plot_I_fc_timecourse <- function(df){
  df %>%
    ggplot(aes(t, I_fc)) +
    annotate("rect", xmin = -20/60, xmax = 0, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    geom_line(aes(group = run), alpha = 0.1, size = 1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    annotate("segment", x = -0.6, xend = -0.1, y = 1.5, yend = 1.3, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 1.5, yend = 1.5, lty = 2, size = 0.5) +
    annotate("rect", xmin = 5.5, xmax = 8.5, ymin = 1.45, ymax = 1.55, fill = "white", color = "white") +
    annotate("text", x = -0.65, y = 1.6, label = "P", size = 5.5, angle = 0, fontface = "bold") +
    annotate("text", x = 6.9, y = 1.5, label = "chase", size = 5.5, angle = 0, fontface = "bold") +
    geom_point(data = jung, aes(time, lysate_fc, fill = fill_col),
               size = 5, shape = 21, color = "white", show.legend = TRUE) +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
    xlab("Time (h)") +
    ylab(TeX(r'($\frac{I}{I_{0h}}$)')) +
    scale_y_continuous(limits = c(-0.05, 1.7), breaks = seq(0, 1.5, by = 0.5)) +
    scale_x_continuous(limits = c(-1, 10), breaks = seq(0, 10, by = 2), expand = c(0,0)) +
    theme_big}

pulsechase_I_tc_grid <- pulsechase_sims_labeled %>%
  jung_plot_I_fc_timecourse() +
  facet_wrap(~label, ncol = 2) +
  theme(legend.position = "none")

pulsechase_I_tc_grid

ggsave(out_png("time_courses/pulsechase_I_tc_grid"),
       pulsechase_I_tc_grid, width = 6, height = 8)
```

### Kinghorn X

All optimized parameter sets produce an identical extracellular time course.

```{r}
kinghorn_opt_X_timecourse <- constitutive_sims_labeled %>%
    ggplot(aes(t, X_frac)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
    geom_point(data = kinghorn, aes(time, X_frac_24h, fill = fill_col), size = 5, 
               shape = 21, color = "white") +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
    labs(fill = "Experimental data") +
    xlab("Time (h)") +
    ylab(TeX(r'($\frac{X}{X_{24h}}$)')) +
    facet_wrap(~label, ncol = 2) +
    theme_big

kinghorn_opt_X_timecourse_nolegend <- kinghorn_opt_X_timecourse +
  theme(legend.position = "none")

# save plot
ggsave(out_png("time_courses/kinghorn_opt_X_timecourse_big"),
       kinghorn_opt_X_timecourse, width = 6, height = 8)

# save plot
ggsave(out_png("time_courses/kinghorn_opt_X_timecourse_big_nolegend"),
       kinghorn_opt_X_timecourse_nolegend, width = 6, height = 8)

kinghorn_opt_X_timecourse
```

### Kinghorn I

In constitutive simulations, I remains at steady state for the entire experimentally observed period.

```{r}
kinghorn_opt_I_timecourse <- constitutive_sims_labeled %>%
  ggplot(aes(t, I_fc)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
  geom_point(data = kinghorn, aes(time, I_fc_0h, fill = fill_col), size = 5, 
               shape = 21, color = "white") +
  scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
  xlab("Time (h)") +
    ylab(TeX(r'($\frac{I}{I_{0h}}$)')) +
    ylim(c(0, 1.5)) +
  facet_wrap(~label, ncol = 2) +
    theme_big

# save plot
ggsave(out_png("time_courses/kinghorn_opt_I_timecourse"),
       kinghorn_opt_I_timecourse, width = 6, height = 8)

kinghorn_opt_I_timecourse
```


### Hornig X

```{r}
hornig_plot_X_timecourse_numcell <- function(df){
  df %>%
    ggplot(aes(t, X / conv_factor_numcell_to_ngml)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
    geom_point(data = hornig, aes(nominal_time, X_numcell, fill = fill_col), size = 5, 
               shape = 21, color = "white") +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
    xlab("Time (h)") +
    ylab("X\n(#/cell)") +
    theme_paper +
    scale_x_continuous(breaks = seq(0, 72, by = 24))
}

hornig_opt_X_timecourse_numcell <- hornig_plot_X_timecourse_numcell(constitutive_sims_labeled) +
  facet_wrap(~label, ncol = 2) #+
  #ggtitle("Hornig extracellular sFLT1 (X) - All fits")

hornig_opt_X_timecourse_numcell_nolegend <- hornig_opt_X_timecourse_numcell +
  theme_paper +
  theme(legend.position = "none")

ggsave(out_png("time_courses/hornig_opt_X_timecourse_numcell_nolegend"),
       hornig_opt_X_timecourse_numcell_nolegend, width = 6, height = 8)

hornig_opt_X_timecourse_numcell_nolegend
```

The simulations track well with the data through 48h, but the 72h point is not well fit.

### Constitutive combo X

```{r}
const_opt_X_timecourse_combo <- constitutive_sims_labeled %>%
    ggplot(aes(t, X / X_24)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
    geom_point(data = hornig, aes(nominal_time, X_norm_24, fill = fill_col), 
               size = 5, shape = 21, color = "white") +
#    geom_line(data = hornig, aes(nominal_time, X_norm_24), linewidth = 1, color = cb_pal[2]) 
  geom_point(data = kinghorn, aes(time, X_frac_24h, fill = fill_col), 
             size = 5, shape = 21, color = "white", show.legend = FALSE) +
#  geom_line(data = kinghorn, aes(time, X_frac_24h), linewidth = 1, color = cb_pal[7]) +
  scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
  xlab("Time (h)") +
    theme_big +
  facet_wrap(~label, ncol = 2) +
    scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_y_continuous(name = TeX(r'($\frac{X}{X_{24h}}$)'),
                     sec.axis = sec_axis(trans = ~.*hornig_X_numcell_24,
                                         name = "X\n(#/cell)")) +
  theme_paper +
  theme(axis.title.y.right = element_text(angle = 0, vjust = 0.5, size = 16))

ggsave(out_png("time_courses/const_opt_X_timecourse_combo_big"),
       const_opt_X_timecourse_combo, width = 6, height = 8)

const_opt_X_timecourse_combo
```

### Combined figure

```{r}
p1 <- pulsechase_X_tc_grid
p2 <- pulsechase_I_tc_grid
p3 <- const_opt_X_timecourse_combo
p4 <- kinghorn_opt_I_timecourse

candidate_combo_plot <- (p1 + p3) / (p2 + p4) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(plot.title = element_blank(), 
        legend.position = "bottom",
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(angle = 0, vjust = 0.5, size = 20),
        axis.title.y.right = element_text(angle = 0, vjust = 0.5, size = 20),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18)) 

ggsave(out_png("time_courses/candidate_combo_plot"),
       candidate_combo_plot, width = 12, height = 16)

candidate_combo_plot
```

