---
title: "sFLT1 ODE modeling - Import simulation data"
author: "Amy Gill"
date: "2023-08-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file analyzes data from optimization of the 2 equation sFLT1 ODE model to 1 pulse-chase and 2 constitutive secretion (Kinghorn, Hornig) time course datasets. These data will be analyzed using the script "01_ode_model/R/ode_opt_workspace_2023-08-14.Rmd".

The optimization was originally performed 8/14/23 using the Matlab script "01_ode_model/Matlab/ode_driver.m" with following options: 

run_mode = "opt_100";
jung_cost_option = "omit_last_Sx_relative";
hornig_cost_option = "relative_omit3h";
kinghorn_cost_option = "all_pts_relative";
num_param_sets = 100;    (# per candidate model)

Optimizing to summary statistics only (mean at each time point) for Kinghorn optimization.

This optimization uses 4 species (I, I_deg, X, X_deg) and optimizes 7 parameters ($\alpha$, $\beta$, $\gamma$, $\delta$, $\epsilon$, $\kappa$, $\tau$) by calculating absolute cost in comparison to all of the following datasets:

- Jung X (normalized to 8h)
- Jung I (normalized to 0h)
- Hornig basal X concentration (absolute, ng/mL -> #/cell)
- Kinghorn X (normalized to 24h)
- Kinghorn I (normalized to 0h)


## Setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

## Import data

### Load files

```{r}
# specify file names
param_file_name <- "../../saved-data/01_ode_model/ode_opt_input.mat"
opt_file_name <- "../../saved-data/01_ode_model/hjk_ode_opt.mat"

# import file data
param_file <- readMat(param_file_name)
opt_file <- readMat(opt_file_name)
```

### Extract initial parameter values

```{r}
# extract the randomly generated parameter sets
# (cleaner to import the parameter matrix from param_mat, but this is equivalent)
init_df <- data.frame(alpha = param_file$initial.alpha,
                      beta = param_file$initial.beta,
                      gamma = param_file$initial.gamma,
                      delta = param_file$initial.delta)

# designate these values as initial (rather than optimized)
names(init_df) <- paste0("INIT_", names(init_df))

# add index to each parameter set
init_df <- init_df %>%
  mutate(run = 1:nrow(init_df))

# create tidy version of param_mat
init_df_tidy <- init_df %>%
  pivot_longer(INIT_alpha:INIT_delta, names_to = "parameter", values_to = "value") %>%
  mutate(parameter = factor(str_remove(parameter, "INIT_")))
```

### Extract species and parameter names, key numbers

```{r}
# extract parameter names
param_names <- levels(init_df_tidy$parameter)

# extract species names
species_names <- str_replace(names(param_file$sp[,,1]), "\\.", "_")

# extract number of parameter sets
num_param_sets <- param_file$num.param.sets[1,1]

# extract conversion factor to move X concentrations (ng/mL) to corresponding #/cell
conv_factor_numcell_to_ngml <- param_file$conv.factor.ngml[1,1]
```

### Extract optimized parameter values

- Optimal parameter values returned by `lsqnonlin` for each optimization run
- Cost associated with this parameter set (sum of squares returned by `lsqnonlin` using the `resnorm` output argument)

```{r}
opt_df <- as.data.frame(opt_file$optimal)

# add prefix to denote these are initial values
names(opt_df) <- paste0("OPT_", param_names)

# add run index and cost
opt_df <- opt_df %>%
  mutate(run = 1:nrow(opt_df),
         cost_hornig_x_ssd = opt_file$hornig.X.cost.ssd,
         cost_jung_x_ssd = opt_file$jung.X.cost.ssd,
         cost_jung_i_ssd = opt_file$jung.I.cost.ssd,
         cost_kinghorn_x_ssd = opt_file$kinghorn.X.cost.ssd,
         cost_kinghorn_i_ssd = opt_file$kinghorn.I.cost.ssd,
         total_cost = cost_hornig_x_ssd + cost_jung_x_ssd + cost_jung_i_ssd +
         cost_kinghorn_x_ssd + cost_kinghorn_i_ssd,
         log2_cost = log2(total_cost))

opt_df
```

```{r}
meta_df <- init_df %>%
  left_join(opt_df, by = "run") %>%
  select(run, everything())

meta_df
```

### Extract time points

```{r}
# extract Jung time points
pulsechase_t_all <- opt_file$jung.T.all[,1]

# Extract Hornig time points
constitutive_t_all <- opt_file$hornig.T.all[,1] # just the first column, since all the same
```

### Extract time courses

```{r}
# format data frames of simulated species values for time courses
make_sim_Y_df <- function(Y_mat){
  # confirm matrix has correct dimensions - warn if not
  test_that("matrix has correct dimensions",{
    expect_equal(dim(Y_mat)[2], length(species_names))
    expect_equal(dim(Y_mat)[3], num_param_sets)
    # not checking dim 1 (num time points) bc differs by model
  })
  
  # flatten matrix to data frame of time course data
  # (this isn't very "R-style", but it works, so tell Hadley to chill)
  for (i in 1:dim(Y_mat)[3]) {
    if (i == 1) {    # first run: initialize output data frame
      # extract matrix for first parameter set to a data frame
      Y_temp <- as.data.frame(Y_mat[,,1])
      names(Y_temp) <- species_names
      Y_temp$run <- 1
        
      # initialize full Y data frame to the time course from this first run
      Y_df <- Y_temp
      
    } else {    # later runs: append to data frame
      Y_temp <- as.data.frame(Y_mat[,,i])
      names(Y_temp) <- species_names
      Y_temp$run <- i
      Y_df <- rbind(Y_df, Y_temp)
    }
  }
  
  
  Y_df <- Y_df %>%
    mutate(X_numcell = X / conv_factor_numcell_to_ngml) # convert units from ng/ml to #/cell
  
  Y_df
}

# extract simulated data for Jung and Hornig
pulsechase_Y_df <- make_sim_Y_df(opt_file$jung.Y.all)
constitutive_Y_df <- make_sim_Y_df(opt_file$hornig.Y.all)
```


### Combine time points and time courses

```{r}
test_that("Jung times and time courses have same dimension",
          {expect_equal(length(pulsechase_t_all) * num_param_sets,
                        nrow(pulsechase_Y_df))})

test_that("Hornig times and time courses have same dimension",
          {expect_equal(length(constitutive_t_all) * num_param_sets,
                        nrow(constitutive_Y_df))})

pulsechase_tcs <- cbind(t = rep(pulsechase_t_all, num_param_sets), pulsechase_Y_df)
constitutive_tcs <- cbind(t = rep(constitutive_t_all, num_param_sets), constitutive_Y_df)
```

### Combine full dataset

```{r}
pulsechase_full_df <- meta_df %>%
  left_join(pulsechase_tcs, by = "run")

constitutive_full_df <- meta_df %>%
  left_join(constitutive_tcs, by = "run")

# clean up repeat data
rm(constitutive_Y_df, constitutive_tcs, pulsechase_tcs, pulsechase_Y_df)
```

## Add normalized values

### Pulse-chase normalization (for comparison to Jung)

```{r}
pulsechase_I_0h <- pulsechase_full_df %>%
  filter(t==0) %>%
  select(run, I_0h = I) %>%
  mutate(rownum = 1:nrow(.)) %>%
  filter(rownum %% 2 == 0) %>%   # remove duplicated rows (not sure why unique/distinct don't work)
  select(-rownum)

pulsechase_X_8h <- pulsechase_full_df %>%
  filter(t == 8) %>%
  select(run, X_8h = X) 

pulsechase_full_df <- pulsechase_full_df %>%
  left_join(pulsechase_I_0h, by = "run", "I_0h") %>%
  left_join(pulsechase_X_8h, by = "run", "X_8h") %>%
  mutate(I_fc = I / I_0h,
         X_frac = X / X_8h)
```

### Constitutive normalization (for comparison to Kinghorn)

```{r}
constitutive_I_0 <- constitutive_full_df %>%
  filter(t == 0) %>%
  select(run, I_0 = I)

constitutive_X_24 <- constitutive_full_df %>%
  filter(t == 24) %>%
  select(run, X_24 = X)

constitutive_full_df <- constitutive_full_df %>%
  left_join(constitutive_I_0) %>%
  left_join(constitutive_X_24) %>%
  mutate(I_fc = I / I_0,
         X_frac = X / X_24)

constitutive_full_df
```

## Save objects

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix <- "../../saved-data/01_ode_model/"
dir.create(outfile_prefix, recursive = TRUE)

save(pulsechase_full_df, constitutive_full_df, meta_df, opt_df, conv_factor_numcell_to_ngml,
     file = paste0(outfile_prefix, "pulsechase_constitutive_opt_sims.rda"))
```


