---
title: "sFLT1 ODE modeling - Analysis"
author: "Amy Gill"
date: "2023-08-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file analyzes data from optimization of the 2 equation sFLT1 ODE model to 1 pulse-chase and 2 constitutive secretion (Kinghorn, Hornig) time course datasets. These data were imported using the script "01_ode_model/R/ode_hjk_opt_import.Rmd".

The optimization was originally performed 8/14/23 using the Matlab script "01_ode_model/Matlab/ode_driver.m" with following options: 

run_mode = "opt_100";
jung_cost_option = "omit_last_Sx_relative";
hornig_cost_option = "relative_omit3h";
kinghorn_cost_option = "all_pts_relative";
num_param_sets = 100;    (# per candidate model)

Optimizing to summary statistics only (mean at each time point) for Kinghorn optimization.

This optimization uses 4 species (I, I_deg, X, X_deg) and optimizes 7 parameters ($\alpha$, $\beta$, $\gamma$, $\delta$, $\epsilon$, $\kappa$, $\tau$) by calculating absolute cost in comparison to all of the following datasets:

- Jung X (normalized to 8h)
- Jung I (normalized to 0h)
- Hornig basal X concentration (absolute, ng/mL -> #/cell)
- Kinghorn X (normalized to 24h)
- Kinghorn I (normalized to 0h)


## Setup

### Common setup

```{r}
# load packages, define helper functions, define color palettes, load experimental data
source("../../helper_scripts/setup.R")
```

### Add "big" theme

```{r}
theme_big <- theme1 +
  theme(legend.position = "bottom",
        axis.title.y = element_text(angle = 0, vjust = 0.5, size = 16),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 10))
```


### Set output file

```{r}
# define folder where output will be saved (for example, png images from ggsave)
outfile_prefix = "../../saved-figs/01_ode_model/"
dir.create(outfile_prefix, recursive = TRUE)
```

### Import optimization runs

```{r}
# load optimization of ODEs to Hornig, Jung, Kinghorn time course datasets
# (if file is not present, run "ode_hjk_opt_import.Rmd" to create it)
load("../../saved-data/01_ode_model/pulsechase_constitutive_opt_sims.rda")
```

### Plot experimental data

Provides examples for how to plot these datasets

#### Hornig X

```{r}
hornig %>%
    ggplot(aes(nominal_time, X_norm_24)) +
    geom_point(size = 5, color = cb_pal[2]) +
    geom_line(linewidth = 1, color = cb_pal[2]) +
    xlab("Time (h)") +
    ggtitle("Extracellular sFLT1") +
    theme1 +
    scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_y_continuous(name = "Fold change vs 24h",
                     sec.axis = sec_axis(trans = ~.*hornig_X_numcell_24,
                                         name = "#/cell"))
```

#### Jung I, X

```{r}
# plot lysate fold change over time
p1 <- jung %>%
  ggplot(aes(time, lysate_fc)) +
  geom_point(size = 5, color = cb_pal[2]) +
  geom_line(size = 1, color = cb_pal[2]) +
  xlab("Time (h)") +
  ylab("Fold change vs 0h") +
  ggtitle("Intracellular sFLT1 (I)") +
  scale_x_continuous(breaks = seq(0, 10, by = 2)) +
  theme1

# plot media fraction maximum over time
p2 <- jung %>%
  ggplot(aes(time, media_frac_max)) +
  geom_point(size = 5, color = cb_pal[2]) +
  geom_line(size = 1, color = cb_pal[2]) +
  xlab("Time (h)") +
  ylab("Fold change vs 8h") +
  ggtitle("Extracellular sFLT1 (X)") +
  scale_x_continuous(breaks = seq(0, 10, by = 2)) +
  theme1

jung_expt_timecourse <- p2 / p1

jung_expt_timecourse
```

#### Kinghorn I, X

Important: values are mean values

```{r}
# plot lysate fold change over time
k_p1 <- kinghorn %>%
  ggplot(aes(time, I_fc_0h)) +
  geom_point(size = 5, color = cb_pal[2]) +
  geom_line(size = 1, color = cb_pal[2]) +
  xlab("Time (h)") +
  ylab("Fold change vs 0h") +
  ggtitle("Intracellular sFLT1 (I)") +
  scale_x_continuous(breaks = seq(0, 24, by=4)) +
  theme1 +
  ylim(0,1.5)

# plot media fraction maximum over time
k_p2 <- kinghorn %>%
  ggplot(aes(time, X_frac_24h)) +
  geom_point(size = 5, color = cb_pal[2]) +
  geom_line(size = 1, color = cb_pal[2]) +
  xlab("Time (h)") +
  ylab("Fold change vs 24h") +
  ggtitle("Extracellular sFLT1 (X)") +
  scale_x_continuous(breaks = seq(0, 24, by=4)) +
  theme1

kinghorn_expt_timecourse <- k_p2 / k_p1

kinghorn_expt_timecourse
```

#### Constitutive datasets combined

```{r}
hornig %>%
    ggplot(aes(nominal_time, X_norm_24)) +
    geom_point(size = 5, color = cb_pal[2]) +
    geom_line(linewidth = 1, color = cb_pal[2]) +
  geom_point(data = kinghorn, aes(time, X_frac_24h), size = 5, color = cb_pal[7]) +
  geom_line(data = kinghorn, aes(time, X_frac_24h), linewidth = 1, color = cb_pal[7]) +
    xlab("Time (h)") +
    ggtitle("Extracellular sFLT1") +
    theme1 +
    scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_y_continuous(name = "Fold change vs 24h",
                     sec.axis = sec_axis(trans = ~.*hornig_X_numcell_24,
                                         name = "#/cell"))
```

## Preprocessing

### Format data frames for analysis

```{r}
# param_df: a 4-column data frame
# run = run number 1:1000
# parameter = factor, parameter definition(alpha, beta, gamma, delta)
# init_opt = factor, INIT = initial guess, OPT = optimized value
# value = parameter value
param_df <- meta_df %>%
  select(run:OPT_delta) %>%
  pivot_longer(INIT_alpha:OPT_delta, names_to = "parameter", values_to = "value") %>%
  separate(parameter, into = c("init_opt", "parameter"), sep = "_") %>%
  mutate(init_opt = ifelse(init_opt == "INIT", "initial", "optimal"),
                           init_opt = factor(init_opt),
         parameter = factor(parameter,
                            levels = c("alpha", "beta", "gamma", "delta"))) 

# add cost to param_df
param_df <- meta_df %>%
  select(run, cost = total_cost) %>%
  right_join(param_df, by = "run")

# extract parameter names
param_names <- levels(param_df$parameter)
```

### Cost distribution

#### Density plot

```{r}
# create function to plot density function of cost
plot_cost_distr_density <- function(df){
  df %>%
    ggplot(aes(total_cost, ..scaled..)) +
    geom_density(fill = cb_pal[3]) +
    #scale_x_log10() +
    #scale_x_continuous(trans = "log10", breaks = 10^(-2:5)) +
    theme1 +
    xlab("Sum of squared difference") +
    ylab("Scaled density")
}

# plot cost distribution
cost_distr <- plot_cost_distr_density(meta_df) #+
    #ggtitle("Cost function values for optimal parameters - all fits")
  
  
cost_distr
```

All n=100 optimization runs for the ODE model converged to the same cost (mean 1.591, SD 2.153e-7), defined as the sum of squared differences between simulated values and experimental observations at fixed time points.

(better to just report stats than to show the SSD density plot above or violin plot below)

```{r}
mean(meta_df$total_cost)
sd(meta_df$total_cost)
```

#### Violin plot

(similar to cost density plot, better to just report stats)

```{r}
# create function to make violin plot of cost distribution
plot_cost_violin <- function(df){
  df %>%
    ggplot(aes(factor(" "), total_cost)) +
    geom_violin(scale = "width", fill = cb_pal[2]) +   # not trimming b/c hard min value 
    geom_boxplot(width = 0.3, fill = "white") +
    #scale_y_continuous(trans = "log10", breaks = 10^(-2:5)) +
    xlab("") +
    ylab("Sum of squared difference") +
    coord_flip() +
    theme1
}

cost_violin <- plot_cost_violin(meta_df) #+
  #ggtitle("Violin plot of cost function values for optimal parameters")

cost_violin
```

## Time courses with experimental data

### Add fill color columns to experimental data

```{r}
jung <- jung %>%
  mutate(fill_col = factor("Jung", levels = c("Jung", "Kinghorn", "Hornig")))

kinghorn <- kinghorn %>%
  mutate(fill_col = factor("Kinghorn", levels = c("Jung", "Kinghorn", "Hornig")))

hornig <- hornig %>%
  mutate(fill_col = factor("Hornig", levels = c("Jung", "Kinghorn", "Hornig")))
```

### Jung X

```{r}
jung_plot_X_frac_timecourse <- function(df){
  df %>%
    ggplot(aes(t, X_frac)) +
    annotate("rect", xmin = -20/60, xmax = 0, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    #annotate("rect", xmin = 0, xmax = Inf, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    geom_line(aes(group = run), alpha = 0.1, size = 1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    #annotate("text", x = -0.75, y = 0.55, label = "Start pulse", angle = 90, size = 6) +
    #annotate("text", x = 0.35, y = 0.75, label = "Start chase", angle = 90, size = 6) +
    annotate("segment", x = -0.7, xend = -0.17, y = 1.05, yend = 0.65, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 1.14, yend = 1.14, 
             lty = 2, size = 0.5) +
    annotate("rect", xmin = 2.9, xmax = 5.1, ymin = 1.11, ymax = 1.19, fill = "white", color = "white") +
    annotate("text", x = -0.65, y = 1.15, label = "P", size = 5.5, angle = 0, fontface = "bold") +
    annotate("text", x = 4, y = 1.15, label = "chase", size = 7, family = "serif") +
    geom_point(data = jung, aes(time, media_frac_max, fill = fill_col),
               size = 5, shape = 21, color = "white",
               show.legend = TRUE) +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
    labs(fill = "Experimental data") +    
    scale_y_continuous(limits = c(-0.05, 1.2), breaks = seq(0, 1, by = 0.25)) +
    scale_x_continuous(limits = c(-1, 10), breaks = seq(0, 10, by = 2), expand = c(0.01,0.01)) +
    #scale_color_viridis_b() +
    xlab("Time (h)") +
    ylab(TeX(r'($\frac{X}{X_{8h}}$)')) +

    #ylab("X (Normalized to 8h)") +
    theme_big
}

jung_opt_X_timecourse <- jung_plot_X_frac_timecourse(pulsechase_full_df) #+
  #ggtitle("Jung extracellular sFLT1 (X) - All fits")

jung_opt_X_timecourse_no_legend <- jung_plot_X_frac_timecourse(pulsechase_full_df) +
  theme(legend.position = "none")

jung_opt_X_timecourse_poster <- jung_opt_X_timecourse + theme_poster

# save plot
ggsave(out_png("jung_opt_X_timecourse_big"),
       jung_opt_X_timecourse, width = 4, height = 3)

# save plot
ggsave(out_png("jung_opt_X_timecourse_poster"),
       jung_opt_X_timecourse_poster, width = 4, height = 3)

# save plot without legend
ggsave(out_png("jung_opt_X_timecourse_big_nolegend"),
               jung_opt_X_timecourse_no_legend, width = 4, height = 3)

jung_opt_X_timecourse_poster
```

### Jung I

```{r}
jung_plot_I_fc_timecourse <- function(df){
  df %>%
    ggplot(aes(t, I_fc)) +
    annotate("rect", xmin = -20/60, xmax = 0, ymin = -Inf, ymax = Inf, fill = cb_pal[6]) +
    geom_line(aes(group = run), alpha = 0.1, size = 1) +
        geom_vline(xintercept = 0) +
    geom_vline(xintercept = -20/60) +
    annotate("segment", x = -0.7, xend = -0.2, y = 1.05, yend = 0.7, size = 0.5) +
    annotate("segment", x = 0.05, xend = 9.95, y = 0, yend = 0, lty = 2, size = 0.5) +
    annotate("rect", xmin = 2.9, xmax = 5.1, ymin = -.02, ymax = .02, fill = "white", color = "white") +
    annotate("text", x = -0.65, y = 1.15, label = "P", size = 5.5, angle = 0, fontface = "bold") +
    annotate("text", x = 4, y = 0.01, label = "chase", size = 7, family = "serif") +
    geom_point(data = jung, aes(time, lysate_fc, fill = fill_col),
               size = 5, shape = 21, color = "white", show.legend = TRUE) +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
    
    xlab("Time (h)") +
    ylab(TeX(r'($\frac{I}{I_{0h}}$)')) +
    scale_y_continuous(limits = c(-0.05, 1.2), breaks = seq(0, 1, by = 0.25)) +
    scale_x_continuous(limits = c(-1, 10), breaks = seq(0, 10, by = 2), expand = c(0.01,0.01)) +
    theme_big}

jung_opt_I_timecourse <- jung_plot_I_fc_timecourse(pulsechase_full_df)
    #ggtitle("Jung intracellular sFlt1 (I) - All fits") +
    
jung_opt_I_timecourse_nolegend <- jung_opt_I_timecourse +
  theme(legend.position = "none")

jung_opt_I_timecourse_poster <- jung_opt_I_timecourse +
  theme_poster

# save plot
ggsave(out_png("jung_opt_I_timecourse_big"),
       jung_opt_I_timecourse, width = 4, height = 3)

# save plot
ggsave(out_png("jung_opt_I_timecourse_big_nolegend"),
       jung_opt_I_timecourse_nolegend, width = 4, height = 3)

# save plot
ggsave(out_png("jung_opt_I_timecourse_poster"),
       jung_opt_I_timecourse_poster, width = 4, height = 3)

jung_opt_I_timecourse_poster
```

### Kinghorn X

All optimized parameter sets produce an identical extracellular time course.

```{r}
kinghorn_opt_X_timecourse <- constitutive_full_df %>%
  ggplot(aes(t, X_frac)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
  geom_point(data = kinghorn, aes(time, X_frac_24h, fill = fill_col), size = 5, 
               shape = 21, color = "white") +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
    labs(fill = "Experimental data") +
    xlab("Time (h)") +
      scale_x_continuous(breaks = seq(0, 72, by = 24)) +
    ylab(TeX(r'($\frac{X}{X_{24h}}$)')) +
    theme_big

kinghorn_opt_X_timecourse_nolegend <- kinghorn_opt_X_timecourse +
  theme(legend.position = "none")

kinghorn_opt_X_timecourse_poster <- kinghorn_opt_X_timecourse + theme_poster

# save plot
ggsave(out_png("kinghorn_opt_X_timecourse_big"),
       kinghorn_opt_X_timecourse, width = 4, height = 3)

# save plot
ggsave(out_png("kinghorn_opt_X_timecourse_big_nolegend"),
       kinghorn_opt_X_timecourse_nolegend, width = 4, height = 3)

# save plot
ggsave(out_png("kinghorn_opt_X_timecourse_poster"),
       kinghorn_opt_X_timecourse_poster, width = 4, height = 3)

kinghorn_opt_X_timecourse_poster
```

### Kinghorn I

In constitutive simulations, I remains at steady state for the entire experimentally observed period.

```{r}
kinghorn_opt_I_timecourse <- constitutive_full_df %>%
  ggplot(aes(t, I_fc)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
  geom_point(data = kinghorn, aes(time, I_fc_0h, fill = fill_col), size = 5, 
               shape = 21, color = "white") +
  scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
  xlab("Time (h)") +
    ylab(TeX(r'($\frac{I}{I_{0h}}$)')) +
      scale_x_continuous(breaks = seq(0, 72, by = 24)) +

    ylim(c(0, 1.5)) +
    theme_big

kinghorn_opt_I_timecourse_poster <- kinghorn_opt_I_timecourse + theme_poster

# save plot
ggsave(out_png("kinghorn_opt_I_timecourse"),
       kinghorn_opt_I_timecourse, width = 4, height = 3)

ggsave(out_png("kinghorn_opt_I_timecourse_poster"),
       kinghorn_opt_I_timecourse_poster)

kinghorn_opt_I_timecourse_poster
```


### Hornig X

```{r}
hornig_plot_X_timecourse_numcell <- function(df){
  df %>%
    ggplot(aes(t, X / conv_factor_numcell_to_ngml)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
    geom_point(data = hornig, aes(nominal_time, X_numcell, fill = fill_col), size = 5, 
               shape = 21, color = "white") +
    scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
        scale_x_continuous(breaks = seq(0, 72, by = 24)) +

    xlab("Time (h)") +
    ylab("X\n(#/cell)") +
    theme_big
}

hornig_opt_X_timecourse_numcell <- hornig_plot_X_timecourse_numcell(constitutive_full_df) #+
  #ggtitle("Hornig extracellular sFLT1 (X) - All fits")

hornig_opt_X_timecourse_numcell_nolegend <- hornig_opt_X_timecourse_numcell +
  theme_big +
  theme(legend.position = "none")

hornig_opt_X_timecourse_poster <- hornig_opt_X_timecourse_numcell + theme_poster

ggsave(out_png("hornig_opt_X_timecourse_numcell_nolegend"),
       hornig_opt_X_timecourse_numcell_nolegend, width = 4, height = 3)

ggsave(out_png("hornig_opt_X_timecourse_numcell_poster"),
       hornig_opt_X_timecourse_poster, width = 4, height = 3)

hornig_opt_X_timecourse_poster
```

The simulations track well with the data through 48h, but the 72h point is not well fit.

### Hornig I

```{r}
hornig_plot_I_timecourse_numcell <- function(df){
  df %>%
    ggplot(aes(t, I)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
    xlab("Time (h)") +
    ylab("I\n(#/cell)") +
    theme_big +
    scale_x_continuous(breaks = seq(0, 72, by = 24)) +
    scale_y_log10(limits = c(5e4, 5e7))
}

hornig_opt_I_timecourse_numcell <- hornig_plot_I_timecourse_numcell(constitutive_full_df) +
  theme_paper +
  theme(legend.position = "none")
#  ggtitle("Hornig intracellular sFLT1 (I) - All fits")

ggsave(out_png("hornig_opt_I_timecourse_numcell"),
       hornig_opt_I_timecourse_numcell, width = 4, height = 3)

hornig_opt_I_timecourse_numcell
```

The absolute number of intracellular sFLT1 varies across parameter sets within a range of 1.9e5 - 2.1e7 molecules per cell, with a median of 7.6e5 #/cell.

```{r}
range(constitutive_full_df$I)
median(constitutive_full_df$I)
```


### Constitutive combo X

```{r}
const_opt_X_timecourse_combo <- constitutive_full_df %>%
    ggplot(aes(t, X / X_24)) +
    geom_line(aes(group = run), size = 1, alpha = 0.25) +
    geom_point(data = hornig, aes(nominal_time, X_norm_24, fill = fill_col), 
               size = 5, shape = 21, color = "white") +
#    geom_line(data = hornig, aes(nominal_time, X_norm_24), linewidth = 1, color = cb_pal[2]) 
  geom_point(data = kinghorn, aes(time, X_frac_24h, fill = fill_col), 
             size = 5, shape = 21, color = "white", show.legend = FALSE) +
#  geom_line(data = kinghorn, aes(time, X_frac_24h), linewidth = 1, color = cb_pal[7]) +
  scale_fill_manual(labels = c("Jung", "Kinghorn", "Hornig"),
                    values = c(cb_pal[5], cb_pal[2], cb_pal[7]), drop = FALSE) +  
  labs(fill = "Experimental data") +
  xlab("Time (h)") +
    theme_paper +
    scale_x_continuous(breaks = seq(0, 72, by = 24)) +
  scale_y_continuous(name = TeX(r'($\frac{X}{X_{24h}}$)'),
                     sec.axis = sec_axis(trans = ~.*hornig_X_numcell_24,
                                         name = "X\n(#/cell)")) +
  theme(axis.title.y.right = element_text(angle = 0, vjust = 0.5, size = 16))

const_opt_X_timecourse_combo
```


```{r}
ggsave(out_png("const_opt_X_timecourse_combo_big"),
       const_opt_X_timecourse_combo, width = 4, height = 3)

const_opt_X_timecourse_combo_poster <- const_opt_X_timecourse_combo + theme_poster

ggsave(out_png("const_opt_X_timecourse_combo_poster"),
       const_opt_X_timecourse_combo_poster, width = 4, height = 3)

const_opt_X_timecourse_combo_poster
```

### Combined figure

```{r}
p1 <- jung_opt_X_timecourse
p2 <- jung_opt_I_timecourse
p3 <- const_opt_X_timecourse_combo
p4 <- kinghorn_opt_I_timecourse

ode_combo_plot <- (p1 + p3) / (p2 + p4) +
  plot_layout(guides = "collect") &
  theme(plot.title = element_blank(), 
        legend.position = "bottom",
        axis.title.x = element_text(size = 24),
        axis.title.y = element_text(angle = 0, vjust = 0.5, size = 24),
        axis.title.y.right = element_text(angle = 0, vjust = 0.5, size = 24),
        axis.text = element_text(size = 20),
        legend.title = element_text(size = 24, face = "bold"),
        legend.text = element_text(size = 24))

ggsave(out_png("ode_combo_plot"),
       ode_combo_plot, width = 12, height = 8)

ode_combo_plot
```

#### Poster version

```{r}
p1 <- jung_opt_X_timecourse_poster
p2 <- jung_opt_I_timecourse_poster
p3 <- const_opt_X_timecourse_combo_poster
p4 <- kinghorn_opt_I_timecourse_poster

ode_combo_plot_poster <- (p1 + p3) / (p2 + p4) +
  plot_layout(guides = "collect") &
  theme(plot.title = element_blank(), 
        legend.position = "bottom",
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(angle = 0, vjust = 0.5, size = 20),
        axis.title.y.right = element_text(angle = 0, vjust = 0.5, size = 20),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))

ggsave(out_png("ode_combo_plot_poster"),
       ode_combo_plot_poster, width = 9, height = 6)

ode_combo_plot_poster
```







